### 2021-09-07 1.0-change_by_probes.rmd
first commit on nfs

---
title: "Methyl Race Patterns: Probe change frequency"
author: "Jonah Fisher"
date: "7/9/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: sandstone
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)

library(minfi)
library(tidyverse)
library(reshape2)
library(plotly)
```

# Get data and filter down to the samples of interest

```{r create data matrices, eval = F}
beta <- readRDS("../general_data/betaqc.rds")
pd <- readRDS("../general_data/pdqc.rds")
pd <- filter(pd, ck6ethrace %in% 1:3)

commonids <- Reduce(intersect, list(pd$idnum[pd$childteen == "C"], pd$idnum[pd$childteen == "T"]))

pdlist <- list(
  child = filter(pd, childteen == "C"),
  teen = filter(pd, childteen == "T")
)

pdlist$child <- pdlist$child[match(commonids, pdlist$child$idnum),]
pdlist$teen <- pdlist$teen[match(commonids, pdlist$teen$idnum),]

betalist <- list(
  child = beta[, match(pdlist$child$MethID, colnames(beta))],
  teen = beta[, match(pdlist$teen$MethID, colnames(beta))]
)

betadiffs <- betalist$teen - betalist$child
betadiffs <- abs(betadiffs)
```

# Different stratification

## Code for ethrace stratification --

```{r create dataframe of differences, eval = F}
changedf <- data.frame(
  probes = rownames(betadiffs),
  All = rowMeans(betadiffs),
  NHW = rowMeans(betadiffs[, pdlist$child$ck6ethrace == 1]),
  NHB = rowMeans(betadiffs[, pdlist$child$ck6ethrace == 2]),
  H = rowMeans(betadiffs[, pdlist$child$ck6ethrace == 3]),
  bins = ceiling(rowMeans(betalist$child) * 10)
)

changelong <- melt(changedf, id = c("probes", "bins"))

```

```{r create data frame of numbers of probes in various change degrees by bins, eval = F}
#We want data points of bin, ethrace group, and count
countdf <- data.frame(
  bin = rep(1:10, 4),
  ethrace = rep(c("All", "NHW", "NHB", "H"), each = 10)
)

getcounts <- function(lowerb, upperb, basemat = changedf, newmat = countdf){
  counts <- apply(newmat, 1, function(rowiter){
    basemat2 <- filter(basemat, bins == as.numeric(rowiter[1]))
    sum(basemat2[rowiter[2]] > lowerb & basemat2[rowiter[2]] < upperb)
  })
  data.frame(
    newmat,
    count = counts,
    change = sprintf("%.3f to %.3f", lowerb, upperb)
  )
}

fullcountdf <- rbind(
  getcounts(0, 0.005),
  getcounts(0.005, 0.01),
  getcounts(0.01, 0.03),
  getcounts(0.03, 0.06),
  getcounts(0.06, 1)
)

saveRDS(fullcountdf, "changecounts_bysire.rds")
```

# Overall count

## Barchart

```{r ggplot of the change stacked bar}
fullcountdf <- readRDS("changecounts_bysire.rds")

p1 <- ggplot(fullcountdf, aes(x = factor(bin, levels = 1:10), y = count, fill = change)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ethrace) +
  xlab("Cutpoint bin of methylation") +
  ylab("Overall frequency of probe count")

ggplotly(p1)

library(DT)
widecountdf <- dcast(fullcountdf, bin + change ~ ethrace, value.var = "count")
write.csv(widecountdf, "probecounts_by_sire.csv", row.names = F)

```

## Data table of overall count

```{r}
datatable(widecountdf)
```

# Proportion of probes within each range of change for every given bin and sire

```{r wide count dff}
widecountdff <- widecountdf
widecountdff[, 3:6] <- lapply(colnames(widecountdff)[3:6], function(x){
  lapply(1:10, function(y){
    vec <- filter(widecountdff, bin == y) %>% select(x) %>% unlist
    freqs <- round(vec / sum(vec), digits = 5)
    names(freqs) <- unique(widecountdff$change)
    freqs
  }) %>% unlist
}) %>% bind_cols

fullcountdff <- melt(widecountdff, id.vars = c("bin", "change"), measure.vars = c("All", "NHW", "NHB", "H"))
colnames(fullcountdff) <- c("bin", "change", "ethrace", "prop")

p2 <- ggplot(fullcountdff, aes(x = factor(bin, levels = 1:10), y = prop, fill = change)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ethrace) +
  xlab("Cutpoint bin of methylation") +
  ylab("Proportion")

ggplotly(p2)

```

## Data table of proportion of probes for given bin and SIRE that are within the change bin (0.000 - 0.005, 0.005 - 0.001, etc...)

```{r}
datatable(widecountdff)
```