### 2021-09-07 Reports/FFCleaning2.Rmd

---
title: "Fragile Families Data Cleaning"
author: "Freida Blostein"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    code_folding: show
    highlight: tango
    number_sections: yes
    theme: sandstone
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/nfs/turbo/bakulski1/People/blostein/FF_methylation")
knitr::opts_knit$set(message=FALSE)
knitr::opts_knit$set(warning=FALSE)
library(tidyr)
library(stringr)
library(dplyr)
library(purrr)
library(tidyselect)
library(broom)
library(compareGroups)
library(kableExtra)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(ggplot2)
library(ggpubr)
library(GGally)
library(ggcorrplot)
library(ggalluvial)
library(cowplot)
library(DiagrammeR)
library(sva)
datadir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Data/"
john_datadir<-"/nfs/turbo/bakulski1/People/johndou/Fragile_Families/ProcessedFiles/jd"
outputdir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Output/"
codedir<-"/nfs/turbo/bakulski1/People/blostein/FF_methylation/Code/"
```

# Review of important goals

## Necessary background:

* 11% of pregnant US women reported past month cigarette use in 2018.

* Prenatal maternal smoking is associated with negative health outcomes, including low birth weight, asthama and neurodevelopmental and behavioral issues in exposed children.

* Prenatal maternal smoking is hard to measure: social desirability bias, serum cotinine is short-lived in pregnant women, and in adult cohorts, possibility of recall bias.

* Finding a reliable and specific biomarker of prenatal maternal smoking is desirable.

### Previous work

Previous work has found many associations between prenatal maternal smoking and DNA methylation, a type of epigenetic change to DNA where a methyl attaches to a CpG site.

#  Objectives and  hypotheses

A) Conduct an ancestry-stratified epigenome wide association study of prenatal maternal smoking and DNA methylation from child saliva samples at age nine and age fifteen.

B) Examine differences in methylation trajectories from age nine to fifteen of the most significant methylation sites identified in the epigenome-wide association study between those exposed vs unexposed to prenatal smoking.

C) Develop a saliva epigenetic classifier of prenatal smoke exposure using a machine learning approach and support vector machines and evaluate the agreement of the classifier at two time points.

Hypotheses:
- I hypothesize that different CpG loci will be associated with maternal prenatal smoking in African ancestry groups than in European and Hispanic ancestry groups.

- I hypothesize that the direction of associations between CpGs and prenatal maternal smoking will stay consistent between ages nine and fifteen but that the magnitude of associations will decline over time.

- Finally, I hypothesize that classification of prenatal maternal smoking status based upon DNA methylation information from salivary samples will be accurate and consistent at ages nine and fifteen.

# Reading in data and preprocessing

## Labeling phenotype data
```{r, phenotypedataReadinLabel, include=F}
pheno<-readRDS(paste0(datadir, "OGData/fullpheno.rds"))

#Phenotype labeling & filtering to variables of interest
#labels
#to set labels using sjlabelled, needed to fix multibyte characters 
#https://stackoverflow.com/questions/4993837/r-invalid-multibyte-string
find_offending_character <- function(x, maxStringLength=256){  
  print(x)
  for (c in 1:maxStringLength){
    offendingChar <- substr(x,c,c)
    #print(offendingChar) #uncomment if you want the indiv characters printed
    #the next character is the offending multibyte Character
  }    
}
#lapply(attributes(pheno)$var.labels, find_offending_character)
attributes(pheno)$var.labels<-gsub("\xad", "'",  attributes(pheno)$var.labels)
#lapply(attributes(pheno)$var.labels, find_offending_character)
attributes(pheno)$var.labels<-gsub("\x92", "'",  attributes(pheno)$var.labels)
#lapply(attributes(pheno)$var.labels, find_offending_character)
#define vars of interest 
idvars<-c("id", "ff_id")
demovars<-c("cm1age", "cm1bsex", "cm1ethrace", "cm1edu", "cm1hhinc", "cm1inpov", "m1h3", "m1h3a", "m1h3b", "m1i4", "m1i4a", "m1i4b", "cf1age", "cf1ethrace", "f1h3", "f1h3a", "f1h3b", "ch5agem","ch6yagem", "c6yagey", "cm5b_age", "cm5b_ageyrs", "cp6yagem", "cp6yagey", "hv5_agem", "hv6_yagem", "hv6_yagey")
smokingvars<-c("m1g4", "f1g4", "f2j5", "f2j5a", "f2j7", "f2j7a", "f3j31", "f3j32", "f4j18", "f4j19", "f5g17", "f5g18", "k5f1k", "k5f1l", "k6d40", "k6d41", "k6d41", "k6d42", "k6d43", "k6d45", "k6d46", "k6d47", "m2j5", "m2j5a", "m2j7", "m2j7a", "m4j18", "m4j19", "m5g17", "m5g18", "n5f17", "n5f18", "p3a22", "p3a23", "p3a23a", "p3a23b", "p3a24", "p4a22", "p4a23", "p5h15", "p5h15b", "p5q3cr", "p6h74", "p6h75", "p6h76", "p6h77", "p6h78")
prenataldrugusevar<-c("m1g2", "m1g3", "m1g5", "m1g6")

#filter to these variables
#FF has data in original form +sjlabelled value labels
#since these dont' show in compareGroups
#also will make FF_labeled, where categorical, binary variables
#will be replaced with their labels
#FF<-pheno %>% select(one_of(idvars), one_of(demovars), one_of(smokingvars), one_of(prenataldrugusevar))
varIDX<-which(names(pheno)%in%c(idvars, demovars, smokingvars, prenataldrugusevar))
FF<-pheno[,varIDX]
varLabels<-attributes(pheno)$var.labels[varIDX]
varLabels[69]<-"ID"
FF<-set_label(FF, c(varLabels))
#create FF_labeled
#read in csv with metadata about variables 
FFmeta<-read.csv(paste0(datadir, "FFMetadata_v07.csv"))
#create long data dictionary  
FFlong<-FFmeta%>% select(old_name, contains('value'), contains('label')) %>% pivot_longer(cols = contains("value"), names_to="value", values_to="old_values")
FFlong<-FFlong%>%pivot_longer(cols = contains("label"), names_to="label", values_to="new_labels")
recodes_long<-FFlong %>% filter(gsub("value", "", value)==gsub("label", "", label)& !is.na(old_values)) %>% select(old_name, old_values, new_labels) 
#store continuous, binary, categorical etc
FFmyMeta<-FFmeta %>% filter(old_name %in% c(idvars, demovars, smokingvars, prenataldrugusevar))
myContinuous<-FFmyMeta %>% filter(type=="Continuous") %>%pull(old_name);
myContinuous<-colnames(FF)[colnames(FF) %in% myContinuous]
myBinary<-FFmyMeta%>%filter(type=="Binary")%>%pull(old_name); myBinary<-colnames(FF)[colnames(FF) %in% myBinary]
myCategorical_O<-FFmyMeta%>%filter(type=="Ordered Categorical")%>%pull(old_name); myCategorical_O<-colnames(FF)[colnames(FF) %in% myCategorical_O]
myCategorical_U<-FFmyMeta%>%filter(type=="Unordered Categorical")%>%pull(old_name); myCategorical_U<-colnames(FF)[colnames(FF) %in% myCategorical_U]
#filter to only vars of interest - not continuous variables  
myRecodes<-recodes_long %>% filter(old_name %in% c(myBinary, myCategorical_O, myCategorical_U))
#apply matchmaker
FF_labeled<-match_df(FF, dictionary=myRecodes, from = 'old_values', to='new_labels', by='old_name')
#reapply correct variable class types and na_codes
na_codes<-c("-1 Refuse", "-2 Don't know", "-3 Missing", "-9 Not in wave")
FF_labeled<-FF_labeled %>% mutate_at(vars(any_of(c(myBinary, myCategorical_O, myCategorical_U))), funs(as.factor(ifelse(. %in% na_codes, "Missing", .)))) %>% 
          mutate_at(vars(any_of(myContinuous)), funs(as.numeric(na_if(na_if(na_if(na_if(., "-3"), '-9'), '-6'), '-2'))))
#read back in the variable labels 
FF_labeled<-set_label(FF_labeled, get_label(FF))
FF_labels<-FF_labeled
```

## Cleaning pdqc data

```{r, pdqcReadin}
pdqc_all<-readRDS(paste0(datadir, 'OGData/', "pd_qc.rds"))
pdqc<-pdqc_all %>% filter(probe_fail_10==0 & sex==predicted_sex)
#What FF ids have methylation data?
methy_ids<-pdqc$id
#create slide variable fo pdqc 
pdqc$slide <- gsub('_.*', '', pdqc$MethID)
pdqc$array<-gsub('.*_', '', pdqc$MethID)
#ids with 9visit, 15visit or both visit
visit9ids<-pdqc%>%filter(childteen=='C')%>%pull(idnum)%>%unique()
visit15ids<-pdqc%>%filter(childteen=='T')%>%pull(idnum)%>%unique()
bothvisit_ids<-visit9ids[visit9ids %in% visit15ids]
visit9only_ids<-visit9ids[!(visit9ids %in%bothvisit_ids)]
visit15only_ids<-visit15ids[!(visit15ids %in%bothvisit_ids)]
```

```{r, createAnalyticSubsetFlags}
#create flag in FF for having methylation data
#and flag for having 2 visits of methylation data
FF_labeled<-FF_labeled %>% 
  mutate(MethylData=ifelse(id %in% methy_ids, "Analysis subset", "Not in analysis subset"),
         TwoVisitData=ifelse(id%in%bothvisit_ids,"Analysis subset - both visits", "Not in both visit analysis subset"))
#table(FF_labeled$MethylData)
#table(FF_labeled$TwoVisitData)
```

```{r, checkIDlengths}
if((length(visit15only_ids)+length(visit9only_ids)+length(bothvisit_ids)!=length(FF_labeled %>% filter(MethylData=="Analysis subset")%>%pull(id)))){print("WARNING ID LENGTHS DO NOT ADD UP")}
if(length(bothvisit_ids)!=length(FF_labeled %>% filter(TwoVisitData=="Analysis subset - both visits")%>%pull(id))){print("WARNING ID LENGTHS DO NOT ADD UP")}
```

### Filter technical replicates 
I filtered to only 1 technical replicate per person-visit, picking the replicate with the lower probe fail percentage

```{r}
#technical replicates will be duplicate id + childteen variables from pdqc
pdqc<-pdqc %>% mutate(newID=paste(idnum, childteen, sep='_'))
techrep_ids<- pdqc %>% select(newID) %>% group_by(newID)%>%filter(n()>1)
techreps<-pdqc %>% filter(newID %in% techrep_ids$newID)%>%arrange(idnum,childteen)
hiPctPF<-techreps %>% group_by(newID) %>%filter(probe_fail_pct==max(probe_fail_pct))%>%pull(MethID)
pdqc_clean<-pdqc %>% filter(!(MethID %in% hiPctPF) & childteen!='M')
myMethIDs<-pdqc_clean %>% pull(MethID)
```

### Plate, slide and array data

```{r read in  batch data}
batchData<-read.csv(file.path(datadir, 'Methylation_450K_array_batch_information.csv'))
```

# Creating new variables

## Creation of new sociodemographic and behavioral variables

### Prenatal maternal smoking exposure variable: binary smoking variable

```{r create binary smoking variable}
FF_labeled<-FF_labeled %>%mutate(smkPreg_binary=ifelse(m1g4%in% c("1 2+pk/d","2 1<pk<2","3 <1pk/d" ), "Yes", ifelse(m1g4%in%c("-2 Don't know", "-3 Missing", "-9 Not in wave"), NA, ifelse(m1g4== "4 None", "No", ifelse(m1g4=="Missing", "Missing", "FLAG")))))
createTable(compareGroups(m1g4~smkPreg_binary, data=FF_labeled))
```

### Ancestry variables

Ancestry is based on csv files of ancestry-specific PCS created by Dr. Erin Ware (per Jonah meeting Oct 6 2020). If an ID was in {african/european/hispanic}, then the individual is labeled as such.

```{r read in pc data and make ancestry variable}
# read in PCs
euroPC<-read.csv(paste0(datadir, "OGData/pcs/european.csv"))
afrPC<-read.csv(paste0(datadir, "OGData/pcs/african.csv"))
hisPC<-read.csv(paste0(datadir, "OGData/pcs/hispanic.csv"))
allPC<-read.table(paste0(datadir, "OGData/pcs/global_pcs.txt"), col.names=c("X", "idnum", paste("PC", 1:20, sep='')))
#individuals in each ancestry specific pc group can be labeled as that ancestry for categorical ancestry variable 
FF_labeled<-FF_labeled %>% mutate(ancestry=case_when(MethylData=="Analysis subset" & id %in% hisPC$idnum~"Hispanic ancestry", MethylData=="Analysis subset" & id %in%afrPC$idnum~"African ancestry", MethylData=="Analysis subset" & id %in% euroPC$idnum~"European ancestry", MethylData=="Analysis subset" & !(id %in%c(hisPC$idnum, afrPC$idnum, euroPC$idnum)) & (id %in% allPC$idnum)~"Other ancestry", MethylData=="Analysis subset" & !(id %in%c(hisPC$idnum, afrPC$idnum, euroPC$idnum)) & !(id %in% allPC$idnum)~"Missing PC data", MethylData!="Analysis subset"~"Not in analysis subset"))
#check ancestry variable
#table(FF_labeled$ancestry)
#with(FF_labeled, xtabs(~ancestry+cm1ethrace))
MissingPCdata<-FF_labeled %>% filter(ancestry=="Missing PC data")%>%select(id)
save(MissingPCdata, file = paste0(datadir, "CreatedData/MissingPC.Rdata"))
#add PC data to FF_labeled 
FF_labeled<-left_join(FF_labeled, allPC %>% mutate(id=as.character(idnum)) %>% select(id, PC1, PC2)%>%rename_at(vars(-id), function(x) paste0('global_', x, sep='')))
localPC<-rbind(euroPC, afrPC, hisPC)
FF_labeled<-left_join(FF_labeled, localPC  %>% mutate(id=as.character(idnum)) %>% select(id, PC1, PC2)%>% rename_at(vars(-id), function(x) paste0('local_', x, sep='')))
```

There were `r length(MissingPCdata$id)` individuals without PC data. None of these individuals had data in the global PC file, so I assume they are actually missing genetic data (for whatever reason, possibly QC).

### Other prenatal substance exposure variables

Recoded frequency of prenatal alcohol drinking and drug use into Yes/no binary variables for drinking and drug use during pregnancy. 
```{r PrenatalExposure}
#table(FF_labeled$m1g2) # during preg, alchohol freq
#table(FF_labeled$m1g3) # during preg, drugs freq 
#with(FF_labeled, table(m1g2, m1g3))
FF_labeled<-FF_labeled %>% mutate_at(c("m1g2", "m1g3"), .funs=funs(YesNoPreg=ifelse(.=="5 Never", "No", ifelse(. =="Missing", "Missing", "Yes"))))
with(FF_labeled, table(m1g2, m1g2_YesNoPreg))
with(FF_labeled, table(m1g3, m1g3_YesNoPreg))
#table(FF_labeled$m1g5) # in last year, alc/drugs interfered with work/relationships
#table(FF_labeled$m1g6) # ever sought help for drug/alc problems
```
### Postnatal smoke exposure

Maternal postnatal smoking at different survey points is fairly correlated, and alluvial plots of changes over time show broad consistency, so I decided to create a single smoking variable for maternal smoking at age 1 and age 5 - if mothers were smoking at either age 1 or age 5 then "Maternal smoking at age 1 or 5", if not smoking at age 1 and age 5 then "No maternal smoking at age 1 and age 5" and then if missing data at one age and no at the other, then "Missing" (and if missing at both ages, "Missing"). Also created a similar dose variable capturing maternal packs per day that could be used instead. Decided to use just age 1 and age 5 because at these two time points the mother was asked (as opposed to PCG, also the dataset of Fragile Families metadata is missing PCG questions from age 3 about smoking) and they both preceded the actual saliva collection for both saliva collection age points. Then use a smoking dose variable from the mother or PCG interview at wave of saliva collection as an additional second hand smoke exposure variable.

```{r PostnatalSmoke investigation}
#Baseline (1) -> Year 1/Age 1 (2) ->Year 3/Age3 (3) -> Year 5 (4) ->Year 9 (5) -> year 15 (6)
#Do you smoke in past month (how many packs per day)
#Mom            ->m2j5 (m2j5a)    ->                  -> m4j18 (m4j19) /p4 -> m5g17 (m5g18) / n5f17 (n5f18) 
#PCG                              p3a23a(p3a23b)                              n5f17 (n5f18)-> p6h76 (p6h77)
#seem to be missing p3 variables (PCG survey year 3) 
#number of people in household who smoke 
#p3a23->p5h15b->p6h78

#hours child around smoke 
#p3a21 -> p4a24->p5h15c

#with(FF_labeled, table(m2j5, m4j18))
#with(FF_labeled, table(m4j18, m5g17))
#with(FF_labeled, table(n5f17, m5g17))
#with(FF_labeled, table(p6h76, m5g17))

cor_plot_postnatalsmk<-plot_grid(FF_labeled %>% select(m2j5, m4j18, m5g17, n5f17, p6h74) %>% mutate_all(funs(as.numeric(.))) %>%ggcorr(label=T)+ggtitle("Correlation between Yes/No smoking (mother) \n at years 1, 5, & 9 & PCG at years 9 & 15")
,FF_labeled %>% select(m2j5a, m4j19, m5g18, n5f18, p6h77) %>% mutate_all(funs(as.numeric(.))) %>%ggcorr(label=T)+ggtitle("Correlation between packs per day smoking (mother) \n at years 1, 5, 9 & PCG at years 9 and 15"), nrow=2)

yesnosmkalluv<-to_lodes_form(as.data.frame(FF_labeled %>% select(m2j5, m4j18, m5g17, p6h74) %>% mutate(m5g17=stringr::str_to_title(m5g17))%>% group_by(m2j5, m4j18, m5g17, p6h74) %>% count()), axes=1:4, id='Cohort')

#ggplot(yesnosmkalluv, aes(x=x, stratum = stratum, alluvium=Cohort, fill=stratum, label=stratum, y=n))+geom_flow(stat="alluvium", lode.guidance="backfront", color='darkgrey')+geom_stratum()+scale_x_discrete(labels=c("Age 1", "Age 5", "Age 9"))

smkdose<-to_lodes_form(as.data.frame(FF_labeled %>% select(m2j5a, m4j19, m5g18, p6h77)%>% mutate_at(vars(c("m2j5a", "m4j19", "m5g18")), funs(factor(substring(., 1, 1), labels=levels(FF_labeled$m2j5a)))) %>% group_by(m2j5a, m4j19, m5g18, p6h77) %>% count()), axes=1:4, id='Cohort')
#ggplot(smkdose, aes(x=x, stratum = stratum, alluvium=Cohort, fill=stratum, label=stratum, y=n))+geom_flow()+geom_stratum()+scale_x_discrete(labels=c("Age 1", "Age 5", "Age 9"))

alluv_plot_postnatalsmk<-plot_grid(ggplot(yesnosmkalluv, aes(x=x, stratum = stratum, alluvium=Cohort, fill=stratum, label=stratum, y=n))+geom_flow()+geom_stratum()+scale_x_discrete(labels=c("Age 1", "Age 5", "Age 9", "Age 15"))
+ggtitle("Alluvial plot of Yes/No smoking variables from age 1 to age 9 (mother) and age 15 (PCG)"), ggplot(smkdose %>% filter(!(stratum %in% c("-6 Skip", "Missing"))), aes(x=x, stratum = stratum, alluvium=Cohort, fill=stratum, label=stratum, y=n))+geom_flow()+geom_stratum()+scale_x_discrete(labels=c("Age 1", "Age 5", "Age 9", "Age 15"))+ggtitle("Alluvial plot of smoking per day among those smoking from age 1 to age 9 (mother) and 15 (PCG)"), nrow=2)

#ggplot(FF_labeled, aes(x=p5h15b, y=p6h78))+geom_jitter()+stat_cor()+scale_x_continuous('# smoking in house age 9')+scale_y_continuous('# smoking in house age 15')+ggtitle("# individuals smoking in house correlation age 9 and age 15")
```

```{r plots for postnatal smoking}
cor_plot_postnatalsmk
alluv_plot_postnatalsmk
```

```{r make maternal postantal smoking vars for 1 and 5}
#Make any maternal postnatal smoking variable & dose categorization for maternal smoking at ages 1 & 5 
FF_labeled<-FF_labeled %>% mutate(PostnatalMaternalSmokingAny = case_when(
                                m2j5 =="1 Yes" ~ "Maternal smoking at age 1 or age 5",
                                m4j18=="1 Yes" ~ "Maternal smoking at age 1 or age 5",
                                m2j5 == "Missing" & m4j18=="Missing" ~ "Missing",
                                m2j5 == "Missing" & m4j18=="2 No" ~ "Missing",
                                m2j5 == "2 No" & m4j18=="Missing" ~ "Missing",
                                m2j5 == "2 No" & m4j18=="2 No" ~ "No maternal smoking at age 1 and age 5")) %>%
  mutate(PostnatalMaternalSmokingDose=case_when(
                                m2j5a %in% c("2 1pk/d", "3 1.5pk/d", "4 2pk/d", "5 >2pk/d") &
                                m4j19 %in% c("2 About pack/day", "3 Pack and half/day", "4 2 packs/day", "5   
                                          More 2 packs/day") ~ "Mom - pack/d or greater at both age 1 & 5", 
                                m2j5a %in% c("2 1pk/d", "3 1.5pk/d", "4 2pk/d", "5 >2pk/d") &
                                !(m4j19 %in% c("2 About pack/day", "3 Pack and half/day", "4 2 packs/day", "5 
                                          More 2 packs/day")) ~ "Mom - pack/d or greater at either age 1 or 5",
                                !(m2j5a %in% c("2 1pk/d", "3 1.5pk/d", "4 2pk/d", "5 >2pk/d")) &
                                m4j19 %in% c("2 About pack/day", "3 Pack and half/day", "4 2 packs/day", "5 
                                          More 2 packs/day") ~ "Mom - pack/d or greater at either age 1 or 5",
                                m2j5a %in% c("1 <1/2 pk/d") | m4j19 %in% c("1 Less half pack/day")~"Mom - less than half pack/d at either age 1 or 5", 
                                m2j5a %in% c("-6 Skip") & m4j19 %in% c("-6 Skip")~
                                  "No maternal smoking at age 1 or 5", 
                                m2j5a %in% c("Missing") & m4j19 %in% c("-6 Skip", "Missing")~"Missing", 
                                m2j5a %in% c("Missing", "-6 Skip") & m4j19 %in% c("Missing")~"Missing"))

with(FF_labeled, table(PostnatalMaternalSmokingAny, m2j5, m4j18))
#with(FF_labeled, table(m2j5a, m4j19, PostnatalMaternalSmokingDose))
#with(FF_labeled, table(PostnatalMaternalSmokingAny, PostnatalMaternalSmokingDose))
```

## Import and create methylation summary variables

### Cell type proportions

Cell type proportions currently estimated using [epidDISH](https://www.bioconductor.org/packages/release/bioc/vignettes/EpiDISH/inst/doc/EpiDISH.html#how-to-estimte-cell-type-fractions-using-dnam-data) package and the EpiFibIC reference (generic for epithelial tissues). This will be replaced with the saliva specific reference panel in [Middleton et al.](https://www.medrxiv.org/content/medrxiv/early/2020/09/15/2020.09.14.20191361.full.pdf).

```{r read in betaqc & celltype proportions}
betaqc<-readRDS(file=paste0(datadir, 'OGData/', "noob_filtered.rds"))

#cell type proportions 
celltypes<-epidish(beta.m = betaqc, ref.m = centEpiFibIC.m, method = "RPC")
estF_FF<-as.data.frame(celltypes$estF)
estF_FF$MethID = row.names(estF_FF)
```

In addition to the site-specific analysis proposed, Kelly and I decided to investigate certain DNA methylation summary metrics that are commonly used in the field. These include global methylation and methylation clocks (which are outcome-naive), and an ooutcome-predicated measure, polymethylation scores.

### Global methylation

We calculated global methylation as the mean of the CpG site-specific methylation values.

```{r global methylation and cpg sites of interest}
aprioriCG<-c("cg05575921", "cg04180046", "cg05549655", "cg14179389")
globalmethy<-colMeans(betaqc, na.rm=T)
globalmethdf<-as.data.frame(cbind("MethID"=names(globalmethy), "globalmethylation"=globalmethy))
aprioriCGdf<-as.data.frame(t(betaqc[aprioriCG, ]))
aprioriCGdf$MethID = colnames(betaqc)
```

### Methylation (or epigenetic) clocks

Methylation or epigenetic clocks are measures which exploit certain sets of CpG sites whose DNA methylation levels associate with subject age. These clocks are highly accurate molecular correlates of chronological age. However, variation in epigenetic clock measures among individuals of the same chronological age do exist and are assumed to reflect biological age.

Four different methylation clocks were created, based on different papers that used different tissue types and age groups for their training samples: [Horvath13](https://genomebiology.biomedcentral.com/articles/10.1186/gb-2013-14-10-r115), [Horvath18](https://www.nature.com/articles/s41576-018-0004-3), [Pediatric](https://www.pnas.org/content/117/38/23329) and [Levine](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5940111/). Since this is a pediatric population, the pediatric clock is probably the best bet, but Jonah also suggested we look at the Levine clock.

```{r readInclocks}
clocks<-read.csv(file=paste0(datadir, "OGData/ffcw_n1776_8clocks.csv"), header = T)
colnames(clocks)[1]<-"MethID"
```

### Polymethylation score

Here, the idea is similar to a polygenetic score. Essentially, we want to create a summary metric of CpG sites that have associated with an exposure or outcome of interest in a previous independent cohort. We identify a previous study which conducted a regression analysis of our exposure of interest on CpG sites and extract those regression coefficients. The regression coefficients become weights for the CpG site values in our cohort. We sum the weighted CpG site values across sites for each sample to come up with a single per-sample measure of exposure-related methylation.

$Score_i=\sum^{m}_j Weight_j*CpG_{ij}$ where

$i$ indexes samples in our study; $j$ indexes CpG sites

$Weight_j$ is the regression coefficient from the independent external study for that $j$ CpG site

$CpG_{ij}$ is the CpG methylation value for that $j$ CpG in that $i$ sample in our study

$Score_i$ is the polymethylation score.

Because polymethylation scores are not widely used (yet), there isn't a standardized approach. Importantly, much of the time researchers use the (confusingly named) 'beta' value to measure methylation at a CpG site, which is the ratio of the methylated probe intensity to the overall intensity and ranges from 0-1. We were concerned that this could lead to sites with more methylation (closer to 1) being upweighted in the resulting score), so we implemented versions of the score where the $CpG_{ij}$ values were first mean-centered or z-score standardized before being used in the score.

For this analysis, we used a 2016 genome-wide consortium meta-analysisof prenatal maternal smoking and DNA methylation in blood samples, which used the same DNA methylation array. This study actually provided regression coefficients for all CpGs statistically significant after FDR correction ($m_{CpG}=6073$) in 4 different regressions:

1. Regression with sustained prenatal maternal smoking as the exposure, methylation in newborn cord blood samples as outcomes
2. Regression 1 adjusted for cell type confounding
3. Regression of sustained prenatal maternal smoking as the exposure, methylation in peripheral blood samples of older children as outcomes
4. Regression of any prenatal maternal smoking es the exposure, methylation in newborn cordblood as outcomes

These regression used normalized betas as the outcomes but found very little variation in results when using raw betas as the outcome (Spearman $\rho$=0.96 for regression coefficients, $\rho$=0.98 for $log_{10}(pvalues$))

```{r make polymethylaton score}
#read in cpgs from PMC4833289
cpgs<-read.csv(paste0(datadir, 'PMC4833289_SuppFile2CpGs.csv'), stringsAsFactors = F, header=F, skip=2)
#make headers
header1<-scan(paste0(datadir, 'PMC4833289_SuppFile2CpGs.csv'), nlines=1, what=character(), sep=',')
header1[7:26]<-c(rep("SSnewborn", 5), rep("SSnewbornCT", 5), rep("SSolder", 5), rep("anynewborn", 5))
header2<-paste(header1, scan(paste0(datadir, 'PMC4833289_SuppFile2CpGs.csv'), nlines=1, skip=1, what=character(), sep=','), sep='_')
names(cpgs)<-header2

#four meta-EWASes: SS_newborn (sustained smoking in newborns); SS_newbornCT (sustained smoking in newborns controlling for cell type) SS_older (sustained smoking in older children); any_newborn (any smoking in newborns)
#pivot_longer to get single column of coefs and a column for the analysis type
#and group into a list by analysis type
cpgs<-cpgs %>% pivot_longer(cols = c(matches("_Coef|_SE|_P|_FDR|_Direction")), names_to = c("set", ".value"), names_pattern="(.+)_(.+)")%>%rename_with(~gsub('_', '', .x))%>%group_split(set)
cpgs<-cpgs%>%setNames(lapply(cpgs, function(x){x$set %>% unique()}))

#make scores - no transform, zscore and centermean
#first, pull script 
source(file.path(codedir, 'UsefulCode', 'makePolyEpiScores.R'))
scores=makePolyEpiScore(m=betaqc, b=cpgs)
scores_z=makePolyEpiScore(m=betaqc, b=cpgs, transformopt = 'zscore')
scores_center=makePolyEpiScore(m=betaqc, b=cpgs, transformopt = 'meancenter')
polyscores=list('notransform'=scores, 'zscore'=scores_z, 'center'=scores_center)
polyscores<-lapply(polyscores, function(x) x%>% mutate(MethID=colnames(betaqc)))
save(polyscores, file = file.path(datadir, 'CreatedData', 'polymethylationscores.RDS'))
polyscores_wide<-lapply(seq_along(polyscores), function(i) polyscores[[i]] %>% setNames(c(paste0(colnames(polyscores[[i]])[1:4], '_', names(polyscores)[i]), 'MethID')))%>%reduce(left_join, by='MethID')
```

### Surrogate variables

```{r sva}

var.filter=5000
#make a matrix of model data 
#filter to only samples with methylation data (& kids??) 
pheno=FF_labeled%>%filter(MethylData=="Analysis subset")%>%left_join(pdqc%>%mutate(id=idnum)%>%select(MethID, id))
#check that dim of pheno and betaqc match

#want to make sure that the order of pheno and betaqc match
pheno=pheno[match(colnames(betaqc), pheno$MethID), ]
if(identical(colnames(betaqc), pheno$MethID)==FALSE){print('STOP: Sample names do not match')}

#make model matrix including variable of interest as factor variable & adjustment variables 
mod=model.matrix(~as.factor(smkPreg_binary), data=pheno)

#make null moddel containing only adjustment variables
mod0=model.matrix(~1, data=pheno)

#next, estimate the surrogate variables 
svobj=sva(betaqc, mod, mod0, vfilter=var.filter)
sv<-data.frame(svobj$sv); colnames(sv)<-paste('sv', 1:ncol(sv), sep='')
pheno=cbind(pheno, sv)

#bring the heatmap function
bring.the.heatmap.sv <- function(pd,covar,folder){
  library(gplots)

  nsv <- length(grep('sv',names(pd)))

  #generate correlation matrix between SVs and variables
  correlations <- matrix(0,nrow=length(covar),ncol=nsv)
  colnames(correlations) <- paste('sv',1:nsv,sep='')
  rownames(correlations) <- covar

  #take svs
  svs <- pd[,colnames(correlations)]

  #correlations between svs and variables:
  # cor test if continuous
  # anova is multi category
  # t-test if two category
  for(col in covar){
    var <- pd[,col]
    #check to see if it is numeric and isn't really a two category thing
    if(class(var)=='numeric' & dim(table(var))>2){
      correlations[col,] <- apply(svs,2,FUN = function(x,y) cor.test(x,y)$p.value,y=var)
    }else{
      if(length(levels(factor(var)))>2){
        correlations[col,] <- apply(svs,2,FUN = function(x,y) anova(lm(x~y))$'Pr(>F)'[1],y=var)
      }else{
        correlations[col,] <- apply(svs,2,FUN = function(x,y) t.test(x~y)$p.value,y=var)
      }
    }
  }

  pdf(paste0(folder,'heatmap-sv.pdf'))
  color.gradient <- colorRampPalette(c('firebrick','ivory3','royalblue'),bias=3)
  heatmap.2(correlations,Rowv=FALSE,Colv=FALSE,cexRow=1,cexCol=1,dendrogram='none',density.info='none',trace='none',
            col=color.gradient(20), key.xlab='p-value', main='Associations Between SVs\n and Given Covariates',
            lmat=rbind(c(0,3,3,3),c(2,0,4,0),c(0,1,1,1)), lhei=c(0.5,1,3.5), lwid=c(0.12,0.145,0.4,0.335), margin=c(4,7))
  dev.off()

  return('check your plot')
}

pheno=left_join(pheno, batchData)
bring.the.heatmap.sv(pheno, covar=c('Sample_Plate', 'Slide', 'Array'))

```

```{r make methyl dataframe}
clocks2<-clocks %>% mutate(idnum=gsub('a', '', idnum))%>%filter(MethID %in% pdqc$MethID)%>%select(MethID, idnum, childteen, horvath, pediatric, levine, grim) # currently clocks are missing 3 individuals who were in johns pipeline but not jonahs

methyldata<-left_join(left_join(left_join(left_join(left_join(left_join(pdqc_clean %>% select(MethID, slide, array,childteen, idnum), globalmethdf), clocks2), estF_FF), aprioriCGdf), polyscores_wide), batchData)
methyldata[, c("globalmethylation", aprioriCG)]<-lapply(methyldata[, c("globalmethylation", aprioriCG)],function(x) as.numeric(as.character(x)))
#lapply(methyldata, class)
methyldata_wide <- methyldata %>% select(-MethID)%>% pivot_wider(., names_from=childteen, values_from=any_of(colnames(methyldata)[4:length(colnames(methyldata))]))
```

```{r individual who is missing global methylation data but had apriori CpG and PMS data?, eval=F}
#weirdID<-methyldata %>% filter(is.na(globalmethylation))%>%pull(MethID)
#globalmethy[weirdID] #not actually missing global methylation data
#clocks %>% filter(MethID%in%weirdID) # #idnum from  not the same as in
#pdqc_clean %>% filter(MethID==weirdID) #this is just a typo in the idnum from data entry, see format of other #idnums
#pdqc_clean %>% pull(idnum) %>% head()
#solution: when creating methyldata, first trim alphabet from clocks$idnum that contain alphabetic characters --> implemented
#also misssing clocks for 3 sex outlier samples that weren't in jonahs' pipeline: 
table(is.na(methyldata$horvath13))
weirdIDs<-methyldata %>% filter(is.na(horvath13))%>%pull(MethID)
#pdqc_all %>% filter(MethID %in% weirdID) %>% xtabs(~predicted_sex_outlier, data=.)
```
## Distributions & checks of methylation variables

### Add methylation data to metadata

```{r makeMyFF}
varLabels2<-c(paste(colnames(FF_labeled)[1:66], varLabels), "Has any methylation data?", "Has two visits of methylation data", "Any maternal smoking during pregnancy", "Ancestry from PCA of child's genetic data", "PC 1 (global)", "PC 2 (global)", "PC 1 (within ancestry category)", "PC 2 (within ancestry category)",
"Any alcohol during pregnancy", "Any drugs during pregnancy", "Any postnatal maternal smoking (age 1 & 5)", "Postnatal maternal smoking dose at age 1 & 5")
FF_labeled<-set_label(FF_labeled, varLabels2)

myFF<-FF_labeled %>% filter(MethylData=="Analysis subset")%>%mutate(idnum=id)
myFF<-left_join(myFF, methyldata)
varLabels3<-c(varLabels2, "ID", "Methylation ID", "Slide", "Array (RC)", "Visit", "Global methylation", "horvath methylation clock", "pediatric methylation clock", "levine methylation clock", "GRIM methylation clock", "Epithelial cell proportion", "Fibroid cell proportion", "Immune cell proportion", "cg05575291 methylation", "cg04180046 methylation",  "cg05549655 methylation", "cg14179389 methylation", "PES: Any smoking, newborn cordblood", "PES: Sustained smoking, newborn cordblood", "PES: Sustained smoking, newborn cordblood (+CT control)", "PES: Sustained smoking, older children peripheral blood", "PES: Any smoking, newborn cordblood - zscore", "PES: Sustained smoking, newborn cordblood - zscore", "PES: Sustained smoking, newborn cordblood (+CT control) - zscore", "PES: Sustained smoking, older children peripheral blood - zscore", "PES: Any smoking, newborn cordblood - mean center", "PES: Sustained smoking, newborn cordblood - mean center", "PES: Sustained smoking, newborn cordblood (+CT control) - mean center", "PES: Sustained smoking, older children peripheral blood - mean center", 'Plate', 'Slide', 'Array')
myFF<-set_label(myFF, varLabels3)
```

## Correlation and distribution of global methylation and epigenetic clocks
```{r correlation global and clocks}
chart.Correlation(myFF %>%select(globalmethylation, levine, pediatric, grim))
```

### Correlation and distribution of polymethylation scores
```{r correlation polymethylation scores}
chart.Correlation(myFF %>%select(SSnewborn_notransform, SSnewbornCT_notransform, SSolder_notransform, anynewborn_notransform))
```

```{r save correlation chart}
postscript(file.path(outputdir, 'CorrelationPMSWeights.ps'))
chart.Correlation(myFF %>%select(SSnewborn_notransform, SSnewbornCT_notransform, SSolder_notransform, anynewborn_notransform))
dev.off()
```
### Correlation and distribution of polymethylation score coefficients

```{r correlation score coefficients}
coefs_methyl<-cbind('SSolderCoefs'=cpgs$SSolder$Coef, 'SSnewbornCTCoefs'=cpgs$SSnewbornCT$Coef, 'SSnewbornCoefs'=cpgs$SSnewborn$Coef, 'anynewbornCoefs'=cpgs$anynewborn$Coef)
chart.Correlation(coefs_methyl)
```

### Distributions of poly methylation scores with different transformations on beta methylation matrix

```{r compare PolyMethylation scores distributions, fig.height=11}
#lapply(polyscores, function(x){summary(x)})
#lapply(polyscores, function(x){apply(x, 2, sd)})
dist_plots=lapply(seq_along(polyscores), function(i){polyscores[[i]]%>%pivot_longer(cols=1:4, names_to='Score_Type', values_to='Score')%>%ggplot(aes(x=Score))+geom_histogram()+facet_wrap(~Score_Type)+ggtitle(names(polyscores)[[i]])+theme_classic()})
plot_grid(dist_plots[[1]], dist_plots[[2]], dist_plots[[3]], ncol=1)
```

```{r only newborn celltype controlled comparison of transform methods }
transform_labels<-c('Mean center', 'No transformation', 'Z-score standardization')
names(transform_labels)<-c('center', 'notransform', 'zscore')
postscript(file.path(outputdir, 'DistributionPMStransformation.ps'))
polyscores %>% bind_rows(.id='transform')%>%ggplot(aes(x=SSnewbornCT))+geom_histogram()+facet_wrap(~transform, scales='free', labeller=labeller(transform=transform_labels))+theme_classic()+theme(text=element_text(size=16))
dev.off()

```

As Kelly thought, the zscore standardized have a larger range and the biggest standard deviations, the standard deviations of the no transform and centered methods are the same, which makes sense as $V(aX+b)=a^2V(X)\implies V(\sum\beta_i*[x_i-\bar{x}])=V(\sum\beta_i*x_i-\beta_i*\bar{x})=V(\sum \beta_i*x_i-k)=V(\sum \beta_i*x_i)$.

### Correlation between methylation data
```{r}
methylation_data<-myFF %>% select(globalmethylation, pediatric, levine, Epi, Fib, IC, anynewborn_center, SSnewbornCT_center, SSolder_center)
chart.Correlation(methylation_data, histogram=T)
```

### Methylation variables over time or by visit

```{r celltype over time}
#Cell type
cellbytime<-myFF %>% pivot_longer(cols=c(Fib, Epi, IC), names_to="Celltype", values_to = "Proportion")

celltype_labels<-c('Epithelial cells', 'Fibroblasts', 'Immune cells'); names(celltype_labels)<-c('Epi', 'Fib', 'IC')

ggplot(cellbytime, aes(x=childteen, y=Proportion, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+stat_compare_means(label="..p.signif..")+ggtitle("Cell type proportions by visit")+theme_classic()+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()
```
```{r celltypes by visit spaghetti}
cairo_ps(file.path(outputdir, 'SpaghettiCellTypes.ps'))
ggplot(cellbytime, aes(x=childteen, y=Proportion, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~Celltype, labeller=labeller(Celltype=celltype_labels))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+ylab('Proportion')+theme(text=element_text(size=16))
dev.off()
```

```{r clocks and global over time, fig.width=9}
global<-ggplot(myFF, aes(x=childteen, y=globalmethylation, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Global methylation")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
pediatric<-ggplot(myFF, aes(x=childteen, y=pediatric, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Pediatric methylation clock")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
levine<-ggplot(myFF, aes(x=childteen, y=levine, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Levine methylation clock")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
plot_grid(global, pediatric, levine, nrow=1)
```

```{r scores over time, fig.width=9}
#ggplot(myFF, aes(x=childteen, y=cg05575921, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Methylation at cg05575921 by visit")
any<-ggplot(myFF, aes(x=childteen, y=anynewborn_center, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Any smoking, \n newborn polymethylation score")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
SSnewborn<-ggplot(myFF, aes(x=childteen, y=SSnewbornCT_center, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Sustained smoking, \n newborn CT polymethylation score ")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
SSolder<-ggplot(myFF, aes(x=childteen, y=SSolder_center, color=childteen))+geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+ggtitle("Sustained smoking, \n older children polymethylation score")+scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+theme_classic()+theme(legend.position = 'bottom')
plot_grid(any, SSnewborn,SSolder, ncol=3)

```

Cell type proportions do differ between 9 and 15 year visits and although global methylation does not look different between these visits, methylation clocks do (a good check).

Possibly higher polymethylation scores at teen visit, although inconsistent across different score types.

### Correlation betwen polymethylation scores over time
```{r correlaton b/t polymethylation scores over time, fig.width=9}
PMS<-methyldata %>%select(idnum, childteen, colnames(polyscores_wide), -MethID)%>%pivot_longer(cols=any_of(colnames(polyscores_wide)), names_to="PMS_type", values_to="PolymethylationScore")%>%pivot_wider(names_from = childteen, values_from=PolymethylationScore)
PMSType_label<-c('Any smoking;\n newborn', 'Sustained smoking;\n newborn', 'Sustained smoking + cell type;\n newborn', 'Sustained smoking;\n older')
names(PMSType_label)<-c('anynewborn_center', 'SSnewborn_center', 'SSnewbornCT_center', 'SSolder_center')
ggplot(PMS%>% filter(str_detect(PMS_type, '_center')), aes(x=C, y=T))+geom_point()+facet_grid(~PMS_type, labeller=labeller(PMS_type=PMSType_label))+stat_cor()+geom_smooth(method=lm)+theme_classic()+scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')
```

```{r save correlation over time image}
postscript(file.path(outputdir, 'ScoreCorrelationVisits.ps'))
ggplot(PMS%>% filter(str_detect(PMS_type, '_center')), aes(x=C, y=T))+geom_point()+facet_grid(~PMS_type, labeller=labeller(PMS_type=PMSType_label))+stat_cor()+geom_smooth(method=lm)+theme_classic()+scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+theme(text=element_text(size=16))
dev.off()
```

### Spaghetti plots of polymethylation scores over time
```{r Polymethylation scores over time, fig.width=9}
methyldata %>% pivot_longer(cols=contains('_center'), names_to='measure', values_to='Score')%>%ggplot(aes(x=childteen, y=Score, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~measure, labeller=labeller(measure=PMSType_label))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+ylab('Score')
```

```{r save spaghetti poly}
cairo_ps(file.path(outputdir, 'SpaghettiScores.ps'))
methyldata %>% pivot_longer(cols=contains('_center'), names_to='measure', values_to='Score')%>%ggplot(aes(x=childteen, y=Score, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~measure, labeller=labeller(measure=PMSType_label))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+ylab('Score')+theme(text=element_text(size=16))
dev.off()
```

## Make visit specific variables

### Child's age at time of interview/visit
```{r}
myFF %>% filter(is.na(cp6yagem) &is.na(hv6_yagem) & childteen=='T')%>%dim() #3
myFF %>% filter(is.na(cm5b_age) &is.na(hv5_agem) &childteen=='C')%>%dim() #1 
myFF %>% filter(!is.na(cm5b_age) & is.na(hv5_agem) & childteen=='C')%>%dim()#1
myFF %>% filter(!is.na(cp6yagem) &is.na(hv6_yagem))%>%dim() #990

myFF <- myFF %>% 
  mutate(ChildAgeAtIn=case_when(childteen=='C'~ cm5b_age, childteen=='T'~cp6yagem))%>%
  mutate(ChildAgeAtVisit=case_when(childteen=='C'~hv5_agem, childteen=='T'~hv6_yagem))%>% 
  mutate(ChildAgeComposite=case_when(childteen=='C' & !is.na(hv5_agem)~hv5_agem, 
         childteen=='C' & is.na(hv5_agem)~cm5b_age, 
         childteen=='T' & !is.na(hv6_yagem)~hv6_yagem, 
         childteen=='T' & is.na(hv6_yagem)~cp6yagem))

myFF %>% filter(childteen=='C')%>%group_by(ChildAgeAtVisit, ChildAgeComposite)%>%summarise(n=n())%>%filter(ChildAgeAtVisit!=ChildAgeComposite)

myFF %>% filter(childteen=='T')%>%group_by(ChildAgeAtVisit, ChildAgeComposite)%>%summarise(n=n())%>%filter(ChildAgeAtVisit!=ChildAgeComposite)

ggplot(myFF%>%filter(childteen=='C'), aes(cm5b_age, hv5_agem))+geom_point()+stat_cor()+ggtitle("Correlation between childs age at mom interview (9 yr) and childs age at home visit (9 yr)")
age5resid<-myFF %>% filter(childteen=='C')%>%mutate(age5_resid=cm5b_age-hv5_agem)%>%pull(age5_resid)
hist(age5resid)

age6resid<-myFF %>% filter(childteen=='T')%>%mutate(age6_resid=cp6yagem-hv6_yagem)%>%pull(age6_resid)
ggplot(myFF%>%filter(childteen=='T'), aes(cp6yagem,hv6_yagem))+geom_point()+stat_cor()+ggtitle("Correlation between childs age at mom interview (15 yr) and childs age at home visit (15 yr)")
hist(age6resid)

table(FF_labeled$cp6yagem-FF_labeled$hv6_yagem)
table(FF_labeled$cm5b_age-FF_labeled$hv5_agem)
table(is.na(myFF$ChildAgeComposite)) # 4

```

### Maternal or PCG smoking dose at time of sample

```{r visitMaternalSmoke}
myFF<-myFF %>% mutate(SmkAtVisitPastmonth=case_when(childteen=='C' & m5g17=='1 yes' & m5g18 %in% c("2 about a pack", "3 a pack and a half", "4 about 2 packs", 
                              "5 more than two  packs") ~ "Pack or more a day", 
                              childteen=='C' & m5g17=='1 yes' & m5g18=="1 less than half a pack a day" ~ "Less than pack a day", 
                              childteen=='C' & n5f17=="1 yes" & n5f18 =='2 about a pack' ~ "Pack or more a day",
                              childteen=='C' & n5f17=="1 yes" & n5f18 =="1 less than half a pack day" ~ "Less than pack a day",
                              childteen=='C' & m5g17=="1 yes" & m5g18 %in% c("-6 Skip", "Missing")~"Missing",
                              childteen == 'C' & n5f17%in%c("-7 N/A", "2 no") & m5g17=="2 no" ~ "No smoking", 
                              childteen == 'C' & n5f17%in%c("2 no") & m5g17%in%c("2 no", "Missing") ~ "No smoking", 
                              childteen == 'C' & n5f17 %in% c("Missing", "-7 N/A") & m5g17 =="Missing" ~ "Missing", 
                              childteen=='T' & p6h76 %in% c("-6 Skip", "0 None") ~ "No smoking", 
                              childteen=='T' & p6h77%in% c("1 5 or fewer cigarettes per day", "2 About half a pack a day, 10 cigarettes")~"Less than pack a day", 
                              childteen=='T' & p6h77 %in% c("3 About a pack a day, 20 cigarettes", "4 More than 1 pack a day")~"Pack or more a day",
                              childteen=='T' & p6h77 == "Missing"~"Missing"))%>%
              mutate(SmkAtVisitYesNo=case_when(SmkAtVisitPastmonth=="No smoking"~"No", 
                                               SmkAtVisitPastmonth %in% c("Less than pack a day", "Pack or more a day") ~ "Yes", 
                                               childteen=='T'& SmkAtVisitPastmonth=="Missing" & !(p6h76 %in% c("-6 Skip", "0 None", "Missing"))~"Yes",
                                               childteen=='C' & SmkAtVisitPastmonth=="Missing" & m5g17=='1 yes'~"Yes", 
                                               childteen=='C' & SmkAtVisitPastmonth=="Missing" & m5g17!='1 yes'~"Missing", 
                                               childteen=='T'& SmkAtVisitPastmonth=="Missing" & p6h76 %in% c("-6 Skip", "0 None", "Missing")~"Missing"))

myFF %>% filter(childteen=='C') %>% xtabs(~SmkAtVisitPastmonth+SmkAtVisitYesNo, data=.)
myFF %>% filter(childteen=='T') %>% xtabs(~SmkAtVisitPastmonth+SmkAtVisitYesNo, data=.)
myFF$SmkAtVisitPastmonth<-factor(myFF$SmkAtVisitPastmonth)
```
### Child smoking
Exclude children who reported smoking at age 9 and age 15. Interestingly, the children who report smoking at age 9 do not report smoking at age 15, and the parental perception of child smoking at age 9 does not match the children who say they have ever smoked a cigarette. For now using the 'ever smoked' variable from age 15 for the exclusion, but there is also a 'smoked in past month' that we could use instead (lose fewer children).

```{r, ChildSmoke}
with(myFF, table(k5f1l,childteen)) # 6 kids smoking at age 9 
with(myFF, table(p5q3cr, k5f1l)) # not the same kids whose parents say somewhat or very true that the child is using tobacco
with(myFF, table(k6d40, childteen)) # 41 kids say ever smoked a cigarette at age 15
#with(myFF, summary(k6d41))
with(myFF, table(k6d42, childteen)) # of these only 13 say they've smoked in the past month 
#with(myFF, summary(k6d43))
with(myFF, table(k5f1l, k6d40, childteen))# not the same kids at age 9 and 15
```

```{r}
varLabels4<-c(varLabels3, "Child's age at maternal/PCG (9/15 yr) interview (months)", "Child's age at home vist (9/15 years) (months)", "Constructed child's age at home visit (in months) (9/15 years)", "Mother/PCG (9/15 yr) smoking in month prior to interview (pk/day)", "Mother/PCG (9/15 yr) smoking in month prior to interview (yes/no)")
myFF<-set_label(myFF, varLabels4)
```

# Analysis

## Randomization of prenatal maternal smoking across slide numbers

```{r slide number vs prenatal maternal smoking, fig.width=8}
slide_count<-myFF %>% mutate(slide=as.factor(slide))%>% group_by(smkPreg_binary, slide, .drop=F)%>%count()%>%pivot_wider(id_cols = slide, names_from=smkPreg_binary, values_from=n)%>%arrange(Yes)%>%mutate(Proportion=Yes/(No+Yes))
hist_data<-slide_count %>% ggplot(aes(x=Proportion))+geom_histogram(binwidth = 0.1, fill='blue', alpha=0.5, breaks=seq(0, 1, 0.1))+ggtitle('Histogram proprotion smk (from data)')
hist_data_c<-slide_count %>% ggplot(aes(x=Yes))+geom_histogram( fill='blue', alpha=0.5, breaks=c(0:12))+scale_x_continuous(breaks=c(0:12))+ggtitle('Histogram # Yes smk (from data)')

kable(slide_count)%>%kable_styling()%>%scroll_box(width='100%', height='400px')

#simulation
set.seed(seed=20210216)
n<-nrow(myFF)
yes<-table(myFF$smkPreg_binary)[2]; no<-table(myFF$smkPreg_binary)[1]
values<-data.frame('smk'=sample(c('No', 'Yes'), n, replace=T, prob=c(no/(no+yes), yes/(no+yes))))
values_slides<-values %>% 
    group_by((row_number()-1) %/% (n()/152)) %>%
    nest()%>%unnest(cols=c(data))
colnames(values_slides)<-c('slide', 'smk')
slide_sim<-values_slides%>%mutate(slide=as.factor(slide)) %>% group_by(smk, slide, .drop=F)%>%count()%>%pivot_wider(id_cols = slide, names_from=smk, values_from=n)%>%arrange(Yes)%>%mutate(Proportion=Yes/(No+Yes))
hist_sim<-slide_sim%>% ggplot(aes(x=Proportion))+geom_histogram(binwidth = 0.1, fill='red', alpha=0.5, breaks=seq(0, 1, 0.1))+ggtitle('Histogram proprotion smk (from simulation)')
hist_sim_c<-slide_sim%>% ggplot(aes(x=Yes))+geom_histogram(fill='red', alpha=0.5, breaks=c(0:12))+scale_x_continuous(breaks=c(0:12))+ggtitle('Histogram # Yes smk (from simulation)')

hist_overlay<-slide_count %>% ggplot(aes(x=Proportion))+geom_histogram(fill='blue', binwidth = 0.1, alpha=0.5, breaks=seq(0, 1, 0.1))+geom_histogram(aes(x=slide_sim$Proportion), breaks=seq(0, 1, 0.1), fill='red', binwidth = 0.1, alpha=0.5)+xlim(0,1)
hist_overlay_c<-slide_count %>% ggplot(aes(x=Yes))+geom_histogram(fill='blue', alpha=0.5, breaks=c(0:12))+geom_histogram(aes(x=slide_sim$Yes), fill='red',  alpha=0.5, breaks=c(0:12))+scale_x_continuous(breaks=c(0:12))

#plot_grid(plot_grid(hist_data, hist_sim, nrow=2), hist_overlay, nrow=1)
plot_grid(plot_grid(hist_data_c, hist_sim_c, nrow=2), hist_overlay_c, nrow=1)

```

## Association of prenatal maternal smoking with plate

This was added 05/05/21 following 04/17/21 email from Jonah with batch variables from Princeton (thanks Jonah!)
```{r PreMSmk & plate (ie batch)}
Platechisq<-chisq.test(table(myFF$Sample_Plate, myFF$smkPreg_binary))
kable(table(myFF$Sample_Plate, myFF$smkPreg_binary))
```

Prenatal maternal smoking is not significantly associated with plate/batch ($\chi^2=$`r round(Platechisq$statistic, 2)`; P-value=`r round(Platechisq$p.value, 2)` )

## Flow chart of sample exclusions

```{r MakeFlowChart}
methylCohort<-pdqc_all %>% filter(childteen!='M')

basemodelvar<-c("smkPreg_binary", "ancestry", "cm1bsex", "cm1inpov", "Epi", "IC", "ChildAgeComposite")
basemodeldata<-myFF %>% filter_at(all_of(basemodelvar), all_vars(!(. %in%c("Missing", "Missing PC data"))& !is.na(.)))
child_base<-basemodeldata%>%filter(childteen=='C')
teen_base<-basemodeldata%>%filter(childteen=='T')

prenatalexposurevar<-c(basemodelvar, "m1g2_YesNoPreg", "m1g3_YesNoPreg")
prenatalexdata<-basemodeldata %>%filter_at(all_of(prenatalexposurevar), all_vars(!(. %in% c("Missing"))))
child_prenatal<-prenatalexdata%>%filter(childteen=='C')
teen_prenatal<-prenatalexdata%>%filter(childteen=='T')

secondhandsmokevars<-c(prenatalexposurevar, "PostnatalMaternalSmokingAny", "SmkAtVisitPastmonth")
secondhandsmkdata<-prenatalexdata%>%filter_at(all_of(secondhandsmokevars), all_vars(!(. %in% c("Missing"))))
child_secondhand<-secondhandsmkdata %>% filter(childteen=='C')
teen_secondhand<-secondhandsmkdata%>%filter(childteen=='T')
twovisit_secondhand<-secondhandsmkdata %>% filter(idnum %in% child_secondhand$idnum & idnum %in% teen_secondhand$idnum)

child_smoke<-secondhandsmkdata %>% filter(k5f1l!= "2 no" & childteen=='C')
child_nosmoke<-secondhandsmkdata %>% filter(k5f1l=="2 no" & childteen=='C')
teen_smoke<-secondhandsmkdata %>% filter(k6d40!="2 No" & childteen=='T')
teen_nosmoke<-secondhandsmkdata %>% filter(k6d40=="2 No" & childteen=='T')  
twovisit_nosmoke<-secondhandsmkdata %>% filter(idnum %in% child_nosmoke$idnum & idnum %in% teen_nosmoke$idnum)

fathersmkdata<-secondhandsmkdata %>% filter(f1g4!="Missing")

#Make Labels
n_fullFF<-paste0("Fragile Families cohort \n", "n=", FF_labeled %>% nrow())
n_methylQC<-paste0("Cohort with 450K methylation data \n", "n=", methylCohort$idnum%>%unique()%>%length(), "individuals with m=", methylCohort$MethID%>%length(), "samples")

n_QCloss<-paste0("Methylation data QC: 12 low-intensity samples dropped, \n", pdqc_all%>%filter(probe_fail_10==T)%>%nrow()-12, " samples with >10% of probes failing dropped, \n", pdqc_all%>%filter(sex!=predicted_sex & probe_fail_10==F & childteen!='M')%>%nrow(), " discordant sex samples dropped, \n & ", nrow(pdqc)-nrow(pdqc_clean), " duplicate samples dropped")
n_methylFF<-paste0("Cohort with quality controlled 450K data \n","n=", myFF %>% select(id) %>% unique() %>% nrow(), " individuals with m=", myFF%>%nrow(), " samples")

n_basemodel<-paste0("Base models: \n", "n=", basemodeldata %>% select(id) %>% unique() %>% nrow(), " individuals with m=", basemodeldata %>%nrow(), " samples")
n_baseexclusions<-paste0("n=",  myFF %>% filter_at(all_of(basemodelvar), any_vars((. %in%c("Missing", "Missing PC data"))))%>% select(id) %>% unique() %>% nrow(), " individuals & m=", myFF %>% filter_at(all_of(basemodelvar), any_vars((. %in%c("Missing", "Missing PC data")))) %>% nrow(), " samples missing principal components of ancestry; \n n=", myFF %>% filter_at(all_of(basemodelvar), any_vars((is.na(.))))%>%select(id)%>%unique()%>%nrow(), " individuals & m=", myFF %>% filter_at(all_of(basemodelvar), any_vars((is.na(.))))%>%nrow(), " samples missing child age at maternal/PCG interview")

n_prenatalexposures<-paste0("Prenatal exposures models: \n","n=", prenatalexdata %>% select(id) %>% unique() %>% nrow(), " individuals with m=", prenatalexdata %>%nrow(), " samples")
n_prenatalexclusions<-paste0("n=", basemodeldata %>% filter_at(all_of(prenatalexposurevar), any_vars(. %in% c("Missing"))) %>% select(id) %>% unique() %>% nrow(), " individuals & m=", basemodeldata %>% filter_at(all_of(prenatalexposurevar), any_vars(. %in% c("Missing"))) %>% nrow(), " samples missing \n data on maternal prenatal alcohol and drug use")

n_secondhandSmk<-paste0("Secondhand smoke exposure models: \n","n=", secondhandsmkdata %>% select(id) %>% unique() %>% nrow(), " individuals with m=", secondhandsmkdata %>%nrow(), " samples")
n_secondhandexclusions<-paste0("n=", prenatalexdata$id[!(prenatalexdata$id %in% secondhandsmkdata$id)] %>% unique()%>% length(), " individuals & m=", prenatalexdata %>% filter_at(all_of(secondhandsmokevars), any_vars(. %in% c("Missing"))) %>% nrow(), " samples missing data \n on maternal or primary care giver postnatal smoking")

n_Child<-paste0('n=', child_secondhand  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', child_nosmoke %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', child_smoke%>%nrow(), " individual \n smoking at 9 year visit \n or missing focal child smoking data")
n_Teen<-paste0('n=', teen_secondhand  %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', teen_nosmoke %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', teen_smoke%>%nrow(), " individuals \n smoking at 15 year visit \n or missing focal child smoking data")

n_sensFatherSmoking<-paste0("Sensitivity analysis: compare effects of maternal and paternal prenatal smoking \n", "n=", fathersmkdata %>% select(id) %>% unique() %>% nrow(), " individuals with m=", fathersmkdata %>%nrow(), " samples", "\n with prenatal paternal smoking data")
n_sensFatherSmokingEx<- paste0("n=", secondhandsmkdata$id[!(secondhandsmkdata$id %in% fathersmkdata$id)] %>% unique()%>% length(), " individuals & m=", secondhandsmkdata %>% filter(f1g4 %in% c("Missing")) %>% nrow(), " samples missing data on paternal prenatal smoking")

n_twovisits<-paste0("n=", twovisit_secondhand%>%pull(idnum) %>%unique() %>%length(), " individuals with both a 9 and 15 year visit")
n_twovisitsSmkExcluded<-paste0("Exclude children who smoke: \n", "n=", twovisit_nosmoke%>%pull(idnum) %>%unique() %>%length(), " individuals with both a 9 and 15 year visit")
n_twovisitSmkExclusions<-paste0("n=", (twovisit_secondhand%>%pull(idnum) %>%unique() %>%length())-(twovisit_nosmoke%>%pull(idnum) %>%unique() %>%length()), " individuals smoking at \n either 9 and 15 year visit or \n missing focal child smoking data")

```

```{r drawFlowchart}
graph1<-DiagrammeR::grViz("
   digraph graph2 {
   compound=true;
   graph [layout = dot]

   node [shape = rectangle, width=4]
   a [label = '@@1']
   b [label = '@@2']
   c [label = '@@3']
   d [label = '@@4']
   e [label = '@@5']
   f [label = '@@6']
   g [label = '@@7']
   h [label = '@@8']
   i [label = '@@9', group=g1]
   j [label = '@@10']
   k [label = '@@16']
   l [label = '@@17']
   m [label = '@@19']

   #fake nodes for connecting to edges
   node [shape=point, width=0.01, height=0.01]
   a1
   b1
   c1
   d1
#   e1
  f1 [group=g1]
  g1
  k1

  #exclusion nodes
  node [shape = rectangle, width=4]
  aa [label = '@@20']
  bb [label = '@@11']
  cc [label = '@@12']
  dd [label = '@@13']
  ff [label = '@@14', width=3]
  gg [label = '@@15', width=3]
  kk [label = '@@18', width=3]

  #draw connections
  a->m
  m->a1 [arrowhead=none]
  a1->b
  b->b1 [arrowhead=none]
  b1->c
  c->c1 [arrowhead=none]
  c1->d
  d->d1 [arrowhead=none]
  d1->e
  e->f; e->g; e->k
{rank=same; h->e [dir=back, minlen=2, style=dashed]}
  f->f1 [arrowhead=none]
  g->g1 [arrowhead=none]
  k->k1 [arrowhead=none]
  f1->i 
  g1->j 
  k1->l

{rank=same; a1->aa [minlen=2]}   
{rank=same; b1->bb [minlen=2]} 
{rank=same; c1->cc [minlen=2]}
{rank=same; d1->dd [minlen=2]}
{rank=same; f1->ff [minlen=2]}
{rank=same; g1->gg [minlen=2]}
{rank=same; k1->kk [minlen=2]}

#invisible constraints to force ordering
ff->i [style=invis]
   }

   [1]: n_fullFF
   [2]: n_methylFF
   [3]: n_basemodel
   [4]: n_prenatalexposures
   [5]: n_secondhandSmk
   [6]: n_Child
   [7]: n_Teen
   [8]: n_sensFatherSmoking
   [9]: n_ChildSmkexcluded
   [10]: n_TeenSmkexcluded
   [11]: n_baseexclusions
   [12]: n_prenatalexclusions
   [13]: n_secondhandexclusions
   [14]: n_ChildSmkexclusions
   [15]: n_TeenSmkexclusions
   [16]: n_twovisits
   [17]: n_twovisitsSmkExcluded
   [18]: n_twovisitSmkExclusions
   [19]: n_methylQC
   [20]: n_QCloss
")
```

```{r drawGraph, fig.height=11, fig.1idth=8}
graph1
```

## Comparison of analytic sample: whole sample of Fragile Families vs methylation data vs analytic subset

```{r pick analysis subset}
analysis_data<-rbind(child_nosmoke, teen_nosmoke)
```

```{r, warning=F, message=F}
methyl1<-createTable(compareGroups(MethylData~smkPreg_binary+cm1inpov+cm1ethrace+cm1bsex+m1g2_YesNoPreg+m1g3_YesNoPreg+PostnatalMaternalSmokingAny+k5f1l+k6d40+f1g4, data=FF_labeled), show.all=T)
export2md(methyl1, caption="Exclusion table of all Fragile Families vs those with quality controlled methylation data for main predictors and key demographic variables from first wave")
```

```{r final analytic subset, warning=F, message=F}
myFF$MethylData2<-ifelse(myFF$MethID %in% c(child_nosmoke$MethID, teen_nosmoke$MethID), 'Final analysis group', 'Not in final analysis group')
set_label(myFF$MethylData2)<-'Exclusions based on missing data'
methyl2<-createTable(compareGroups(MethylData2~smkPreg_binary+globalmethylation+pediatric+Epi+IC+anynewborn_center+SSnewbornCT_center+SSolder_center+cm1inpov+ancestry+cm1bsex+m1g2_YesNoPreg+m1g3_YesNoPreg+PostnatalMaternalSmokingAny+SmkAtVisitPastmonth, data=myFF), show.all=T)
export2md(strataTable(methyl2, 'childteen'), caption="Exclusion table of subset with quality controlled methylation data vs final analytic subset for main predictor, outcomes and key demographic variables")

```

# Datasets for modeling
Variables

Base model: Binary smoking variable [X], ancestry categorization, baby sex [X], maternal income to poverty ratio [X],  child's age in month at visit of interest (note, currently using a constructed child's age at visit, where individuals having a missing child's age at home visit was supplemented with child's age at home interview) [X], cell types (Epi + IC, may need to take one out),

Prenatal exposures: Base model variables + prenatal maternal alcohol use (Yes/no) [X], prenatal maternal drug use (yes/no)

Secondhand smoke exposure models: Prenatal exposures model+ any maternal smoking at age 1 & 5 [X], maternal or PCG smoking dose at visit of interest (age 9 or 15) [X]

First hand smoking: Secondhand smoke exposure models + exclude children with missing or reported firsthand smoking at visit of interest (age 9 or 15)

Paternal smoking: Filter to individuals with data on prenatal paternal smoking and compare effect size of prenatal paternal to prenatal maternal (as per Wiklund et al. 2019, idea is that this captures additional confounding)

Ancestry specific models - by ancestry groups; replace ancestry control with control for local PCs within ancestry group

Longitduinal models - no additional filtering needed

```{r, saveData}
save(child_base, file=paste0(datadir, 'CreatedData/', 'ChildBaseModelData.Rdata'))
save(teen_base, file=paste0(datadir, 'CreatedData/', 'ChildBaseModelData.Rdata'))
save(basemodeldata, file=paste0(datadir, 'CreatedData/', 'BaseModelData.Rdata'))

save(child_prenatal, file=paste0(datadir, 'CreatedData/', 'ChildPrenatalExModelData.Rdata'))
save(teen_prenatal, file=paste0(datadir, 'CreatedData/', 'PrenatalExModelData.Rdata'))
save(prenatalexdata, file=paste0(datadir, 'CreatedData/', 'PrenatalExModelData.Rdata'))

save(child_secondhand, file=paste0(datadir, 'CreatedData/', 'ChildSecondHandSmkModelData.Rdata'))
save(teen_secondhand, file=paste0(datadir, 'CreatedData/', 'TeenSecondHandSmkModelData.Rdata'))
save(twovisit_secondhand, file=paste0(datadir, 'CreatedData/', 'TwoVisitSecondHandSmkModelData.Rdata'))
save(secondhandsmkdata, file=paste0(datadir, 'CreatedData/', 'SecondHandSmkModelData.Rdata'))

save(child_nosmoke, file=paste0(datadir, 'CreatedData/', 'ChildNoSmkModelData.Rdata'))
save(teen_nosmoke, file=paste0(datadir, 'CreatedData/', 'TeenNoSmkModelData.Rdata'))
save(twovisit_nosmoke, file=paste0(datadir, 'CreatedData/', 'TwoVisitNoSmkModelData.Rdata'))

save(fathersmkdata, file=paste0(datadir, 'CreatedData/', 'FatherSmkModelData.Rdata'))

save(list=c("basemodeldata", "prenatalexdata", "secondhandsmkdata", "child_secondhand", "teen_secondhand", "twovisit_secondhand", "twovisit_nosmoke", "child_nosmoke", "teen_nosmoke", "fathersmkdata"), file=paste0(datadir, 'CreatedData/', Sys.Date(), 'CreatedData.Rdata'))
```