### 2021-09-07 Reports/Sept22FF.Rmd

---
title: "September22_FFReport"
author: "F blostein"
date: "9/23/2020"
output: html_document
---
Goals: 
1. Create linked dataset with variables of interest from full pheno data 
2. Create table comparing various analytic subsets (methylation data; two-visit methylation data) to full Fragile Families dataset 
3. Address technical replicates

Questions for Kelly: 
1. Review plan for principal components of ancestry control 
2. Check plan for technical replicates 
3. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/scratch/bfoxman_root/bfoxman/blostein/FragileFamilies/")
library(tidyr)
library(dplyr)
library(compareGroups)
library(kableExtra)
```

```{r}
pheno<-readRDS("/scratch/bfoxman_root/bfoxman/blostein/FragileFamilies/Files/OGData/fullpheno.rds")
#betaqc<-readRDS("/scratch/bfoxman_root/bfoxman/blostein/FragileFamilies/Files/OGData/betaqc.rds")
pdqc<-readRDS("/scratch/bfoxman_root/bfoxman/blostein/FragileFamilies/Files/OGData/pdqc.rds")
```

```{r}
source("/scratch/bfoxman_root/bfoxman/blostein/FragileFamilies/Code/Reports/DataDictionary.R")
```

```{r, fullpheno}
#check column names of pheno for restricted health data which appear
# to begin with suffix ch (i.e. chc10e (tobacco use-ever) and
#chc10p (tobacco use-during preg)
colnames(pheno)[grep('ch', colnames(pheno))] 
#find the id varaibles
colnames(pheno)[grep('id', colnames(pheno))] 
#filter for variables of interest

FF<-pheno %>% select(one_of(idvars), one_of(demovars), one_of(smokingvars), one_of(prenataldrugusevar))
#Warning message:
#Unknown columns: `p3a22`, `p3a23`, `p3a23a`, `p3a23b`, `p3a24`, `p4a22`, `p4a23` 
```

```{r, idgroups}
#What FF ids have methylation data?
methy_ids<-pdqc$id
#technical replicates will be duplicate id + childteen variables from pdqc 
techrep_ids<-pdqc %>%select(idnum, childteen, MethID)%>%group_by(idnum, childteen)%>%filter(n()>1)%>%pull(MethID)
#ids with 9visit, 15visit or both visit
visit9ids<-pdqc%>%filter(childteen=='C')%>%pull(idnum)%>%unique()
visit15ids<-pdqc%>%filter(childteen=='T')%>%pull(idnum)%>%unique()
bothvisit_ids<-visit9ids[visit9ids %in% visit15ids]
visit9only_ids<-visit9ids[!(visit9ids %in%bothvisit_ids)]
visit15only_ids<-visit15ids[!(visit15ids %in%bothvisit_ids)]
```

```{r}
#create flag in FF for having methylation data
#and flag for having 2 visits of methylation data
FF<-FF %>% 
  mutate(MethylData=ifelse(id %in% methy_ids, "Analysis subset", "Not in analysis subset"),
         TwoVisitData=ifelse(id%in%bothvisit_ids,"Analysis subset - both visits", "Not in both visit analysis subset"))
table(FF$MethylData)
table(FF$TwoVisitData)
```

```{r, checks}
if((length(visit15only_ids)+length(visit9only_ids)+length(bothvisit_ids)!=length(FF %>% filter(MethylData=="Analysis subset")%>%pull(id)))){print("WARNING ID LENGTHS DO NOT ADD UP")}
if(length(bothvisit_ids)!=length(FF %>% filter(TwoVisitData=="Analysis subset - both visits")%>%pull(id))){print("WARNING ID LENGTHS DO NOT ADD UP")}
```

```{r, labelMetaData}
myBinary<-myBinary[myBinary %in% colnames(FF)]
myCategorical_O<-myCategorical_O[myCategorical_O %in% colnames(FF)]
myCategorical_U<-myCategorical_U[myCategorical_U%in%colnames(FF)]
FF<-FF %>% mutate_at(vars(myBinary, myCategorical_O, myCategorical_U), funs(as.factor(.)))
```

```{r}
methyl<-createTable(compareGroups(MethylData~.-id-ff_id, data=FF), show.all=T)
twovisit<-createTable(compareGroups(TwoVisitData~.-id-ff_id, data=FF), show.all=T)
my_FF<-FF %>% filter(MethylData=="Analysis subset")%>%mutate(smkPreg_binary=ifelse(m1g4%in% c(1, 2, 3), "Yes", ifelse(m1g4%in%c(-3, -2, -9), NA, ifelse(m1g4==4, "No", "FLAG"))))
xtabs(~m1g4+smkPreg_binary, data=my_FF)
smk<-createTable(compareGroups(smkPreg_binary~.-id-ff_id, data=my_FF))
```

```{r}
export2md(methyl)
export2md(twovisit)
export2md(smk)
```

```{r}
#explore technical replicates
#pdqc %>% filter(MethID %in% techrep_ids) %>% table(~qc_flag_any, data=.)
techrep_qcF<-pdqc %>% filter(MethID %in% techrep_ids & qc_flag_any==T)
techrep_qcP<-pdqc %>% filter(MethID %in% techrep_ids & qc_flag_any==F)
```

```{r}

```