### 2021-09-08 Reports/Poster/IGES_poster_vertical.Rmd

---
main_topsize: 0.1 #percent coverage of the poster
main_bottomsize: 0.05
#ESSENTIALS
title: Polymethylation Scores for Prenatal Maternal Smoke Exposure Persist Until Age
  15 and Are Detected in Saliva
title_textsize: 100
author:
- name: Freida Blostein
  affil: 1
  twitter: blostein
  email: blostein@umich.edu
  orcid: 0000-0002-1099-3857
  main: yes
- name: Jonah Fisher
  affil: 2
- name: John Dou
  affil: 1
- name: Lisa Schenper
  affil: 3
- name: Daniel A. Notterman
  affil: 3
- name: Colter Mitchell
  affil: 2
- name: Kelly M. Bakulski
  affil: 1
affiliation:
- num: 1
  address: Department of Epidemiology, University of Michigan School of Public Health,
    Ann Arbor, Michigan, USA
- num: 2
  address: Institute for Social Research, University of Michigan, Ann Arbor, Michigan,
    USA
- num: 3
  address: Department of Molecular Biology, Princeton University, Princeton, New Jersey,
    USA
main_findings:
  - "**Polymethylation scores** for prenatal maternal smoke exposure **persist over time** and are **robust across ancestries**"
logoleft_name: 'Images/Epid_informal_horiz_white.png'
logoright_name: 'Images/Fragile_Families_Logo.png'
logocenter_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/qr-code-black.png
primary_colour: "#0048A4"
secondary_colour: "#5885AF"
accent_colour: "#E35A01"
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: packages.bib
link-citations: true
---

<style>
#main-img-left {
 width: 20%
}
#main-img-center {
 width: 2%
}
#main-img-right {
width=20%
}
</style>

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

# Introduction

This is the `posterdown_betterport` template for the {posterdown} package! I was inspired by the twitter thread of [Mike Morrison](https://mobile.twitter.com/mikemorrison/status/1110191245035479041) and wanted to apply the `#betterposter` concept to the reproducible (yet simple to use) functionality of the {posterdown} package [@R-posterdown]. If you're not an R user don't sweat as you do **NOT** need to use it at all! Feel free to use only the Markdown functionality of this package :)

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

## Gaps

1. Pick a template layout.
2. Write/ create your poster content distraction free.
3. Let posterdown do its thing!

# Methods

I will show here how to include poster elements that may be useful, such as an equation using mathjax:

$$
E = mc^2
$$

To reference a citation you can add your `.bib` file to the working directory and name it in the YAML metadata or generate an automated one as done here, then you only need to reference the label value in the `.bib` file. For example this package is built on top of the wonderful {pagedown} package and I will cite it at the end of this sentance using this in the rmd `[@R-pagedown]` [@R-pagedown].

To get a better understanding of how to include features like these please refer to the {posterdown} [wiki](https://github.com/posterdown/wiki).

**_Now on to the results!_**

# Results

Here you may have some figures to show off, bellow I have made a scatterplot with the infamous Iris dataset and I can even reference to the figure automatically like this, `Figure \@ref(fig:irisfigure)`, Figure \@ref(fig:irisfigure).

```{r, irisfigure, fig.cap='Here is a caption for the figure. This can be added by using the "fig.cap" option in the r code chunk options, see this [link](https://yihui.name/knitr/options/#plots) from the legend himself, [Yihui Xie](https://twitter.com/xieyihui).', out.width="80%"}
par(mar=c(2,2,0,1))
plot(x = iris$Sepal.Length, y = iris$Sepal.Width,
     col = iris$Species, pch = 19, xlab = "Sepal Length",
     ylab = "Sepal Width")
```

Maybe you want to show off some of that fancy code you spent so much time on to make that figure, well you can do that too! Just use the `echo=TRUE` option in the r code chunk options, Figure \@ref(fig:myprettycode)!

```{r myprettycode, echo=TRUE, fig.cap='Boxplots, so hot right now!', fig.height=6, out.width="80%"}
#trim whitespace
par(mar=c(2,2,0,0))
#plot boxplots
boxplot(iris$Sepal.Width~iris$Species,
        col = "#008080", 
        border = "#0b4545",
        ylab = "Sepal Width (cm)",
        xlab = "Species")
```

How about a neat table of data? See, Table \@ref(tab:iristable):

```{r, iristable}
knitr::kable(
  iris[1:15,1:5], format = "html",
  caption = "A table made with the **knitr::kable** function.",
  align = "c", col.names = c("Sepal <br> Length",
                             "Sepal <br> Width",
                             "Petal <br> Length",
                             "Petal <br> Width",
                             "Species"),
  escape = FALSE)
```

# References