### 2021-09-13 Papers/Polymethylation_scores/98_declarations.Rmd
added paper sections

# Declarations

## Ethics approval and consent to participate

## Consent for publication

Not applicable