### 2021-09-30 Reports/ROC_report.Rmd
September 30 updated pipeline and results

---
title: "ROC curve analysis"
author: "F blostein"
date: "9/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source()
```

```{r}
#investigate adding in covariates and effect on ROC curve 
#question for kelly --> only use methylation or use methylation + base model
#don't think it's reasonable to use prenatal exposures + postantal smoking 
# could we predict peoples prenatal smoke exposure when we don't have info on maternal smoking/behaviors

global_roc%>%dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric'))%>%
  ggplot(aes(x=spec, y=sens, color=covariates))+geom_line()+
  facet_grid(childteen~methylation, labeller=labeller(childteen=age_labels, methylation=outcome_labels))+
  theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=6,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')
```

```{r}

#is prediction better when using PMS? 
global_roc%>%dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(str_detect(predictors, '\\+'), covariates, 'No covariates'))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') &
           !str_detect(covariates, 'm1g2|sv'))%>%
  ggplot(aes(x=spec, y=sens, color=methylation, linetype=modeltype))+geom_line()+
  facet_grid(childteen~., labeller=labeller(childteen=age_labels))+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')
```

```{r}

global_roc%>%dplyr::select(predictors, covariates, methylation, childteen, test)%>%
  mutate(modeltype=ifelse(str_detect(predictors, '\\+'), covariates, 'No covariates'))%>%
  unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') &
           covariates=='')%>%
  ggplot(aes(x=spec, y=sens, color=methylation, linetype=modeltype))+geom_line()+
  facet_grid(childteen~., labeller=labeller(childteen=age_labels))+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')

```

```{r}
#is best methylation predictor consistent across ancestries
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates=='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()

#things get weird when using the base model vars
local_roc %>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'SSnewbornCT|cg055|global|pediatric') & covariates!='')%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()
```

```{r}
#investigated change in ROC when adding in each base model var individually
local_roc_test%>% dplyr::select(covariates, predictors, childteen, ancestry, test)%>%unnest(test)%>% 
  filter(str_detect(predictors, 'global'))%>%
  ggplot(aes(x=spec, y=sens, color=predictors))+
  facet_grid(childteen~ancestry)+geom_line()
#big jump in ROC curve for European & hispanic - batch variable 
#big jump in ROC curve for European - poverty and hispanic - local pcs
modeldata %>%filter(ancestry=='Hispanic ancestry' & childteen=='T')%>%xtabs(~Sample_Plate+smkPreg_binary, data=.)

```

```{r}
#is prediction improved when using two time points?

global_roc%>%dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') & covariates=='')%>%
  rbind(long_global_roc%>% mutate(childteen='C+T', methylation=gsub('_C.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>%
          unnest(test) %>% 
          filter(str_detect(methylation, 'SSnewbornCT|cg055|global|pediatric') & covariates==''))%>%
  ggplot(aes(x=spec, y=sens, color=childteen))+geom_line()+
  facet_wrap(~methylation)+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')

global_roc%>%dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT|cg055') & covariates=='')%>%
  rbind(long_global_roc%>% mutate(childteen='C+T', methylation=gsub('_C.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>%
          unnest(test) %>% 
          filter(str_detect(methylation, 'SSnewbornCT|cg055') & covariates==''))%>%
  ggplot(aes(x=spec, y=sens, color=childteen))+geom_line()+
  facet_wrap(~methylation)+
  scale_color_discrete(labels=outcome_labels)+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')

```