### 2021-10-14 Reports/Results.Rmd
05 ewas script and results section improvements

---
title: "Polymethylation scores results"
author: "F blostein"
date: "9/30/2021"
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: no
    theme: sandstone
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(ggplot2)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(stringr)
library(dplyr)
library(compareGroups)
library(kableExtra)
library(gtsummary)
library(gtools)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(purrr)
library(broom)
library(broom.mixed)
library(ggplot2)
library(multcomp)
library(ggpubr)
library(cowplot)
library(geepack)
library(lme4)
library(pROC)
library(lmerTest)
library(stringr)
library(sva)
```

```{r temp read n data}
codedir=''
datadir="/nfs/turbo/bakulski1/People/blostein/FF_methylation/Data"
load(paste0(datadir, '/CreatedData/allPhenoData.Rdata'))
load(paste0(datadir, '/CreatedData/allmethyl.Rdata'))
load(paste0(datadir, '/CreatedData/completeCaseSVA.Rdata'))
load(paste0(datadir, '/CreatedData/methylation_summarymodels.Rdata'))
```

```{r plotting variables}
theme_set(theme_bw())
```

```{r}
n_total=modeldata %>% dplyr::select(id)%>%unique()%>%nrow()
n_obs=modeldata%>%nrow()

all_methyl=c("globalmethylation", "pediatric", "grim", "Epithelial.cells_saliva", "Leukocytes_saliva", "SSnewborn_center", "SSnewbornCT_center", "SSolder_center", "cg05575921")
main_methyl=c("globalmethylation", "pediatric", "SSnewbornCT_center", "cg05575921")

all_methyl_labels=modeldata%>%dplyr::select(any_of(all_methyl))%>%get_label()
all_methyl_labels[str_detect(all_methyl_labels, 'Polymethylation', )]=paste0(rep('Score:', 3), c('Sustained smoking \n newborns', 'Sustained smoking \n newborns (w/ CT control)', 'Sustained smoking \n older children'))
names(all_methyl_labels)=all_methyl

all_methyl_levels=c("SSnewbornCT_center", "SSnewborn_center",  "SSolder_center",
                                       "pediatric" , "grim", "cg05575921", 
                                      "globalmethylation","Epithelial.cells_saliva", "Leukocytes_saliva")

main_methyl_labels=modeldata%>%dplyr::select(any_of(main_methyl))%>%get_label()

my_round=function(x, d){
  y=c()
  d_orig=d
  for(i in 1:length(x)){
    d=d_orig
    while(round(x[i], d)==0){
    d=d+1
    }
    y[i]=round(x[i], d)
  }
  return(y)
}
```

# Current issues

# Target journal and deadline

[Clinical epigenetics](https://www.biomedcentral.com/collections/epibio): Special issue on biomarkers; due date December 1st
 >Epigenetic biomarkers have significantly contributed to improved understanding of the origins and progression of disease.  Moreover, increasing evidence shows that epigenetic biomarkers have potential for personalized medicine. However, routine use of epigenetic biomarkers in in-vitro diagnostics (IVD) is lagging behind the discoveries of new biomarkers. In this thematic series of Clinical Epigenetics, we discuss current progress in development of epigenetic biomarkers for clinical use in common complex diseases of ageing and welcome both original research and reviews.
The sections of the collection include:
  - Types of epigenetic biomarkers and readiness for IVD use
    - DNA methylation as biomarker
    - Histone modification as biomarker
    - Upcoming field of circulating free nucleosomes as biomarker
    - Progress towards clinical use of biomarker
  - *Biomarkers of environmental exposures and causality of the biomarker*
  - Disease predisposition biomarkers
  - Biomarkers used for disease detection
  - Clinical disease prevention and management: predictive and prognostic
  - Monitoring of chronic disease
  - *Single versus multiple biomarkers and risk scores*
  - Major stages of clinical validation of the biomarker
  - Biomarker detection technologies suitable for IVD use and significance of bioinformatics
  - Ethics and regulatory aspects of IVD epigenetic biomarkers

# Figure 1 - Selection of samples from the Fragile Families and Child Wellbeing (FFCW) study into this analysis. N represents the number of individuals at each step in the selection procedure, M represents the number of samples, as individuals with repeated measures can have more than one sample.

```{r, eval=T}
#function
N_M_count=function(filter_statement, statement, df){
  paste0('m=', df %>% filter(eval(parse(text = filter_statement)))%>%nrow(), 
         ' samples & ', 
         'n=', df %>% filter(eval(parse(text = filter_statement))) %>% dplyr::select(idnum) %>%unique()%>%nrow(), 
         ' individuals with', statement)
}
#Make Labels
n_fullFF<-paste0("Fragile Families cohort \n", "n=", FF_labeled %>% nrow())
n_methylQC<-paste0("Cohort with 450K methylation data \n", "n=", methylCohort$idnum%>%unique()%>%length(), " individuals with m=", methylCohort$MethID%>%length(), " samples")

n_QCloss<-paste0("Methylation data QC: ", 
                 N_M_count("probe_fail_10==T & childteen!='M'", " with >10% of probes failing dropped, \n ", pdqc_all),
                 N_M_count("sex_new!=predicted_sex & probe_fail_10==F & childteen!='M'", 
                           "  with discordant sex samples dropped, \n ", pdqc_all),
                 N_M_count("predicted_sex_outlier==T & sex_new==predicted_sex & probe_fail_10==F & childteen!='M'", 
                           ' with abnormal sex chromosome intensity dropped & \n ', pdqc_all),
                 length(reps_remove), " duplicate samples dropped")

n_methylFF<-paste0("Cohort with quality controlled 450K data \n","n=", myFF %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", myFF%>%nrow(), " samples")

n_missingmaternalprenataldata=paste0('Missing maternal data: \n', 
                                     N_M_count("is.na(smkPreg_binary)", " with missing prenatal maternal smoking \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & is.na(m1g2_YesNoPreg)", " with missing prenatal alcohol \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & is.na(m1g3_YesNoPreg)", 
                                               " with missing prenatal other drug use \n ", myFF),
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & is.na(cm1inpov)", 
                                               " with missing maternal income to poverty ratio data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny=='Missing'", " with missing postnatal maternal smoking data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth=='Missing'", 
                                               " with missing past month maternal/primary care giver smoking packs/day", myFF))

maternaldata=myFF %>% filter(!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) & PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth!='Missing')  
n_withmaternaldata=paste0('m=', maternaldata %>% nrow(), ' samples & n=', maternaldata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with maternal covariate data')

n_missingchilddata=paste0('Missing child data: \n', 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data'", 
                                    " with missing ancestry/genetic data  \n ", maternaldata), 
                          N_M_count("!is.na(ancestry) & ancestry!='Missing PC data' & is.na(cm1bsex)", 
                                    " with missing child sex data \n ", maternaldata), 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data' & !is.na(cm1bsex) & is.na(ChildAgeComposite)", 
                                    " with missing child age at sample \n ", maternaldata))

childdata= maternaldata %>% filter(ancestry!='Missing PC data' & !is.na(cm1bsex) & !is.na(ChildAgeComposite))
n_withchilddata=paste0('m=', childdata %>% nrow(), ' samples & n=', childdata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with child covariate data')

n_Child<-paste0('n=', childdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata%>%filter(childteen=='C' & k5f1l=='2 no')%>% nrow(), 
                           " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', childdata%>%filter(k5f1l!= "2 no" & childteen=='C')%>%nrow(), 
                             " individuals smoking \n or missing child smoking data \n at 9 year visit")
n_Teen<-paste0('n=', childdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata %>%filter(childteen=='T' & k6d40=="2 No") %>% nrow(), 
                          " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', childdata %>% filter(k6d40!="2 No" & childteen=='T') %>% nrow,
                            " individuals smoking \n or missing focal child smoking data \n at 15 year visit") 
if(childdata %>% filter((childteen=='C' & k5f1l=='2 no') | (childteen=='T' & k6d40=='2 No'))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

```

```{r}
graph1<-DiagrammeR::grViz("
   digraph graph2 {
   compound=true;
   graph [layout = dot]

   node [shape = rectangle, width=4]
   a [label = '@@1']
   b [label = '@@2']
   c [label = '@@3']
   d [label = '@@4']
   e [label = '@@5']
   f [label = '@@6']
   g [label = '@@7']
   h [label = '@@11']
   i [label = '@@12']

   #fake nodes for connecting to edges
   node [shape=point, width=0.01, height=0.01]
   a1
   b1
   c1
   d1
   f1 [group=g1]
   g1
   h1

  #exclusion nodes
  node [shape = rectangle, width=4]
  aa [label = '@@8']
  bb [label = '@@9']
  cc [label = '@@10']
  hh [label = '@@13']
  ii [label = '@@14']

  #draw connections
#  a->m
#  m->a1 [arrowhead=none]
  a->a1 [arrowhead=none]
  a1->b
  b->b1 [arrowhead=none]
  b1->c
  c->c1 [arrowhead=none]
  c1->d
  d->d1 [arrowhead=none]
  d1->e
  e->f; e->g
  f->g1 [arrowhead=none]; g->h1 [arrowhead=none]
  g1->h; h1->i

{rank=same; b1->aa [minlen=2]}   
{rank=same; c1->bb [minlen=2]} 
{rank=same; d1->cc [minlen=2]}
{rank=same; g1->hh [minlen=2]}
{rank=same; h1->ii [minlen=2]}

#invisible constraints to force ordering
#ff->i [style=invis]
   }

   [1]: n_fullFF
   [2]: n_methylQC
   [3]: n_methylFF
   [4]: n_withmaternaldata
   [5]: n_withchilddata
   [6]: n_Child
   [7]: n_Teen
   [8]: n_QCloss
   [9]: n_missingmaternalprenataldata
   [10]: n_missingchilddata
   [11]: n_ChildSmkexcluded
   [12]: n_TeenSmkexcluded
   [13]: n_ChildSmkexclusions
   [14]: n_TeenSmkexclusions
")

graph1
```

# Supplemental Table 1 - Descriptive statistics of excluded vs included samples with methylation data available from the Fragile and Families and Child Wellbeing study

```{r}
myFF$MethylData2<-ifelse(myFF$MethID %in% modeldata$MethID, 'Final analysis group', 'Not in final analysis group')
set_label(myFF$MethylData2)<-'Exclusions based on missing data'
myFF=left_join(myFF, methyldata%>%mutate(idnum=as.character(idnum)))
methyl2<-createTable(compareGroups(MethylData2~smkPreg_binary+m1g2_YesNoPreg+m1g3_YesNoPreg+cm1inpov+PostnatalMaternalSmokingAny+SmkAtVisitPastmonth+ancestry+cm1bsex+globalmethylation+pediatric+Epithelial.cells_saliva+Leukocytes_saliva+SSnewbornCT_center+SSolder_center+cg05575921, data=myFF, include.miss = T), show.all=T)
export2md(strataTable(methyl2, 'childteen'), caption="Exclusion table of subset with quality controlled methylation data vs final analytic subset for main predictor, outcomes and key demographic variables")
```
# Supplemental Figure - Comparison of choice of A) regression coefficient and B) transform of methylation beta value on distributions of polymethylation scores

```{r}
scores=gsub('_center', '', colnames(modeldata)[str_detect(colnames(modeldata), 'center')])
age_long_score=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary')),
                contains(scores))%>%
  pivot_longer(cols=contains(scores),
               names_to='Scores',
               values_to='Polymethylation score')

plot_grid(
  age_long_score%>%filter(str_detect(Scores, 'no')) %>% 
    ggplot(aes(x=`Polymethylation score`))+geom_histogram()+
    facet_wrap(~Scores, labeller=labeller(Scores=get_label(modeldata%>%dplyr::select(contains('notransform'))))), 
  age_long_score%>%filter(str_detect(Scores, 'CT')) %>% 
    ggplot(aes(x=`Polymethylation score`))+geom_histogram()+
    facet_wrap(~Scores, scales='free', ncol=3,
               labeller=labeller(Scores=get_label(modeldata%>%dplyr::select(contains('CT'))))),
  labels=c('A', 'B'), nrow=2
)

```

# Supplemental Figure 2 - Correlation of methylation summary measures

```{r}
library(Hmisc)
library(corrplot)
res2<-modeldata%>%dplyr::select(any_of(all_methyl))%>%cor()
#labels
colnames(res2)=all_methyl_labels
rownames(res2)=all_methyl_labels
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)
```

#Supplemental Figure 3 - Correlation between methylation summary measures between visits

```{r}
age_long_methyl=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))

age_wide_methyl=age_long_methyl%>%
  pivot_wider(names_from=childteen, 
              values_from=`Methylation summary measure`)

age_wide_methyl%>%
  ggplot(aes(x=C, y=T))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))
```

# Supplemental Figure XX: Change in methylation summary measures between age 9 and age 15

```{r}
age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..")+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))+
  theme(legend.position = 'bottom')

age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, group=idnum))+
  geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))+
  theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))
```

# Figure 2: Polymethylation scores are correlated and stable between age 9 and age 15 in a diverse sample of `r n_total` US children

```{r change in scores time, fig.width=12}

#add line for mean ? 
outcome_labels=c('Global methylation', 'Pediatric clock', 'AHHR: cg05575921', 'Polymethylation score \n (newborn, mean centered)', 'Polymethylation score  \n (older children, mean centered')
names(outcome_labels)=c('globalmethylation', 'pediatric', 'cg05575921', 'SSnewbornCT_center','SSolder_center')

fig2a=age_wide_methyl%>%filter(Methyl_factor%in% main_methyl)%>%
  ggplot(aes(x=C, y=T))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), nrow=1)

mean_summary=age_wide_methyl%>%filter(Methyl_factor%in% main_methyl)%>%mutate(diff=T-C)%>%group_by(Methyl_factor)%>%summarise(mean_diff=mean(diff, na.rm=T))

fig2b=age_wide_methyl%>%filter(Methyl_factor%in% main_methyl)%>%mutate(diff=T-C)%>%
  ggplot(aes(x=diff))+geom_histogram()+
  facet_wrap(~Methyl_factor, scales='free_x', labeller = labeller(Methyl_factor=all_methyl_labels), nrow=1)+
  theme_classic()+xlab('Difference in measure (Age 15 - Age 9)')+
  geom_vline(aes(xintercept=0), color='red', linetype=2)+geom_rug()+
  geom_vline(data=mean_summary, aes(xintercept=mean_diff), color='blue')

plot_grid(fig2a, fig2b, nrow=2, align = 'v', axis = 'lr')

```

# Table 1: Bivariate associations between prenatal maternal smoking, DNA methylation summary measures and important covariates among a diverse sample of

```{r bivariate table}
export2md(strataTable(createTable(compareGroups(smkPreg_binary~m1g2_YesNoPreg+m1g3_YesNoPreg+cm1inpov+PostnatalMaternalSmokingAny+SmkAtVisitPastmonth+ancestry+cm1bsex+globalmethylation+pediatric+Epithelial.cells_saliva+Leukocytes_saliva+SSnewbornCT_center+SSolder_center+cg05575921, data=modeldata%>%mutate(childteen=factor(case_when(childteen=='C'~'Age 9', childteen=='T'~ 'Age 15'), levels=c('Age 9', 'Age 15'))), include.miss = T)), 'childteen'))
```

# Supplemental Figure: *A priori* CpG sites methylation and prenatal maternal smoking by age and ancestry

# Supplemental Figure - A) Ancestry and B) age differences in methylation summary measures by prenatal maternal smoking exposure

```{r}
plot_grid(
  age_long_methyl%>%filter(Methyl_factor %in% main_methyl)%>%
    ggplot(aes(x=smkPreg_binary, y=`Methylation summary measure`, color=ancestry))+
    geom_violin()+geom_boxplot(width=0.2, position=position_dodge(0.9))+stat_compare_means(label="..p.signif..")+
    scale_color_discrete(name='Ancestry')+
    scale_x_discrete(name='Prenatal smoke exposure', labels=c('No', 'Yes'))+
    facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))+
    theme(legend.position = 'bottom'),

age_long_methyl%>%filter(Methyl_factor %in% main_methyl)%>%
  ggplot(aes(x=smkPreg_binary, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2, position=position_dodge(0.9))+stat_compare_means(label="..p.signif..")+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Prenatal smoke exposure', labels=c('No', 'Yes'))+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels))+
  theme(legend.position = 'bottom'),

labels=c('A', 'B'), nrow=1)
```

# Supplemental Table XX: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable models among `r n_total` US children

```{r}
my_main_outcomes=c('AHHR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)', 'Sustained smoking polymethylation score (older children)', 'Pediatric clock', 'Global methylation')

test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='secondhand smoke exposure model')%>%
  filter(outcome %in% my_main_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='C'~'Age 9 models', 
                             childteen=='T'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age=tbl_stack(table_list, group_header=names(table_list))

```

```{r}
test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='secondhand smoke exposure model')%>%
  filter(outcome %in% my_main_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry=tbl_stack(table_list, group_header=names(table_list))

```

```{r, width=11, eval=F}
#alternatively 
my_names=c('Methylation measure', paste0(c(paste0(paste0('Unstratified <br> n=', n_total), footnote_marker_alphabet(1, 'html')), paste0(modeldata%>%group_by(ancestry)%>%dplyr::select(ancestry, idnum)%>%unique()%>%summarise(n=n())%>%transmute(label=paste0(ancestry, ' <br> n=', n))%>%pull(label), footnote_marker_alphabet(2, 'html'))), '<br> Beta (95% CI); p-value'))

tb2=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'), longitudinal_local_models%>%dplyr::select(-data))%>%filter(predictors=='secondhand smoke exposure model')%>%filter(outcome %in% my_main_outcomes)%>%unnest(tidied)%>%filter(term=='smkPreg_binaryYes')%>%mutate(estimate=paste0(my_round(estimate, 2), ' (', my_round(conf.low, 2), ', ', my_round(conf.high, 2), ') p=', signif(p.value, 2)))%>%dplyr::select(ancestry, outcome, estimate)%>%pivot_wider(names_from=ancestry, values_from=c(estimate))

names(tb2)=my_names

tb2%>%kable(., escape=F)%>%kable_minimal()%>%
  kableExtra::scroll_box(width = "100%")%>% footnote(alphabet =  c("Model adjusted for:child sex, mother income to poverty ratio at birth, epithelial and immune cell proportions (estimated from methylation data), maternal other drug and other alcohol use during pregnancy, batch of array, genetic ancestry (1st two principal components of genetic data), postnatal smoke exposure (any postnatal maternal smoking and maternal packs per day in month prior to visit).", "Models adjusted as in (a) but principal components of genetic ancestry rerun within ancestry strata and first two components from these stratified principal components analyses used"))%>%kable_styling(font_size = 10)

```

# Figure 3 - Polymethylation scores are associated with prenatal maternal smoke exposure consistently across age and ancestry group in a sample of `r n_total` US children

```{r, fig.width=12}
my_outcomes=c('Global methylation', 'Pediatric clock', 'AHHR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)')

outcome_labels=c('Global methylation', 'Pediatric clock', 'AHHR: cg05575921',  'Polymethylation score \n (newborn, mean centered)')

all_models=rbind(local_models, global_models%>%mutate(ancestry='Unstratified'))

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.25)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='secondhand smoke exposure model')%>%mutate(outcome_f=outcome)

names(outcome_labels)=my_outcomes

all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='secondhand smoke exposure model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('european', 'hispanic', 'african', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))
```

# Supplemental Figure XX: Surrogate variable associations with known covariates

```{r}
#bring the heatmap function
covar=c('Sample_Plate', 'Slide', 'Array', 'cm1bsex', 'smkPreg_binary', 'childteen', 'ancestry')

bring.the.heatmap.sv <- function(pd,covar,folder){

  nsv <- length(grep('sv',names(pd)))

  #generate correlation matrix between SVs and variables
  correlations <- matrix(0,nrow=length(covar),ncol=nsv)
  colnames(correlations) <- paste('sv',1:nsv,sep='')
  rownames(correlations) <- covar

  #take svs
  svs <- pd[,colnames(correlations)]

  #correlations between svs and variables:
  # cor test if continuous
  # anova is multi category
  # t-test if two category
  for(col in covar){
    var <- pd[,col]
    #check to see if it is numeric and isn't really a two category thing
    if(class(var)=='numeric' & dim(table(var))>2){
      correlations[col,] <- apply(svs,2,FUN = function(x,y) cor.test(x,y)$p.value,y=var)
    }else{
      if(length(levels(factor(var)))>2){
        correlations[col,] <- apply(svs,2,FUN = function(x,y) anova(lm(x~y))$'Pr(>F)'[1],y=var)
      }else{
        correlations[col,] <- apply(svs,2,FUN = function(x,y) t.test(x~y)$p.value,y=var)
      }
    }
  }
  color.gradient <- colorRampPalette(c('firebrick','ivory3','royalblue'),bias=3)
  correlations=correlations%>%as.data.frame()%>%tibble::rownames_to_column('var1')%>%
    pivot_longer(-var1, names_to='var2')%>%
    mutate(var2=factor(var2, levels=unique(mixedsort(var2))),
           var1=factor(var1, levels=pd[, covar]%>%colnames(), labels=get_label(pd[, covar])))
  correlations%>%
    ggplot(aes(x=var1, y=var2, fill=value))+geom_tile()+
    scale_fill_gradientn(colours = color.gradient(20), values = seq(0, 1, 0.5))+
    scale_x_discrete(labels=str_wrap(get_label(pd[, covar]), 10))
}

bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
```

# Supplemental Table XX: Surrogate variable models

```{r}
test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_main_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='C'~'Age 9 models', 
                             childteen=='T'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age_svs=tbl_stack(table_list, group_header=names(table_list))

test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_main_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry_svs=tbl_stack(table_list, group_header=names(table_list))
```

# Figure 4 - Polymethylation scores more accurately delineate prenatal smoke exposure status than non-targeted DNA methylation summary measures and *a priori* CpGs

```{r, fig.width=10}

age_labels=c('Age 9', 'Age 15'); names(age_labels)=c('C', 'T')
methyl_labels_roc=paste0('Base model+', all_methyl_labels)
names(methyl_labels_roc)=paste0('Base model+', names(all_methyl_labels))

fig4a= global_roc%>%
    dplyr::select(predictors, covariates, methylation, childteen, test, modeltype)%>%
    unnest(test) %>% 
    filter(str_detect(methylation, 'SSnewbornCT|cg0557|global') | modeltype=='Base model')%>%
    ggplot(aes(x=spec, y=sens, color=modeltype))+geom_line()+
    facet_wrap(childteen~., nrow=1, labeller=labeller(childteen=age_labels))+
    theme_classic()+
    scale_color_manual('', values=c('red', 'darkred', 'darkblue', 'darkgreen', 'blue', 'green'))+
    theme(legend.position = 'right')+guides(color=guide_legend(nrow=2, byrow=TRUE))+
    geom_abline(intercept=1, slope=-1, color='black')+
    xlab('Specificity')+ylab('Sensitivity')+
    theme(legend.position='bottom')+ coord_fixed()

fig4b=global_roc%>%
  dplyr::select(covariates, methylation, childteen, test)%>%unnest(test) %>% 
  filter(str_detect(methylation, 'SSnewbornCT'))%>%
  rbind(long_global_roc%>% 
          mutate(childteen='C+T', methylation=gsub('_C.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>%
          unnest(test) %>% 
          filter(str_detect(methylation, 'SSnewbornCT') 
                 & covariates!=''))%>%
  ggplot(aes(x=spec, y=sens, color=childteen, linetype=childteen))+geom_line(size=2)+
  facet_wrap(~methylation, labeller = labeller(methylation=c('SSnewbornCT_center'='Polymethylation score, \n newborn mean centered')))+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=2,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  scale_color_grey('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  scale_linetype('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

cowplot::plot_grid(fig4a, fig4b, labels=c('A', 'B'), nrow=1, rel_widths=c(0.65, 0.35), align='h', axis='t')

```

```{r}
roc_table=global_roc %>% dplyr::select(childteen, methylation, modeltype, roc, auc)

roc_base=roc_table %>% filter(modeltype=='Base model+No methylation')%>%mutate(roc_basemodel=roc)%>%dplyr::select(childteen, roc_basemodel)
roc_ss=roc_table %>% filter(modeltype=='Sustained smoking w/ cell type correction (newborn)')%>%mutate(roc_pms=roc)%>%dplyr::select(childteen, roc_pms)

roc_table1=roc_table%>%filter(!str_detect(modeltype, 'Base model'))%>%left_join(roc_ss)%>%mutate(roctest=map2(roc, roc_pms, ~roc.test(.x, .y)))
roc_table1%>%mutate(`pval comparison to sustained smoking score (newborns)`=map_dbl(roctest, ~.x$p.value))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to sustained smoking score (newborns)`)

roc_table2=roc_table%>%filter(str_detect(modeltype, 'Base model'))%>%left_join(roc_base)%>%mutate(roctest=map2(roc, roc_basemodel, ~roc.test(.x, .y)))
roc_table2%>%mutate(`pval comparison to base model`=map_dbl(roctest, ~.x$p.value))%>%dplyr::select(childteen, modeltype, auc, `pval comparison to base model`)
```

# Question for group: add EWAS results to this paper

- Arguments in favor: recent paper in Clinical Epigenetics included both 
- Arguments against: enough here for a compellling story; deadline of December 1st; submit and if asked for we have it available.

# Cross sectional EWAS results:

# Longitudinal EWAS results

# METAL meta-analysis of 450K + EPIC chip data