### 2021-10-28 Papers/Polymethylation_scores/98_declarations.Rmd
added declarations

# Declarations

## Ethics approval and consent to participate

## Consent for publication

Not applicable

## Availability of data and materials

## Competing interests

## Funding

## Authors' contributions

## Acknowledgements