### 2021-11-02 Papers/Polymethylation_scores/01_titlepage.Rmd
responses to kelly edits november2

# Title Page

**Title:** Polymethylation Scores for Prenatal Maternal Smoke Exposure Persist Until Age 15 and Are Detected in Saliva

**Authors:** Freida A. Blostein^1, Jonah Fisher^2, John Dou^1, Lisa Schenper^3, Erin Ware^2 Daniel A. Notterman^3, Colter Mitchell^2, Kelly M. Bakulski^1

**Affiliations:**

^1 Department of Epidemiology, University of Michigan School of Public Health, Ann Arbor, Michigan, United States of America

^2 Institute for Social Research, University of Michigan, Ann Arbor, Michigan, United States of America

^3 Department of Molecular Biology, Princeton University, Princeton, New Jersey, United States of America