### 2021-11-02 Papers/Polymethylation_scores/06_discussion.Rmd

---
output:
  word_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
---

# Discussion

In the longitudinal Fragile Families and Child Wellbeing birth cohort, we observed prenatal maternal smoking was associated with DNA methylation in children's saliva samples from ages 9 and 15. Prenatal maternal smoking was associated with polymethylation scores across strata of child age and ancestry. Global methylation and epigenetic clocks were not associated with maternal smoking exposure demonstrating a specific and persistent signal of prenatal smoke exposure in DNA methylation. Polymethylation scores had reasonable accuracy for classifying prenatal maternal smoking. This accuracy was more consistent across time than when using individual *a priori* CpG sites.

Our findings are consistent with the previous literature on associations between prenatal maternal smoking and DNA methylation. We replicated top hits from previous epigenome wide association studies, including a large meta-analysis of DNA methylation in newborn blood samples [@Joubert:2016ue]. There are no previous studies of prenatal maternal smoking and saliva DNA methylation. However, previous work has that the majority of CpG sites were similarly methylated in blood and saliva (reviewed in @Langie2017). DNA methylation at AHRR:cg05575291 can accurately classify own--smoking behavior in both blood (area under the curve 0.995) and saliva (area under the curve 0.971) [@Dawes2019; @Philibert2020].

We advance the literature on prenatal maternal smoking and DNA methylation by evaluating the persistence of the association as children age and its portability across tissue and ancestry. Polymethylation scores built using coefficients from meta-analysis of cord blood DNA methylation from primarily European-ancestry newborns were still associated with prenatal maternal smoking in our independent saliva samples from a diverse cohort at ages 9 and 15 [@Joubert:2016ue]. The portability of other risk scores, such as polygenic risk scores, across ancestries is a complex research area [@Martin2019; ] and evaluating the portability of epigenetic summary measures has been identified as a key area for evaluation [@Hüls2020]. In this case, the polymethylation score for prenatal maternal smoking appears to be portable across ancestry groups. Effect estimates were consistent across ancestries at age 9, although Hispanic and European genetic ancestry samples had lower effect estimates at age 15. This could reflect higher unreported smoking initiation in European and Hispanic ancestry teens in the United States than African ancestry teens. Genetic ancestry correlates with race, and White and Latino teens have much higher rates of teen smoking and earlier ages at initiation than Black children [@El-Toukhy2016]. Though we excluded children who reported own-smoking from our analytic sample, under-reporting of own-smoking could influence DNA methylation at ages 9 and 15, creating outcome misclassification. This could also explain the observed attenuation the effect on AHRR:cg05575921 methylation by age 15 in our sample, as DNA methylation at this probe is known to vary by personal cigarette smoking[@Dawes2019; @Philibert2020].

Our findings are also consistent with previous work on DNA methylation as a biomarker for prenatal maternal smoke exposure. Prenatal maternal smoking classification using saliva DNA methylation in our sample of 9 and 15 year olds () performed comparably to classification using peripheral blood from older adults () [@Richmond:2018tu]. However, previous biomarkers of prenatal maternal smoking using DNA methylation of cord blood from newborns () and peripheral blood from 3-5 year old children () more accurately classified prenatal maternal smoke exposure [@Reese:2017ts; @Ladd-Acosta:2016aa]. Differences in DNA methylation patterns across tissues and over time may influence the performance of DNA methylation biomarkers. The similarity in performance between saliva DNA methylation in our study and peripheral blood DNA methylation in older adults is encouraging for the use of saliva DNA methylation as a readily accessible clinical sample.

Our analysis contributes to the development of an accurate methylation biomarker for prenatal maternal smoking by evaluating the impact of specific methodological choices on accuracy of classification. DNA methylation biomarkers may be especially susceptible to confounding by subject age at methylation measurement, tissue of measurement, and cell-type proportion [@Hüls2020]. Accuracy of polymethylation score classification was similar when using coefficients from cord blood vs peripheral blood samples from older children. Accuracy was not improved by using two time-points of methylation measurement. Accuracy was improved when polymethylation scores used coefficients which incorporated cell-type control. Our analysis therefore suggests that cell-type control when both generating and applying coefficients for polymethylation scores may positively influence accuracy, although more work comparing coefficients from different populations is still needed.

Our analysis suggests that polymethylation scores may be more accurate than using single CpG sites as biomarkers. AHRR:cg05575921 is a consistent marker of prenatal maternal smoke exposure in meta-analyses of newborn cord blood[@Joubert:2016ue] and a very accurate DNA methylation marker of own smoking [@Philibert2020]. However, salivary AHRR:cg05575921 was not a persistent marker of prenatal maternal smoking in an analysis of middle-aged adult women [@Richmond:2018tu]. In our sample, salivary AHRR:CpG05575921 methylation was less accurate at predicting prenatal maternal smoke exposure than polymethylation scores. The accuracy of AHRR:CpG05575921 and other single CpG site biomarkers may be influenced by time-since-exposure and new environmental exposures. Incorporating information across multiple sites of DNA methylation may yield a biomarker more robust to these influences.

Our analysis is not without its limitations. We used maternal self report of prenatal maternal smoker as serum cotinine levels were not available. Due to social desirability bias, this could result in exposure misclassification. We would expect this to bias our results towards the null. In any analysis of a prenatal exposure and postnatal outcome, there is the possibility of selection bias into the cohort due to live birth bias. Selection bias is also possible due to loss-to-follow-up between birth and age 9. Additionally, while we controlled for postnatal secondhand smoke exposure and excluded children who reported any own-smoking, we cannot exclude the possibility of residual confounding in this observational cohort.

However, our analysis also has several strengths. We analyzed samples from a large sample of diverse participants underrepresented in genetic and epigenetic research [@Sirugo2019; @Bentley2017]. While our exposure measurement of prenatal maternal smoking was self reported, it was assessed prospectively and preceded outcome measurements. We analyzed repeated measures of DNA methylation with reproducible array measures conducted in a single batch. We tested associations between prenatal maternal smoking and multiple DNA summary measures to evaluate the specificity of the polymethylation scores. In sensitivity analyses, we adjusted for other prenatal exposures and postnatal smoke exposure to examine the specificity of the biomarker to nature and timing of exposure.