### 2021-11-02 Papers/Polymethylation_scores/08_methods.Rmd

---
output:
  word_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
---

## Methods

### Cohort

The Fragile Families and Child Wellbeing Study is a birth cohort of nearly 5,000 children born in 20 cities in the United States between 1998 and 2000 [@Reichman:2001wv]. Participants were selected at delivery using a three-stage stratified random sample design which oversampled unmarried mothers by a ratio of 3:1 [@Reichman:2001wv]. Participants were excluded on the following criteria: those with parents who planned to place the child for adoption, those where the father was deceased, those who did not speak English or Spanish well enough to be interviewed, births where the mothers or babies were too ill to complete the interview, and those where the baby died before the interview could take place. Children were followed longitudinally with assessments at ages 1, 3, 5, 9 and 15; additional follow up is ongoing. Assessments included medical record extraction, biosample collection, in-home assessments, and surveys of the mother, father, primary caregiver, teacher and child. At ages nine and fifteen a saliva sample was taken from the child [@Reichman:2001wv]. A subsample of the Fragile Families cohort was selected for saliva DNA and DNA methylation processing. <!--#This subsampling was .... PUT SUBSAMPLE technique here -->

### Covariates and exposure measurement

Demographic and prenatal maternal substance use variables were derived from maternal self-report questionnaire data at baseline (child's birth).

Maternal covariates included maternal income to poverty ratio at baseline, prenatal smoking, prenatal maternal drug and alcohol use, and postnatal maternal or primary caregiver smoking. Maternal income to poverty ratio is a constructed variable of the ratio of total household income (as self-reported by the mother) to the official poverty thresholds designated by the United States Census Bureau for the year preceding the interview. At baseline, mother's answered categorical questions about their prenatal smoking, drug and alcohol use. For maternal prenatal smoking, mother's were asked

> During your pregnancy, how many cigarettes did you smoke? Did you smoke...:
>
> 2 or more packs a day
>
> 1 or more but less than 2 packs per day
>
> Less than 1 pack a day
>
> None

Few participants reported smoking a pack or more a day, thus we dichotomized to any vs no prenatal maternal smoking. For maternal prenatal drug and alcohol use the mothers were asked

> During your pregnancy how often did you use drugs/drink alcohol (respectively):
>
> Never
>
> Less than 1 time per month
>
> Several times per month
>
> Several times per week
>
> Every day

When the child was 1, 5, 9 and 15 years of age, primary caregivers responded to questions about maternal and in-home smoking. To encapsulate general early childhood smoke exposure, we created a binary variable for postnatal exposure at ages 1 or 5. To encapsulate recent postnatal smoke exposure, we used a categorical variable for packs per day (no smoking, less than one pack/day, one or more packs/day) in the month prior to the age 9 and 15 interview.

Child covariates included child sex, child report of personal cigarette smoking, and child genetic ancestry. Mothers reported sex (male/female) of their child at baseline. At ages 9 and 15, the child responded to interview questions about ever smoking a cigarette.

Child genetic ancestry was calculated. Principal components of child genetic ancestry were calculated from genetic data (CHIP?) from child saliva samples.

### DNA methylation measurement

Salivary samples from the children were collected at ages nine and fifteen using the Oragene•DNA sample collection kit (DNA Genotek Inc., Ontario). Saliva DNA was extracted manually using DNA Genotek's purification protocol using prepIT L2P. DNA was bisulfite treated and cleaned using the EZ DNA Methylation kit (Zymo Research, California). Samples were randomized and plated across slides by demographic characteristics. Saliva DNA methylation was measured using the Illumina HumanMethylation 450k BeadArray [@Sandoval2011] and imaged using the Illumina iScan system. All samples were run in a single batch to minimize technical variability.

DNA methylation image data were processed in R statistical software (3.5) using the minfi package [@Aryee2014]. The red and green image pairs (n=1811) were read into R and the minfi *preprocessNoob* function was used to normalize dye bias and apply background correction. Further quality control was applied using the ewastools packages [@Heiss2018]. We dropped samples with \>10% of sites have detection p-value \>0.01 (n=43), sex discordance between DNA methylation predicted sex and recorded sex (n=20), and abnormal sex chromosome intensity (n=3). CpG sites were removed if they had detection p-value \>0.01 in 5% of samples (n=26,830) or were identified as cross-reactive (n=27,782) (Figure 1) [@Chen2013]. Relative proportions of immune and epithelial cell types were estimated from DNA methylation measures using a childhood saliva reference panel [@Middleton2021].

We created polymethylation scores for prenatal maternal smoke exposure. From an independent meta-analysis of prenatal smoke exposure and newborn DNA methylation, we extracted the regression coefficients of 6,074 CpG sites associated with prenatal maternal smoking at a false discovery ratio corrected $P$ value \<0.05 [@Joubert:2016ue]. We mean-centered the DNA methylation beta values in our study, weighted them by the independent refression coefficience and took the sum. We calculated polymethylation scores using regression coefficients from 4 different regressions (sustained smoking exposure and DNA methylation in newborn cordblood with cell-type control, sustained smoking exposure and DNA methylation in newborn cordblood without cell-type control, sustained smoking exposure and DNA methylation in peripheral blood from older children, any smoking exposure and DNA methylation in newborn cordblood) [@Joubert:2016ue]. Global DNA methylation was calculated for each sample as the mean methylation value of each sample across the cleaned probe set. Pediatric epigenetic age was calculated for each sample using the coefficients and methods provided by the creators (see [<https://github.com/kobor-lab/Public-Scripts>](https://github.com/kobor-lab/Public-Scripts){.uri}) [@mcewen2019]. The GRIM age clock, including the smoking pack-years subscale, was calculated as previously described [@lu2019].

Single *a priori* CpG sites were selected from the previous large meta-analysis of prenatal maternal smoking and DNA methylation in children's cord and peripheral blood samples [@Joubert:2016ue].

Surrogate variables were calculated from DNA methylation data using the function sva from the R package sva version 3.38.

### Statistical analyses

#### Inclusion exclusion criteria

From the 1692 samples from 884 unique individuals with quality controlled DNA methylation data, we further excluded any samples missing data on: maternal prenatal smoking (0 samples), alcohol (4 samples) or other drug use (6 samples), maternal income to poverty ratio (0 samples), maternal postnatal smoking data ( samples). We also excluded samples with missing child sex (0 samples), child age (0 samples) or genetic data ( samples). Finally, if a child reported ever smoking a cigarette at age 9, we excluded their samples. Age 15 samples from children who reported ever smoking a cigarette at age 15 were also excluded.

#### Primary model

In our primary model we adjusted for maternal income to poverty ratio at baseline, child sex, child age, plate from DNA methylation processing, and estimated immune cell proportion estimated from DNA methylation. In nonstratified models, we adjusted for the first two components of genetic ancestry from principal component analysis. In ancestry-stratified models, the first two principal components from principal component analysis run within each ancestry strata were used. While child sex, child age, plate from DNA methylation processing and immune cell proportions are not confounders (as they cannot casually effect prenatal maternal smoke exposure), these variables can strongly effect DNA methylation and so were adjusted for as precision variables.

#### Sensitivity analyses

In sensitivity analyses we performed additional adjustments: 1) models adjusting for prenatal drug and alcohol use, 2) models adjusting for prenatal drug and alcohol use and postnatal maternal/primary caregiver smoking, 3) models adjusting for surrogate variables calculated from DNA methylation data.

Code to perform all analyses is available (www.github.com/bakulskilab)