### 2021-11-04 Papers/Polymethylation_scores/99_FigureCaptions.Rmd

---
output:
  html_document: default
  word_document: default
bibliography: Methylation_SmokingBib.bib
always_allow_html: true
---

# Figure 1 - Selection of samples from the Fragile Families and Child Wellbeing study into analytic subset. N represents the number of individuals at each step in the selection procedure, M represents the number of samples. Individuals with repeated measures can have more than one sample.

# Figure 2 - Differences in selected DNA methylation summary measures by self-report of prenatal maternal smoking among `r n_total` children in the Fragile Families and Child Wellbeing study (CT=cell-type control)

# Figure 3 - Prental maternal smoke exposure is consistently associated with polymethylation scores at ages 9 and 15 and is portable across genetic ancestry groups in a sample of `r n_total` children in the Fragile Families and Child Wellbeing study

# Figure 4 - A) Polymethylation scores accurately classify maternal report of prenatal smoke exposure status at ages 9 and 15 among a `r n_total` children in the Fragile Families and Child Wellbeing Study B) Inclusion of two-visits of DNA methylation data does not improve classification of maternal report of prenatal smoke exposure status among `r n_total` children in the Fragile Families and Child Wellbeing study