### 2021-11-04 Papers/Polymethylation_scores/index_paper.Rmd

---
title: "PAPER DRAFT"
author: "------------"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    df_print: paged
bibliography: Methylation_SmokingBib.bib
csl: springer-vancouver.csl
always_allow_html: true
---

```{r, echo=F, include=F, warning=F, message=F}
library(stringr)
results_run=str_detect(here::here(), 'nfs/turbo/bakulski1')
```

```{r, child = here::here("Papers", "Polymethylation_scores", "01_titlepage.Rmd")}
```

```{r, child = here::here("Papers", "Polymethylation_scores", "02_abstract.Rmd")}
```

```{r, child = here::here("Papers", "Polymethylation_scores", "03_keywords.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "04_background.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "05_results.Rmd"), eval=results_run}
```

```{r child = here::here("Papers", "Polymethylation_scores", "06_discussion.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "07_conclusions.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "08_methods.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "98_declarations.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "99_FigureCaptions.Rmd"), eval=results_run}
```

# References