### 2021-11-04 Reports/Paper1.Rmd

---
title: "Polymethylation scores results"
author: "F blostein"
date: "9/30/2021"
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: no
    theme: sandstone
  pdf_document: default
  word_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(ggplot2)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(stringr)
library(dplyr)
library(compareGroups)
library(kableExtra)
library(gtsummary)
library(gtools)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(purrr)
library(broom)
library(broom.mixed)
library(ggplot2)
library(multcomp)
library(ggpubr)
library(cowplot)
library(geepack)
library(lme4)
library(pROC)
library(lmerTest)
library(stringr)
library(sva)
library(gtsummary)
library(gt)
library(here)
library(DiagrammeRsvg)
library(xml2)
library(ggrepel)

include_ewas=F
stand_alone=T
```

```{r temp read n data}
basedir=here::here()
outdir=file.path(gsub('Code', 'Output', basedir), 'Figures')
datadir=gsub('Code', 'Data', basedir)
load(paste0(datadir, '/CreatedData/allPhenoData.Rdata'))
load(paste0(datadir, '/CreatedData/allmethyl.Rdata'))
load(paste0(datadir, '/CreatedData/completeCaseSVA.Rdata'))
load(paste0(datadir, '/CreatedData/methylation_summarymodels.Rdata'))
```

```{r plotting variables}
theme_set(theme_bw())
```

```{r}
modeldata=modeldata %>% mutate(childteen=factor(childteen, levels=c('Age 9', 'Age 15')))
set_label(modeldata$childteen)='Visit'

n_total=modeldata %>% dplyr::select(id)%>%unique()%>%nrow()
m_obs=modeldata%>%nrow()
n_twovisit=modeldata %>% group_by(id)%>%summarise(n=n())%>%filter(n==2)%>%nrow()

all_methyl=c("globalmethylation", "pediatric", "grim", 'DNAmPACKYRS', "Epithelial.cells_saliva", "Leukocytes_saliva", "SSnewborn_center", "SSnewbornCT_center", "SSolder_center", "cg05575921")
main_methyl=c("globalmethylation", "pediatric", "SSnewbornCT_center", "cg05575921")

all_methyl_labels=modeldata%>%dplyr::select(any_of(all_methyl))%>%get_label()
all_methyl_labels[str_detect(all_methyl_labels, 'Polymethylation', )]=paste0(rep('Score:', 3), c('Sustained smoking \n newborns', 'Sustained smoking \n newborns (w/ CT control)', 'Sustained smoking \n older children'))
names(all_methyl_labels)=all_methyl

all_methyl_levels=c("SSnewbornCT_center", "SSnewborn_center",  "SSolder_center",
                                       "pediatric" , "grim", 'DNAmPACKYRS',  "cg05575921", 
                                      "globalmethylation","Epithelial.cells_saliva", "Leukocytes_saliva")

main_methyl_labels=modeldata%>%dplyr::select(any_of(main_methyl))%>%get_label()

my_round=function(x, d){
  y=c()
  d_orig=d
  for(i in 1:length(x)){
    d=d_orig
    while(round(x[i], d)==0){
    d=d+1
    }
    y[i]=round(x[i], d)
  }
  return(y)
}
```

```{r figure counter}
main_fig=1
sup_fig=1
main_tab=1
sup_tab=1
```

# Figure `r main_fig` - Selection of samples from the Fragile Families and Child Wellbeing (FFCW) study into analytic subset. N represents the number of individuals at each step in the selection procedure, M represents the number of samples, as individuals with repeated measures can have more than one sample.

```{r, eval=T}
#function
N_M_count=function(filter_statement, statement, df){
  paste0('m=', df %>% filter(eval(parse(text = filter_statement)))%>%nrow(), 
         ' samples & ', 
         'n=', df %>% filter(eval(parse(text = filter_statement))) %>% dplyr::select(idnum) %>%unique()%>%nrow(), 
         ' individuals with', statement)
}
#Make Labels
n_fullFF<-paste0("Fragile Families cohort \n", "n=", FF_labeled %>% nrow())
n_methylQC<-paste0("Cohort with 450K methylation data \n", "n=", methylCohort$idnum%>%unique()%>%length(), " individuals with m=", methylCohort$MethID%>%length(), " samples")

n_QCloss<-paste0("Methylation data QC: ", 
                 N_M_count("probe_fail_10==T & childteen!='M'", " with >10% of probes failing dropped, \n ", pdqc_all),
                 N_M_count("sex_new!=predicted_sex & probe_fail_10==F & childteen!='M'", 
                           "  with discordant sex samples dropped, \n ", pdqc_all),
                 N_M_count("predicted_sex_outlier==T & sex_new==predicted_sex & probe_fail_10==F & childteen!='M'", 
                           ' with abnormal sex chromosome intensity dropped & \n ', pdqc_all),
                 length(reps_remove), " duplicate samples dropped")

n_methylFF<-paste0("Cohort with quality controlled 450K data \n","n=", myFF %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", myFF%>%nrow(), " samples")

n_missingmaternalprenataldata=paste0('Missing maternal data: \n', 
                                     N_M_count("is.na(smkPreg_binary)", " with missing prenatal maternal smoking \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & is.na(m1g2_YesNoPreg)", " with missing prenatal alcohol \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & is.na(m1g3_YesNoPreg)", 
                                               " with missing prenatal other drug use \n ", myFF),
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & is.na(cm1inpov)", 
                                               " with missing maternal income to poverty ratio data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny=='Missing'", " with missing postnatal maternal smoking data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth=='Missing'", 
                                               " with missing past month maternal/primary care giver smoking packs/day", myFF))

maternaldata=myFF %>% filter(!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) & PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth!='Missing')

n_withmaternaldata=paste0('m=', maternaldata %>% nrow(), ' samples & n=', maternaldata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with maternal covariate data')

n_missingchilddata=paste0('Missing child data: \n', 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data'", 
                                    " with missing ancestry/genetic data  \n ", maternaldata), 
                          N_M_count("!is.na(ancestry) & ancestry!='Missing PC data' & is.na(cm1bsex)", 
                                    " with missing child sex data \n ", maternaldata), 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data' & !is.na(cm1bsex) & is.na(ChildAgeComposite)", 
                                    " with missing child age at sample \n ", maternaldata))

childdata= maternaldata %>% filter(ancestry!='Missing PC data' & !is.na(cm1bsex) & !is.na(ChildAgeComposite))
n_withchilddata=paste0('m=', childdata %>% nrow(), ' samples & n=', childdata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with child covariate data')

n_Child<-paste0('n=', childdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata%>%filter(childteen=='C' & k5f1l=='2 no')%>% nrow(), 
                           " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', childdata %>% filter((k5f1l=="1 yes" | is.na(k5f1l)) & childteen=='C') %>% nrow(), 
                             " individuals smoking \n or missing child smoking data \n at 9 year visit")
n_Teen<-paste0('n=', childdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata %>%filter(childteen=='T' & k6d40=="2 No") %>% nrow(), 
                          " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', childdata %>% filter((k6d40=="1 Yes" | is.na(k6d40) | k5f1l=='1 yes') & childteen=='T') %>% nrow,
                            " individuals smoking \n or missing focal child smoking data \n at 15 year visit")

if(childdata %>% filter(((k5f1l=="2 no" | (is.na(k5f1l) & k6d40=='2 No')) & childteen=='C') | (childteen=='T' & k6d40=="2 No" & (k5f1l=='2 no' |  is.na(k5f1l))))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

```

```{r, fig.height=12}

graph1<-DiagrammeR::grViz("
   digraph graph2 {
   compound=true;
   graph [layout = dot]

   node [shape = rectangle, width=4]
   a [label = '@@1']
   b [label = '@@2']
   c [label = '@@3']
   d [label = '@@4']
   e [label = '@@5']
   f [label = '@@6']
   g [label = '@@7']
   h [label = '@@11']
   i [label = '@@12']

   #fake nodes for connecting to edges
   node [shape=point, width=0.01, height=0.01]
   a1
   b1
   c1
   d1
   f1 [group=g1]
   g1
   h1

  #exclusion nodes
  node [shape = rectangle, width=4]
  aa [label = '@@8']
  bb [label = '@@9']
  cc [label = '@@10']
  hh [label = '@@13']
  ii [label = '@@14']

  #draw connections
#  a->m
#  m->a1 [arrowhead=none]
  a->a1 [arrowhead=none]
  a1->b
  b->b1 [arrowhead=none]
  b1->c
  c->c1 [arrowhead=none]
  c1->d
  d->d1 [arrowhead=none]
  d1->e
  e->f; e->g
  f->g1 [arrowhead=none]; g->h1 [arrowhead=none]
  g1->h; h1->i

{rank=same; b1->aa [minlen=2]}   
{rank=same; c1->bb [minlen=2]} 
{rank=same; d1->cc [minlen=2]}
{rank=same; g1->hh [minlen=2]}
{rank=same; h1->ii [minlen=2]}

#invisible constraints to force ordering
#ff->i [style=invis]
   }

   [1]: n_fullFF
   [2]: n_methylQC
   [3]: n_methylFF
   [4]: n_withmaternaldata
   [5]: n_withchilddata
   [6]: n_Child
   [7]: n_Teen
   [8]: n_QCloss
   [9]: n_missingmaternalprenataldata
   [10]: n_missingchilddata
   [11]: n_ChildSmkexcluded
   [12]: n_TeenSmkexcluded
   [13]: n_ChildSmkexclusions
   [14]: n_TeenSmkexclusions
")

graph1 %>%
  export_svg() %>%
  read_xml() %>%
  write_xml(file.path(outdir, paste0("Figure", main_fig, ".svg")))

main_fig=main_fig+1            
```
```{r, eval=stand_alone}
graph1
```

# Supplemental Table `r sup_tab` - Descriptive statistics of excluded vs included samples with methylation data available from the Fragile and Families and Child Wellbeing study

```{r}
myFF$MethylData2<-ifelse(myFF$MethID %in% modeldata$MethID, 'Final analysis group', 'Not in final analysis group')
set_label(myFF$MethylData2)<-'Exclusions based on missing data'
myFF=left_join(myFF, methyldata%>%mutate(idnum=as.character(idnum)))
myFF=myFF%>%mutate(across(c(PostnatalMaternalSmokingAny, SmkAtVisitPastmonth), ~ifelse(.x=='Missing', 
                                                       NA, .x)))%>%
  mutate(ancestry=ifelse(ancestry=='Missing PC data', NA, ancestry))%>%copy_labels(myFF)

exclusion_tbl=myFF %>% mutate(childteen=case_when(childteen=='C'~'Age 9', childteen=='T'~'Age 15'))%>%
  dplyr::select(smkPreg_binary, m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,
         SmkAtVisitPastmonth, ancestry, cm1bsex, childteen, globalmethylation, pediatric, grim,
         Epithelial.cells_saliva, Leukocytes_saliva, SSnewbornCT_center, SSolder_center,
         cg05575921, MethylData2)%>%
  tbl_summary(., by=MethylData2, statistic=list(all_continuous()~"{mean} ({sd})"))%>%
  modify_header(stat_by = "**{level}**, M samples = {n}")%>%
  add_n(col_label='M samples')#%>%
  #add_p(test=list(all_continuous()~"t.test"))

sup_tab=sup_tab+1
```

```{r, eval=stand_alone}
exclusion_tbl
```

```{r wide data}
modeldata_wide = modeldata%>%
  mutate(childteen=gsub(' ', '_', childteen))%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'childteen', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary',  all_methyl)))%>%
  pivot_wider(names_from = childteen, values_from=any_of(all_methyl))

model_twovisit=modeldata_wide[complete.cases(modeldata_wide), ]

modeldata_people=modeldata%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary', 'PostnatalMaternalSmokingAny')))%>%unique()
```

# Supplemental Figure `r sup_fig` - Pearson correlation of methylation summary measures from an anlysis of M=`r m_obs` samples and N=`r n_total`children in the Fragile Families and Child Wellbeing cohort

```{r}
library(Hmisc)
library(corrplot)
res2<-modeldata%>%dplyr::select(any_of(all_methyl))%>%cor()
#labels
colnames(res2)=all_methyl_labels
rownames(res2)=all_methyl_labels

pdf(file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)
dev.off()

sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=12, fig.height=12}
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)

```

# Supplemental Figure `r sup_fig` - Correlation between age 9 and age 15 visits for selected DNA methylation summary measures among `r n_twovisit` children with age 9 and age 15 visit data in the Fragile Families and Child Wellbeing study

```{r}
age_long_methyl=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))

age_wide_methyl=age_long_methyl%>%
  pivot_wider(names_from=childteen, 
              values_from=`Methylation summary measure`)

#checks
if(age_wide_methyl%>%filter(Methyl_factor=='globalmethylation')%>%filter(!is.na(`Age 9`) & !is.na(`Age 15`))%>%nrow()!=n_twovisit){stop('check sample size two visit vs age_wide methyl')}

visit_cor=age_wide_methyl%>%
  ggplot(aes(x=`Age 9`, y=`Age 15`))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)

ggexport(visit_cor, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=12, fig.height=12}
visit_cor
```

# Supplemental Figure `r sup_fig`: Distributions of selected methylation summary measures at age 9 and age 15 among `r n_total` children in the Fragile Families and Child Wellbeing study

```{r}
dist_methyl=age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..", label.y.npc=0.87)+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  geom_point(alpha=0)+geom_line(alpha=0, aes(group=idnum))+geom_smooth(aes(group=1), method='lm', color='black')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)+
  theme(legend.position = 'bottom')

ggexport(dist_methyl, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
sup_fig=sup_fig+1

```

```{r, eval=stand_alone, fig.height=11}
dist_methyl
```

# Figure `r main_fig`: Differences in selected DNA methylation summary measures by prenatal maternal smoke exposure among `r n_total` US children in the Fragile Families and Child Wellbeing study

```{r change in scores time, fig.width=12}

#add line for mean ? 
outcome_labels=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Polymethylation score \n (newborn, mean centered)', 'Polymethylation score  \n (older children, mean centered')
names(outcome_labels)=c('globalmethylation', 'pediatric', 'cg05575921', 'SSnewbornCT_center','SSolder_center')

fig2=age_long_methyl%>%
    filter(Methyl_factor%in% main_methyl)%>%
    ggplot(aes(x=childteen, y=`Methylation summary measure`, color=smkPreg_binary))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_compare_means(label='..p.signif..', label.y.npc = 0.88)+
    facet_wrap(~Methyl_factor, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete('Visit', labels=c('Age 9', 'Age 15'))+
    scale_color_grey('', labels=c('Not prenatal maternal smoke exposed', 'Prenatal smoke exposed'))+
    theme(legend.position = 'bottom')

ggexport(fig2, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')))
main_fig=main_fig+1

```

```{r, eval=stand_alone}
fig2
```

# Table `r main_tab`: Bivariate associations between prenatal maternal smoking, selected DNA methylation summary measures and important covariates among a diverse sample of `r n_total` children in the Fragile Families and Child Wellbeing study

```{r bivariate table}
smoking_tbl=modeldata %>% 
  dplyr::select(smkPreg_binary, m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,
         SmkAtVisitPastmonth, ancestry, cm1bsex, globalmethylation, pediatric, grim, DNAmPACKYRS, Epithelial.cells_saliva,
         Leukocytes_saliva, SSnewbornCT_center, SSolder_center, cg05575921, childteen)%>%
  tbl_strata(strata=childteen, .tbl_fun=~.x %>%
               tbl_summary(by=smkPreg_binary, statistic=
                             list(all_continuous()~"{mean} ({sd})"))%>%
               add_n()%>%add_p(test=list(all_continuous()~"t.test")))

smoking_tbl=smoking_tbl%>%as_gt()%>%
  tab_row_group(label='Maternal', rows = 1:11)%>%
  tab_row_group(label='Child', rows=12:25)

main_tab=main_tab+1
```

```{r, eval=stand_alone}
smoking_tbl
```

# Figure `r main_fig` - Polymethylation scores are associated with prenatal maternal smoke exposure consistently across age and ancestry group in a sample of `r n_total` US children

```{r}
my_outcomes=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)')

outcome_labels=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921',  'Polymethylation score \n (newborn, mean centered)')

all_models=rbind(local_models, global_models%>%mutate(ancestry='Unstratified'))

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.25)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='base model')%>%mutate(outcome_f=outcome)

names(outcome_labels)=my_outcomes

forest_plot=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='base model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

ggexport(forest_plot, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')), width=15)
main_fig=main_fig+1
```

```{r, eval=stand_alone, fig.width=12}
forest_plot
```

# Supplemental Figure `r sup_fig`: Sensitivity analyses for controlling for different variable sets in multivariable analyses of prenatal maternal smoking and DNA methylation summary measures

```{r}
with(modeldata, table(ancestry, PostnatalMaternalSmokingAny, smkPreg_binary))%>%kbl()

forest_plot_compareModels=all_models%>%
    mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
    unnest(tidied)%>%
    filter(term=='smkPreg_binaryYes')%>%
    filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
    ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
    facet_grid(predictors~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
    geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
    theme_classic()+
    theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
    scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
    scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

ggexport(forest_plot_compareModels, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=15, height=15)
sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=14}
forest_plot_compareModels
```

# Supplemental Figure `r sup_fig` DNA methylation summary measures across prenatal and postnatal exposures

```{r}

my_comparisons <- list(c("No prenatal smoke exposed & No maternal smoking at age 1 and 5", 
                        "No prenatal smoke exposed & Maternal smoking at age 1 or age 5"), 
                        c("No prenatal smoke exposed & Maternal smoking at age 1 or age 5", 
                          "Yes prenatal smoke exposed & Maternal smoking at age 1 or age 5"), 
                      c("Yes prenatal smoke exposed & No maternal smoking at age 1 and 5", 
                        "Yes prenatal smoke exposed & Maternal smoking at age 1 or age 5"))

n_count=modeldata%>%mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% group_by(childteen, exposed)%>%summarise(n=n())

give.n <- function(x){
  return(c(y = min(x)-0.1, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

dna_prepost=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', 'PostnatalMaternalSmokingAny', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))%>%

  mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% 
  filter(Methyl_factor%in% main_methyl[3:4])%>%
  ggplot(aes(x=exposed, y=`Methylation summary measure`, color=exposed))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_summary(fun.data = give.n, geom = "text", fun.y = max,
                  position = position_dodge(width = 0.75))+
  stat_compare_means(comparisons=my_comparisons)+
    facet_grid(Methyl_factor~childteen, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete(labels=function(x) str_wrap(x, width = 10))+
    theme(legend.position = 'right')+
  guides(color=guide_legend(nrow=4,byrow=TRUE))

ggexport(dna_prepost, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=10, height=10)
sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=14, fig.height=10}
dna_prepost
```

# Supplemental Figure `r sup_fig`: Surrogate variable associations with known covariates

```{r}
#bring the heatmap function
covar=c('Sample_Plate', 'Slide', 'Array', 'cm1bsex', 'smkPreg_binary', 'childteen', 'ancestry')

bring.the.heatmap.sv <- function(pd,covar,folder){

  nsv <- length(grep('sv',names(pd)))

  #generate correlation matrix between SVs and variables
  correlations <- matrix(0,nrow=length(covar),ncol=nsv)
  colnames(correlations) <- paste('sv',1:nsv,sep='')
  rownames(correlations) <- covar

  #take svs
  svs <- pd[,colnames(correlations)]

  #correlations between svs and variables:
  # cor test if continuous
  # anova is multi category
  # t-test if two category
  for(col in covar){
    var <- pd[,col]
    #check to see if it is numeric and isn't really a two category thing
    if(class(var)=='numeric' & dim(table(var))>2){
      correlations[col,] <- apply(svs,2,FUN = function(x,y) cor.test(x,y)$p.value,y=var)
    }else{
      if(length(levels(factor(var)))>2){
        correlations[col,] <- apply(svs,2,FUN = function(x,y) anova(lm(x~y))$'Pr(>F)'[1],y=var)
      }else{
        correlations[col,] <- apply(svs,2,FUN = function(x,y) t.test(x~y)$p.value,y=var)
      }
    }
  }
  color.gradient <- colorRampPalette(c('firebrick','ivory3','royalblue'),bias=3)
  correlations=correlations%>%as.data.frame()%>%tibble::rownames_to_column('var1')%>%
    pivot_longer(-var1, names_to='var2')%>%
    mutate(var2=factor(var2, levels=unique(mixedsort(var2))),
           var1=factor(var1, levels=pd[, covar]%>%colnames(), labels=get_label(pd[, covar])))
  correlations%>%
    ggplot(aes(x=var1, y=var2, fill=value))+geom_tile()+
    scale_fill_gradientn('P value', colours = color.gradient(20), values = seq(0, 1, 0.5))+
    scale_x_discrete('', labels=str_wrap(get_label(pd[, covar]), 10))+
    scale_y_discrete('')
}

pdf(file=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=11, height=13)
bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
dev.off()
sup_fig=sup_fig+1
```

```{r, eval=stand_alone, fig.width=12, fig.height=8}
bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
```

# Supplemental Table `r sup_tab`: Coefficients for multivariable models of prenatal maternal smoke exposure regressed on selected DNA methylation summary variables (outcomes) among children in the Fragile Families and Child Wellbeing study using surrogate variables

```{r}
test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age_svs=tbl_stack(table_list, group_header=names(table_list))

test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry_svs=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1
```

```{r, eval=stand_alone}
regression_table_ancestry_svs
```

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable linear models stratified by age or in multivariable linear mixed effect models among `r n_total` US children

```{r}
#my_main_outcomes=c('AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)', 'Sustained smoking polymethylation score (older children)', 'Pediatric clock', 'Global methylation')

test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1
```

```{r, eval=stand_alone}
regression_table_age
```

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable models among `r n_total` US children by ancestry

```{r}
test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1

```

```{r, eval=stand_alone}
regression_table_ancestry
```

# Supplemental Figure `r sup_fig`: Associations between prenatal maternal smoking and polymethylation scores calculated using alternative coefficient sets among `n n_total` children in the Fragile Families and Child Wellbeing study

```{r} 
 compare_pms=all_models%>%
    mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
    unnest(tidied)%>%
    filter(term=='smkPreg_binaryYes')%>%
    filter(str_detect(outcome, 'smoking'))%>%
  filter(ancestry=='Unstratified')%>%
 ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=outcome))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
    facet_grid(predictors~ancestry, scales='free_x')+
    geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
    theme_classic()

ggexport(compare_pms, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1

```

```{r, eval=stand_alone, fig.width=10}
compare_pms
```

# Supplemental Figure `r sup_fig` - Associations between prenatal maternal smoking and other *a priori* CpG sites selected from previous meta-analyses among `n n_total` children in the Fragile Families and Child Wellbeing study

```{r}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)

#read in Joubert meta top cpgs 
#read in cpgs from PMC4833289
cpgs<-read.csv(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), stringsAsFactors = F, header=F, skip=2)
#make headers
header1<-scan(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), nlines=1, what=character(), sep=',')
header1[7:26]<-c(rep("SSnewborn", 5), rep("SSnewbornCT", 5), rep("SSolder", 5), rep("anynewborn", 5))
header2<-paste(header1, scan(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), nlines=1, skip=1, what=character(), sep=','), sep='_')
names(cpgs)<-header2

#four meta-EWASes: SS_newborn (sustained smoking in newborns); SS_newbornCT (sustained smoking in newborns controlling for cell type) SS_older (sustained smoking in older children); any_newborn (any smoking in newborns)
#pivot_longer to get single column of coefs and a column for the analysis type
#and group into a list by analysis type
cpgs<-cpgs %>% pivot_longer(cols = c(matches("_Coef|_SE|_P|_FDR|_Direction")), names_to = c("set", ".value"), names_pattern="(.+)_(.+)")%>%rename_with(~gsub('_', '', .x))%>%mutate(Gene=`Mapped Gene`, estimate=Coef, conf.low=estimate-1.96*SE, conf.high=estimate+1.96*SE, source='Joubert et al. 2016', std.error=SE, p.value=P, Chr=CHR) %>%mutate(source=paste0(source, ': ', set))%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)%>%filter(set %in% c('SSolder', 'SSnewbornCT'))

myCpGs=all_models%>%filter(str_detect(outcome, 'cg') & predictors=='base model')%>%unnest(tidied)%>%filter(term=='smkPreg_binaryYes')%>%
    filter(ancestry=='Unstratified')%>%mutate(CpG=gsub( '.*: ', '', outcome), Gene=gsub(':.*', '', outcome), source=paste0('Fragile Families:', age))

top_cpgs=myCpGs$CpG

cpg_all<-rbind(myCpGs%>%filter(CpG %in% top_cpgs), cpgs%>%filter(CpG %in% top_cpgs)%>%dplyr::select(-Chr), topCPGs%>%filter(CpG%in%top_cpgs)%>%mutate_at(c('p.value'), as.numeric))

#geom blank to give good axes limits
blank_data <- cpg_all %>%dplyr::select(Gene, CpG, estimate, std.error)%>%group_by(Gene)%>%mutate(min=min(estimate), max=max(as.numeric(std.error), na.rm=T))%>%dplyr::select(-estimate, -std.error)%>%unique()%>%mutate(estimate=min*-1+max, conf.high=estimate, conf.low=estimate)%>%dplyr::select(-max, -min)%>%arrange(Gene)%>%mutate(source='Wiklund et al. 2019')

blank_data2 <- cpg_all %>%dplyr::select(Gene, CpG, estimate, std.error)%>%group_by(Gene)%>%mutate(min=min(estimate))%>%dplyr::select(-estimate, -std.error)%>%unique()%>%mutate(estimate=min*-1+0.005, conf.high=estimate, conf.low=estimate)%>%dplyr::select(-min)%>%arrange(Gene)%>%mutate(source='Fragile Families: Age_9')

plot_aprioriCpG<-ggplot(cpg_all, aes(x=estimate, xmin=conf.low, xmax=conf.high, color=source, y=CpG))+geom_point(position=position_dodge(width=1))+geom_errorbarh(height=0.1, position=position_dodge(width=1))+scale_x_continuous()+geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+facet_wrap(.~Gene, scales='free', ncol=2)+theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=4,byrow=TRUE))+geom_blank(data = blank_data, aes(x = estimate))+scale_color_discrete('', labels=c('Fragile families saliva: age 15', 'Fragile families saliva: age 9', 'Fragile families: mixed effect model', 'Joubert et al. 2016 newborn cord blood', 'Joubert et al. 2016 peripheral blood older children', 'Wicklund et al. peripheral blood >16 years'))

ggexport(plot_aprioriCpG, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1
```
```{r, eval=stand_alone, fig.width=10}
plot_aprioriCpG
```

# Figure `r main_fig` - Polymethylation scores more accurately delineate prenatal smoke exposure status than non-targeted DNA methylation summary measures and *a priori* CpGs

```{r}

age_labels=c('Age 9', 'Age 15'); names(age_labels)=c('Age 9', 'Age 15')
methyl_labels_roc=paste0('Base model+', all_methyl_labels)
names(methyl_labels_roc)=paste0('Base model+', names(all_methyl_labels))

fig4a= global_roc%>%
    dplyr::select(predictors, covariates, methylation, childteen, test, modeltype)%>%
    unnest(test) %>% 
    filter(str_detect(methylation, 'SSnewbornCT|cg0557|global') | modeltype=='Base model')%>%
    ggplot(aes(x=spec, y=sens, color=modeltype))+geom_line()+
    facet_wrap(childteen~., nrow=1, labeller=labeller(childteen=age_labels))+
    theme_classic()+
    scale_color_manual('', values=c('red', 'darkred', 'darkblue', 'darkgreen', 'blue', 'green'), 
                       function(x) str_wrap(x, width = 30))+
    theme(legend.position = 'right')+guides(color=guide_legend(nrow=4, byrow=TRUE))+
    geom_abline(intercept=1, slope=-1, color='black')+
    xlab('Specificity')+ylab('Sensitivity')+
    theme(legend.position='bottom')+ coord_fixed()

fig4b=global_roc%>%
  dplyr::select(covariates, methylation, childteen, test)%>% 
  filter(str_detect(methylation, 'SSnewbornCT') & covariates!='')%>%unnest(test)%>%
  rbind(long_global_roc%>% 
          mutate(childteen='Age 9 & Age 15', methylation=gsub('_Age_9.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>% 
          filter(str_detect(methylation, 'SSnewbornCT') 
                 & covariates!='')%>%
          unnest(test))%>%
  ggplot(aes(x=spec, y=sens, color=childteen, linetype=childteen))+geom_line(size=2)+
  facet_wrap(~methylation, labeller = labeller(methylation=c('SSnewbornCT_center_scale'='Polymethylation score, \n newborn mean centered')))+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  scale_color_grey('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  scale_linetype('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

auc_fig=cowplot::plot_grid(fig4a, fig4b, labels=c('A', 'B'), nrow=1, rel_widths=c(0.65, 0.35), align='h', axis='tb')

ggexport(auc_fig, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')), width=9)
main_fig=main_fig+1
```

```{r, eval=stand_alone, fig.width=10}
auc_fig
```

# Supplemental Table: Area under the curve for models shown in main figure 4 for classifying prenatal maternal smoke exposure, with *P* value when comparing to base model only or polymethylation score

```{r}
roc_table=global_roc %>% dplyr::select(childteen, methylation, modeltype, roc, auc)
roc_base=roc_table %>% filter(modeltype=='Base model+No methylation')%>%mutate(roc_basemodel=roc)%>%dplyr::select(childteen, roc_basemodel)
roc_ss=roc_table %>% filter(modeltype=='Sustained smoking polymethylation score (newborns, cell-type controlled)')%>%mutate(roc_pms=roc)%>%dplyr::select(childteen, roc_pms)

rct1=roc_table%>% 
  filter(methylation %in% c('globalmethylation', 'cg05575921', 'SSnewbornCT_center_scale', ''))%>%
  left_join(roc_base)%>%
  mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
  left_join(roc_ss)%>%
  mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
  mutate(auc=round(auc, 2))%>%
  mutate(`P value vs polymethylation score`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
  mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
  dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score`)

```

```{r, eval=stand_alone}
rct1%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs polymethylation score'))
```

# Supplemental Table: Area under the curve when using different coefficients to construct polymethylation scores for predicting prenatal maternal smoke exposure

```{r} 
rct2=roc_table%>% 
    filter(str_detect(modeltype, 'smoking|No methylation'))%>%
    left_join(roc_base)%>%
    mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
    left_join(roc_ss)%>%
    mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
    mutate(auc=round(auc, 4))%>%
    mutate(`P value vs polymethylation score with cell-type control`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
    mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
    dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score with cell-type control`)
```

```{r, eval=stand_alone}
rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs polymethylation score with cell-type control'))
```