### 2021-11-09 Papers/Polymethylation_scores/99_FigureCaptions.Rmd
added figure captions

---
output:
  html_document: default
  word_document: default
bibliography: Methylation_SmokingBib.bib
always_allow_html: true
---

```{r, echo=F, include=FALSE}
print_figs=TRUE
```

#### Table 1: Bivariate associations between prenatal maternal smoking (no/yes) and selected DNA methylation summary measures and important covariates among a diverse sample of `r n_total` children in the Fragile Families and Child Wellbeing study

```{r,  eval=print_figs, include=T}
smoking_tbl
```

#### Figure 1 - Selection of samples from the Fragile Families and Child Wellbeing study into analytic subset. N represents the number of individuals at each step in the selection procedure, M represents the number of samples. Individuals with repeated measures can have more than one sample.

```{r, echo=F, eval=print_figs, include=T, fig.height=11, fig.width=7}
graph1
```

#### Figure 2 - Differences in selected DNA methylation summary measures by self-report of prenatal maternal smoking among `r n_total` children in the Fragile Families and Child Wellbeing study (CT=cell-type control)

```{r, echo=F, eval=print_figs, include=T, fig.height=5, fig.width=7}
fig2
```

#### Figure 3 - Prental maternal smoke exposure is consistently associated with polymethylation scores at ages 9 and 15 and is portable across genetic ancestry groups in a sample of `r n_total` children in the Fragile Families and Child Wellbeing study

```{r, echo=F, eval=print_figs, include=T, fig.width=12}
forest_plot
```

#### Figure 4 - A) Polymethylation scores accurately classify maternal report of prenatal smoke exposure status at ages 9 and 15 among a `r n_total` children in the Fragile Families and Child Wellbeing Study B) Inclusion of two-visits of DNA methylation data does not improve classification of maternal report of prenatal smoke exposure status among `r n_total` children in the Fragile Families and Child Wellbeing study

```{r, echo=F, eval=print_figs, include=T,  fig.width=12}
auc_fig
```