### 2021-11-10 Reports/Paper1.Rmd

---
title: "Supplemental Materials: Polymethylation Scores for Prenatal Maternal Smoke Exposure Persist Until Age 15 and Are Detected in Saliva"
author: "Freida A. Blostein^1, Jonah Fisher^2, John Dou^1, Lisa Schenper^3, Erin Ware^2 Daniel A. Notterman^3, Colter Mitchell^2, Kelly M. Bakulski^1"
date: "`9/30/2021`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: tango
    number_sections: no
    theme: sandstone
  word_document: default
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message=FALSE)

library(ggplot2)
library(kableExtra)
library(ggpubr)
library(tidyr)
library(stringr)
library(dplyr)
library(compareGroups)
library(kableExtra)
library(gtsummary)
library(gtools)
library(haven)
library(sjlabelled)
library(matchmaker)
library(EpiDISH)
library("PerformanceAnalytics")
library(DescTools)
library(purrr)
library(broom)
library(broom.mixed)
library(ggplot2)
library(multcomp)
library(ggpubr)
library(cowplot)
library(geepack)
library(lme4)
library(pROC)
library(lmerTest)
library(stringr)
library(sva)
library(gtsummary)
library(gt)
library(here)
library(DiagrammeRsvg)
library(xml2)
library(ggrepel)

as_supp=T
stand_alone=F
```

```{r temp read n data}
basedir=here::here()
outdir=file.path(gsub('Code', 'Output', basedir), 'Figures')
datadir=gsub('Code', 'Data', basedir)
load(paste0(datadir, '/CreatedData/allPhenoData.Rdata'))
load(paste0(datadir, '/CreatedData/allmethyl.Rdata'))
load(paste0(datadir, '/CreatedData/completeCaseSVA.Rdata'))
load(paste0(datadir, '/CreatedData/methylation_summarymodels.Rdata'))
```

```{r plotting variables}
theme_set(theme_bw())
set_label(modeldata$cm1bsex)='Child gender'
set_label(modeldata$cm1inpov)='Maternal income to poverty threshold (at baseline)'
set_label(modeldata$PostnatalMaternalSmokingDose)='Primary caregiver smoking past month by dose'
set_label(modeldata$SSnewbornCT_center)='Polymethylation score for smoke exposure'
set_label(modeldata$pediatric)='Pediatric epigenetic clock (years)'
set_label(modeldata$grim)='GRIM epigenetic clock (years)'
set_label(modeldata$globalmethylation)='Percent global DNA methylation'
set_label(modeldata$cg05575921)='AHRR gene: percent cg05575921 methylation'
```

```{r}
modeldata=modeldata %>% mutate(childteen=factor(childteen, levels=c('Age 9', 'Age 15')))%>%droplevels()
set_label(modeldata$childteen)='Visit'

n_total=modeldata %>% dplyr::select(id)%>%unique()%>%nrow()
m_obs=modeldata%>%nrow()
n_twovisit=modeldata %>% group_by(id)%>%summarise(n=n())%>%filter(n==2)%>%nrow()

all_methyl=c("globalmethylation", "pediatric", "grim", 'DNAmPACKYRS', "Epithelial.cells_saliva", "Leukocytes_saliva", "SSnewborn_center", "SSnewbornCT_center", "SSolder_center", "cg05575921")
main_methyl=c("globalmethylation", "pediatric", "SSnewbornCT_center", "cg05575921")

all_methyl_labels=modeldata%>%dplyr::select(any_of(all_methyl))%>%get_label()
all_methyl_labels[str_detect(all_methyl_labels, 'Polymethylation', )]=paste0(rep('Score:', 3), c('Sustained smoking \n newborns', 'Sustained smoking \n newborns (w/ CT control)', 'Sustained smoking \n older children'))
names(all_methyl_labels)=all_methyl

all_methyl_levels=c("SSnewbornCT_center", "SSnewborn_center",  "SSolder_center",
                                       "pediatric" , "grim", 'DNAmPACKYRS',  "cg05575921", 
                                      "globalmethylation","Epithelial.cells_saliva", "Leukocytes_saliva")

main_methyl_labels=modeldata%>%dplyr::select(any_of(main_methyl))%>%get_label()

my_round=function(x, d){
  y=c()
  d_orig=d
  for(i in 1:length(x)){
    d=d_orig
    while(round(x[i], d)==0){
    d=d+1
    }
    y[i]=round(x[i], d)
  }
  return(y)
}
```

```{r figure counter}
main_fig=1
sup_fig=1
main_tab=1
sup_tab=1
```

```{r figure 1, eval=T}
#function
N_M_count=function(filter_statement, statement, df){
  paste0('m=', df %>% filter(eval(parse(text = filter_statement)))%>%nrow(), 
         ' samples & ', 
         'n=', df %>% filter(eval(parse(text = filter_statement))) %>% dplyr::select(idnum) %>%unique()%>%nrow(), 
         ' individuals with', statement)
}
#Make Labels
n_fullFF<-paste0("Fragile Families cohort \n", "n=", FF_labeled %>% nrow())
n_methylQC<-paste0("Cohort with 450K methylation data \n", "n=", methylCohort$idnum%>%unique()%>%length(), " individuals with m=", methylCohort$MethID%>%length(), " samples")

n_QCloss<-paste0("Methylation data QC: ", 
                 N_M_count("probe_fail_10==T & childteen!='M'", " with >10% of probes failing dropped, \n ", pdqc_all),
                 N_M_count("sex_new!=predicted_sex & probe_fail_10==F & childteen!='M'", 
                           "  with discordant sex samples dropped, \n ", pdqc_all),
                 N_M_count("predicted_sex_outlier==T & sex_new==predicted_sex & probe_fail_10==F & childteen!='M'", 
                           ' with abnormal sex chromosome intensity dropped & \n ', pdqc_all),
                 length(reps_remove), " duplicate samples dropped")

n_methylFF<-paste0("Cohort with quality controlled 450K data \n","n=", myFF %>% dplyr::select(id) %>% unique() %>% nrow(), " individuals with m=", myFF%>%nrow(), " samples")

n_missingmaternalprenataldata=paste0('Missing maternal data: \n', 
                                     N_M_count("is.na(smkPreg_binary)", " with missing prenatal maternal smoking \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & is.na(m1g2_YesNoPreg)", " with missing prenatal alcohol \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & is.na(m1g3_YesNoPreg)", 
                                               " with missing prenatal other drug use \n ", myFF),
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & is.na(cm1inpov)", 
                                               " with missing maternal income to poverty ratio data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny=='Missing'", " with missing postnatal maternal smoking data \n ", myFF), 
                                     N_M_count("!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) &
                                               PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth=='Missing'", 
                                               " with missing past month maternal/primary care giver smoking packs/day", myFF))

maternaldata=myFF %>% filter(!is.na(smkPreg_binary) & !is.na(m1g2_YesNoPreg) & !is.na(m1g3_YesNoPreg) & !is.na(cm1inpov) & PostnatalMaternalSmokingAny!='Missing' & SmkAtVisitPastmonth!='Missing')

n_withmaternaldata=paste0('m=', maternaldata %>% nrow(), ' samples & n=', maternaldata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with maternal covariate data')

n_missingchilddata=paste0('Missing child data: \n', 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data'", 
                                    " with missing ancestry/genetic data  \n ", maternaldata), 
                          N_M_count("!is.na(ancestry) & ancestry!='Missing PC data' & is.na(cm1bsex)", 
                                    " with missing child sex data \n ", maternaldata), 
                          N_M_count("is.na(ancestry) | ancestry=='Missing PC data' & !is.na(cm1bsex) & is.na(ChildAgeComposite)", 
                                    " with missing child age at sample \n ", maternaldata))

childdata= maternaldata %>% filter(ancestry!='Missing PC data' & !is.na(cm1bsex) & !is.na(ChildAgeComposite))
n_withchilddata=paste0('m=', childdata %>% nrow(), ' samples & n=', childdata %>% dplyr::select(idnum)%>%unique()%>%nrow(), ' individuals with child covariate data')

n_Child<-paste0('n=', childdata %>%filter(childteen=='C')  %>% nrow(), " individuals with a 9 year visit")
n_ChildSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata%>%filter(childteen=='C' & k5f1l=='2 no')%>% nrow(), 
                           " individuals with a 9 year visit")
n_ChildSmkexclusions<-paste0('n=', childdata %>% filter((k5f1l=="1 yes" | is.na(k5f1l)) & childteen=='C') %>% nrow(), 
                             " individuals smoking \n or missing child smoking data \n at 9 year visit")
n_Teen<-paste0('n=', childdata %>%filter(childteen=='T')   %>% nrow(), " individuals with a 15 year visit")
n_TeenSmkexcluded<-paste0("Exclude children who smoke: \n",'n=', childdata %>%filter(childteen=='T' & k6d40=="2 No") %>% nrow(), 
                          " individuals with a 15 year visit")
n_TeenSmkexclusions<-paste0('n=', childdata %>% filter((k6d40=="1 Yes" | is.na(k6d40) | k5f1l=='1 yes') & childteen=='T') %>% nrow,
                            " individuals smoking \n or missing focal child smoking data \n at 15 year visit")

if(childdata %>% filter(((k5f1l=="2 no" | (is.na(k5f1l) & k6d40=='2 No')) & childteen=='C') | (childteen=='T' & k6d40=="2 No" & (k5f1l=='2 no' |  is.na(k5f1l))))%>%nrow()!=nrow(modeldata)){stop('filter N and completecase data set N do not match')}

```

```{r figure 1 make figure, fig.height=12}

graph1<-DiagrammeR::grViz("
   digraph graph2 {
   compound=true;
   graph [layout = dot]

   node [shape = rectangle, width=4]
   a [label = '@@1']
   b [label = '@@2']
   c [label = '@@3']
   d [label = '@@4']
   e [label = '@@5']
   f [label = '@@6']
   g [label = '@@7']
   h [label = '@@11']
   i [label = '@@12']

   #fake nodes for connecting to edges
   node [shape=point, width=0.01, height=0.01]
   a1
   b1
   c1
   d1
   f1 [group=g1]
   g1
   h1

  #exclusion nodes
  node [shape = rectangle, width=4]
  aa [label = '@@8']
  bb [label = '@@9']
  cc [label = '@@10']
  hh [label = '@@13']
  ii [label = '@@14']

  #draw connections
#  a->m
#  m->a1 [arrowhead=none]
  a->a1 [arrowhead=none]
  a1->b
  b->b1 [arrowhead=none]
  b1->c
  c->c1 [arrowhead=none]
  c1->d
  d->d1 [arrowhead=none]
  d1->e
  e->f; e->g
  f->g1 [arrowhead=none]; g->h1 [arrowhead=none]
  g1->h; h1->i

{rank=same; b1->aa [minlen=2]}   
{rank=same; c1->bb [minlen=2]} 
{rank=same; d1->cc [minlen=2]}
{rank=same; g1->hh [minlen=2]}
{rank=same; h1->ii [minlen=2]}

#invisible constraints to force ordering
#ff->i [style=invis]
   }

   [1]: n_fullFF
   [2]: n_methylQC
   [3]: n_methylFF
   [4]: n_withmaternaldata
   [5]: n_withchilddata
   [6]: n_Child
   [7]: n_Teen
   [8]: n_QCloss
   [9]: n_missingmaternalprenataldata
   [10]: n_missingchilddata
   [11]: n_ChildSmkexcluded
   [12]: n_TeenSmkexcluded
   [13]: n_ChildSmkexclusions
   [14]: n_TeenSmkexclusions
")

graph1 %>%
  export_svg() %>%
  read_xml() %>%
  write_xml(file.path(outdir, paste0("Figure", main_fig, ".svg")))

main_fig=main_fig+1            
```

```{r figure 1 plot figure, eval=stand_alone}
graph1
```

# Supplemental Table `r sup_tab` - Descriptive statistics of excluded vs included samples with methylation data available from the Fragile and Families and Child Wellbeing study. Samples excluded according to the flow diagram in Main Figure 1

```{r supplemental table exclusions}
myFF$MethylData2<-ifelse(myFF$MethID %in% modeldata$MethID, 'Final analysis group', 'Not in final analysis group')
set_label(myFF$MethylData2)<-'Exclusions based on missing data'
myFF=left_join(myFF, methyldata%>%mutate(idnum=as.character(idnum)))
myFF=myFF%>%mutate(across(c(PostnatalMaternalSmokingAny, SmkAtVisitPastmonth), ~ifelse(.x=='Missing', 
                                                       NA, .x)))%>%
  mutate(ancestry=ifelse(ancestry=='Missing PC data', NA, ancestry))%>%copy_labels(myFF)

exclusion_tbl=myFF %>% mutate(childteen=case_when(childteen=='C'~'Age 9', childteen=='T'~'Age 15'))%>%
    dplyr::select(smkPreg_binary, childteen, MethylData2, SSnewbornCT_center, cg05575921, pediatric, globalmethylation, Leukocytes_saliva, Epithelial.cells_saliva, ancestry, cm1bsex,  m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,SmkAtVisitPastmonth )%>%
  mutate(smkPreg_binary=factor(smkPreg_binary, levels=c('No', 'Yes'), labels=c('Not exposed', 'Smoke exposed')), 
         cm1bsex=factor(cm1bsex, levels=c('1 Boy', '2 Girl'), labels=c('Boy', 'Girl')))%>%
  mutate(cm1bsex=set_label(cm1bsex, 'Child gender'), smkPreg_binary=set_label(smkPreg_binary, 'Prenatal maternal smoking'),
         childteen=set_label(childteen, 'Age at DNA methylation sample'))%>%
  tbl_summary(., by=MethylData2, statistic=list(all_continuous()~"{mean} ({sd})"))%>%
  modify_header(stat_by = "**{level}**, M samples = {n}")%>%
  add_n(col_label='M samples')#%>%
  #add_p(test=list(all_continuous()~"t.test"))

sup_tab=sup_tab+1
```

```{r exclusion table include, eval=as_supp}
exclusion_tbl
```

```{r wide data}
modeldata_wide = modeldata%>%
  mutate(childteen=gsub(' ', '_', childteen))%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'childteen', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary',  all_methyl)))%>%
  pivot_wider(names_from = childteen, values_from=any_of(all_methyl))

model_twovisit=modeldata_wide[complete.cases(modeldata_wide), ]

modeldata_people=modeldata%>%
  dplyr::select(any_of(c('idnum', 'ancestry', 'cm1bsex', 'ancestry', 
                         'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'cm1inpov',
                         'smkPreg_binary', 'PostnatalMaternalSmokingAny')))%>%unique()
```

# Supplemental Figure `r sup_fig` - Pearson correlation of selected DNA methylation summary measures from an anlysis of M=`r m_obs` samples and N=`r n_total`children in the Fragile Families and Child Wellbeing cohort. Lower diaganol shows the Pearson rho correlation coefficient, upper triangle shows ellipses illustrating relationship between variables colored by the magnitude and direction of the correlation.

```{r make supplemental figure correlation plot}
library(Hmisc)
library(corrplot)
res2<-modeldata%>%dplyr::select(any_of(all_methyl))%>%cor()
#labels
colnames(res2)=all_methyl_labels
rownames(res2)=all_methyl_labels

pdf(file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)
dev.off()

sup_fig=sup_fig+1
```

```{r supplemental correlation plot plot it , eval=as_supp, fig.width=12, fig.height=12}
corrplot.mixed(res2, upper="ellipse", lower='number', order='hclust', tl.pos ="lt", tl.cex=0.8)

```

# Supplemental Figure `r sup_fig` - Correlation between age 9 and age 15 visit DNA methylation summary measures among `r n_twovisit` children with age 9 and age 15 visit data in the Fragile Families and Child Wellbeing study. Each point represents one individual, x-axis is the DNA methylation summary measure at age 9 and y-axis is the DNA methylation measure at age 15. Blue lines are simple linear regression lines of best fit, with Pearson rho and associated $P$ value shown. From top-left: Polymethylation score for prenatal maternal smoking, score created using coefficients from a regression of prenatal maternal smoking and DNA methylation in newborn cord blood, incorporating cell-type control. Polymethylation score for prenatal maternal smoking using coefficients from a regression of prenatal maternal smoking and DNA methylation in newborn cord blood, no cell-type control. Polymethylation score calculated using regression coefficients from a regression of prenatal maternal smoking and DNA methylation in older children's peripheral blood samples. Pediatric epigenetic clock (years). GRIM epigenetic clock (years). Percent DNA methylation at cg05575921 in AHRR. Percent global DNA methylation (mean across all DNA methylation sites in cleaned probe set). Proportion of epithelial cells calculated from DNA methylation using reference panel of children's saliva. Proportion of immune cells calculated from DNA methylation using a reference panel of children's saliva.

```{r supplemental figure time correlation plot}
age_long_methyl=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))

age_wide_methyl=age_long_methyl%>%
  pivot_wider(names_from=childteen, 
              values_from=`Methylation summary measure`)

#checks
if(age_wide_methyl%>%filter(Methyl_factor=='globalmethylation')%>%filter(!is.na(`Age 9`) & !is.na(`Age 15`))%>%nrow()!=n_twovisit){stop('check sample size two visit vs age_wide methyl')}

visit_cor=age_wide_methyl%>%
  ggplot(aes(x=`Age 9`, y=`Age 15`))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+
  scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)

ggexport(visit_cor, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
sup_fig=sup_fig+1
```

```{r supplemenetal figure time correlation plot, eval=as_supp, fig.width=12, fig.height=12}
visit_cor
```

# Supplemental Figure `r sup_fig`: Distributions of selected methylation summary measures at age 9 and age 15 among `r n_total` children in the Fragile Families and Child Wellbeing study. Captive box plots within violin plots show the distribution of selected DNA methylation measures in samples from age 9 (red) and age 15 (blue). Black lines show lines of best fit. From top-left: Polymethylation score for prenatal maternal smoking, score created using coefficients from a regression of prenatal maternal smoking and DNA methylation in newborn cord blood, incorporating cell-type control. Polymethylation score for prenatal maternal smoking using coefficients from a regression of prenatal maternal smoking and DNA methylation in newborn cord blood, no cell-type control. Polymethylation score calculated using regression coefficients from a regression of prenatal maternal smoking and DNA methylation in older children's peripheral blood samples. Pediatric epigenetic clock (years). GRIM epigenetic clock (years). Percent DNA methylation at cg05575921 in AHRR. Percent global DNA methylation (mean across all DNA methylation sites in cleaned probe set). Proportion of epithelial cells calculated from DNA methylation using reference panel of children's saliva. Proportion of immune cells calculated from DNA methylation using a reference panel of children's saliva.

```{r supplemental figure methylation time distributions}
dist_methyl=age_long_methyl%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, color=childteen))+
  geom_violin()+geom_boxplot(width=0.2)+stat_compare_means(label="..p.signif..", label.y.npc=0.87)+
  scale_color_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  scale_x_discrete(name='Visit', labels=c('Age 9', 'Age 15'))+
  geom_point(alpha=0)+geom_line(alpha=0, aes(group=idnum))+geom_smooth(aes(group=1), method='lm', color='black')+
  facet_wrap(~Methyl_factor, scales='free', labeller = labeller(Methyl_factor=all_methyl_labels), ncol=3)+
  theme(legend.position = 'bottom')

ggexport(dist_methyl, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')))
sup_fig=sup_fig+1

```

```{r supplemental figure methylation time distribution, eval=as_supp, fig.height=11}
dist_methyl
```

```{r main figure 2 change in scores by smoking status, fig.width=12}

#add line for mean ? 
outcome_labels=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Polymethylation score \n (newborn, mean centered)', 'Polymethylation score  \n (older children, mean centered')
names(outcome_labels)=c('globalmethylation', 'pediatric', 'cg05575921', 'SSnewbornCT_center','SSolder_center')

fig2=age_long_methyl%>%
    filter(Methyl_factor%in% main_methyl)%>%
    ggplot(aes(x=childteen, y=`Methylation summary measure`, color=smkPreg_binary))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_compare_means(label='..p.signif..', label.y.npc = 0.88)+
    facet_wrap(~Methyl_factor, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete('Visit', labels=c('Age 9', 'Age 15'))+
    scale_color_grey('', labels=c('Not prenatal maternal smoke exposed', 'Prenatal smoke exposed'))+
    theme(legend.position = 'bottom')

ggexport(fig2, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')))
main_fig=main_fig+1

```

```{r main figure 2 change in scores smoking status, eval=stand_alone}
fig2
```

```{r bivariate table}

smoking_tbl=modeldata %>% 
  dplyr::select(smkPreg_binary, childteen, SSnewbornCT_center, cg05575921, pediatric, globalmethylation, Leukocytes_saliva, Epithelial.cells_saliva, ancestry, cm1bsex,  m1g2_YesNoPreg, m1g3_YesNoPreg, cm1inpov, PostnatalMaternalSmokingAny,SmkAtVisitPastmonth )%>%
  mutate(smkPreg_binary=factor(smkPreg_binary, levels=c('No', 'Yes'), labels=c('Not exposed', 'Smoke exposed')), 
         cm1bsex=factor(cm1bsex, levels=c('1 Boy', '2 Girl', labels=c('Boy', 'Girl'))))%>%
  tbl_strata(strata=childteen, .tbl_fun=~.x %>%
               tbl_summary(by=smkPreg_binary, statistic=
                             list(all_continuous()~"{mean} ({sd})"))%>%
               add_n()%>%add_p(test=list(all_continuous()~"t.test")))

smoking_tbl=smoking_tbl%>%as_gt()%>%
  tab_row_group(label='Child characteristics', rows = 1:13)%>%
  tab_row_group(label='Maternal characteristics', rows=14:23)%>%
     row_group_order(
      groups = c("Child characteristics", "Maternal characteristics")
    )

main_tab=main_tab+1
```

```{r, eval=stand_alone}
smoking_tbl
```

# Supplemental Table `r sup_tab`:  Bivariate associations between prenatal maternal smoking and additional DNA methylation summary measures among a diverse sample of `r n_total` children in the Fragile Families and Child Wellbeing study. Percent DNA methylation shown calculated across all DNA methylation sites available (global DNA methylation) and stratified by genomic region (Island, Open Sea, Shelf, Shore). Polymethylation score for smoke exposure from main analysis and from sensitivity analyses where scores were calculated using alternative available regression coefficients.

```{r supplemental table additional methylation measures}
smoking_tbl_extendedMeth=modeldata %>% 
  dplyr::select(smkPreg_binary, globalmethylation, Island_mean, OpenSea_mean, Shelf_mean, Shore_mean, 
                pediatric, grim, DNAmPACKYRS, Epithelial.cells_saliva, Leukocytes_saliva,
                SSnewbornCT_center, SSnewborn_center, SSolder_center, anynewborn_center,
                matches('cg'),  childteen)%>%
    mutate(smkPreg_binary=factor(smkPreg_binary, levels=c('No', 'Yes'), labels=c('Not exposed', 'Smoke exposed')))%>%
  tbl_strata(strata=childteen, .tbl_fun=~.x %>%
               tbl_summary(by=smkPreg_binary, statistic=
                             list(all_continuous()~"{mean} ({sd})"))%>%
               add_n()%>%add_p(test=list(all_continuous()~"t.test")))

sup_tab=sup_tab+1

```

```{r supplemental table additional methylation measures print, eval=as_supp}
smoking_tbl_extendedMeth
```

```{r main figure 3 forest plot}
my_outcomes=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)')

outcome_labels=c('Global methylation', 'Pediatric clock', 'AHRR: cg05575921',  'Polymethylation score \n (newborn, mean centered)')

all_models=rbind(local_models, global_models%>%mutate(ancestry='Unstratified'))

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.5)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='base model')%>%mutate(outcome_f=outcome)

names(outcome_labels)=my_outcomes

forest_plot=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='base model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

ggexport(forest_plot, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')), width=15)
main_fig=main_fig+1
```

```{r main figure 3 print forest plot, eval=stand_alone, fig.width=12}
forest_plot
```

# Supplemental Figure `r sup_fig`: Sensitivity analyses for controlling for different variable sets in multivariable analyses of prenatal maternal smoking and DNA methylation summary measures. Base model: Child sex, maternal income to poverty ratio, immune cell proportion, plate, 1st two principal components from PCA on child genetic data, child age. Prenatal exposures model: base model + yes/no any prenatal maternal alcohol use, yes/no any prenatal other drug use. Secondhand smoking models: prenatal exposures model + yes/no any postnatal maternal/primary care giver smoking at age 1 or 5, pk/day in month prior to primary caregiver/maternal interview at age 9 or 15. Surrogate variable: including all surrogate variables

```{r supplemental figure all models forest}
#with(modeldata, table(ancestry, PostnatalMaternalSmokingAny, smkPreg_binary))%>%kbl()

forest_plot_compareModels=all_models%>%
    mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
    unnest(tidied)%>%
    filter(term=='smkPreg_binaryYes')%>%
    filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
    ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
    facet_grid(predictors~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
    geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
    theme_classic()+
    theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
    scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
    scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'), labels=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))

ggexport(forest_plot_compareModels, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=15, height=15)
sup_fig=sup_fig+1
```

```{r, eval=as_supp, fig.width=14, fig.height=12}
forest_plot_compareModels
```

# Supplemental Figure `r sup_fig` DNA methylation summary measures across combinations of exposure to prenatal and postnatal  maternal smoking. Sample size is shown at the bottom of violin plots with captured boxplots. POlymethylation score shown in top facets, percent DNA methylation at cg05575921 in AHRR gene shown in bottom facets, age 9 samples shown in left facets, age 15 shown in right facets

```{r supplemental figure pre post smoke exposure}

my_comparisons <- list(c("No prenatal smoke exposed & No maternal smoking at age 1 and 5", 
                        "No prenatal smoke exposed & Maternal smoking at age 1 or age 5"), 
                        c("No prenatal smoke exposed & Maternal smoking at age 1 or age 5", 
                          "Yes prenatal smoke exposed & Maternal smoking at age 1 or age 5"), 
                      c("Yes prenatal smoke exposed & No maternal smoking at age 1 and 5", 
                        "Yes prenatal smoke exposed & Maternal smoking at age 1 or age 5"))

n_count=modeldata%>%mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% group_by(childteen, exposed)%>%summarise(n=n())

give.n <- function(x){
  return(c(y = min(x)-0.1, label = length(x))) 
  # experiment with the multiplier to find the perfect position
}

dna_prepost=modeldata%>%
  dplyr::select(any_of(c('idnum', 'childteen', 'ancestry', 'smkPreg_binary', 'PostnatalMaternalSmokingAny', all_methyl)))%>%
  pivot_longer(cols=any_of(all_methyl),
               names_to='Methylation summary measure type',
               values_to='Methylation summary measure')%>%
  mutate(Methyl_factor=factor(`Methylation summary measure type`, 
                              levels=all_methyl_levels))%>%

  mutate(exposed=paste0(smkPreg_binary, ' prenatal smoke exposed & ', PostnatalMaternalSmokingAny))%>% 
  filter(Methyl_factor%in% main_methyl[3:4])%>%
  ggplot(aes(x=exposed, y=`Methylation summary measure`, color=exposed))+
    geom_violin()+geom_boxplot(width=0.2, position = position_dodge(0.9))+
  stat_summary(fun.data = give.n, geom = "text", fun.y = max,
                  position = position_dodge(width = 0.75))+
  stat_compare_means(comparisons=my_comparisons)+
    facet_grid(Methyl_factor~childteen, scales='free',labeller = labeller(Methyl_factor=all_methyl_labels))+
    scale_x_discrete(labels=function(x) str_wrap(x, width = 10))+
    theme(legend.position = 'right')+
  guides(color=guide_legend(nrow=4,byrow=TRUE))

ggexport(dna_prepost, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=10, height=10)
sup_fig=sup_fig+1
```

```{r supplemental figure pre post smoke exposure print, eval=as_supp, fig.width=14, fig.height=10}
dna_prepost
```

# Supplemental Figure `r sup_fig`: Heatmap of *P* values from Pearson correlations between known covariates and surrogate variables from surrogate variable models

```{r surrogate variables heatmap}
#bring the heatmap function
covar=c('Sample_Plate', 'Slide', 'Array', 'cm1bsex', 'smkPreg_binary', 'childteen', 'ancestry')
modeldata$Sample_Plate=set_label(modeldata$Sample_Plate, 'Batch')
modeldata$cm1bsex=set_label(modeldata$cm1bsex, 'Child gender')
modeldata$PostnatalMaternalSmokingAny=set_label(modeldata$PostnatalMaternalSmokingAny, 'Postnatal maternal smoking (age 1 & 5)')
modeldata$SmkAtVisitPastmonth=set_label(modeldata$SmkAtVisitPastmonth, 'Postnatal maternal smoking (month prior to DNA methylation sample')

bring.the.heatmap.sv <- function(pd,covar,folder){

  nsv <- length(grep('sv',names(pd)))

  #generate correlation matrix between SVs and variables
  correlations <- matrix(0,nrow=length(covar),ncol=nsv)
  colnames(correlations) <- paste('sv',1:nsv,sep='')
  rownames(correlations) <- covar

  #take svs
  svs <- pd[,colnames(correlations)]

  #correlations between svs and variables:
  # cor test if continuous
  # anova is multi category
  # t-test if two category
  for(col in covar){
    var <- pd[,col]
    #check to see if it is numeric and isn't really a two category thing
    if(class(var)=='numeric' & dim(table(var))>2){
      correlations[col,] <- apply(svs,2,FUN = function(x,y) cor.test(x,y)$p.value,y=var)
    }else{
      if(length(levels(factor(var)))>2){
        correlations[col,] <- apply(svs,2,FUN = function(x,y) anova(lm(x~y))$'Pr(>F)'[1],y=var)
      }else{
        correlations[col,] <- apply(svs,2,FUN = function(x,y) t.test(x~y)$p.value,y=var)
      }
    }
  }
  color.gradient <- colorRampPalette(c('firebrick','ivory3','royalblue'),bias=3)
  correlations=correlations%>%as.data.frame()%>%tibble::rownames_to_column('var1')%>%
    pivot_longer(-var1, names_to='var2')%>%
    mutate(var2=factor(var2, levels=unique(mixedsort(var2))),
           var1=factor(var1, levels=pd[, covar]%>%colnames(), labels=get_label(pd[, covar])))
  correlations%>%
    ggplot(aes(x=var1, y=var2, fill=value))+geom_tile()+
    scale_fill_gradientn('P value', colours = color.gradient(20), values = seq(0, 1, 0.5))+
    scale_x_discrete('', labels=str_wrap(get_label(pd[, covar]), 10))+
    scale_y_discrete('')
}

pdf(file=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=11, height=13)
bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
dev.off()
sup_fig=sup_fig+1
```

```{r surrogate variable heatmap plot, eval=as_supp, fig.width=12, fig.height=8}
bring.the.heatmap.sv(modeldata, covar=c('smkPreg_binary', 'Sample_Plate', 'Slide', 'Array', 'ChildAgeComposite', 'ancestry', 'cm1bsex', 'Leukocytes_saliva', 'Epithelial.cells_saliva', 'cm1inpov', 'm1g2_YesNoPreg', 'm1g3_YesNoPreg', 'PostnatalMaternalSmokingAny', 'SmkAtVisitPastmonth'))
```

# Supplemental Table `r sup_tab`: Coefficients for multivariable models of prenatal maternal smoke exposure regressed on selected DNA methylation summary variables (outcomes) among children in the Fragile Families and Child Wellbeing study using surrogate variables

```{r supplemental table surrogate variable table}
test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age_svs=tbl_stack(table_list, group_header=names(table_list))

test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='surrogate variable models')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry_svs=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1
```

```{r surrogate variable table print, eval=as_supp}
regression_table_ancestry_svs
```

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable linear models stratified by age or in multivariable linear mixed effect models among `r n_total` US children. All models controlled for:  first two principal components of child genetic ancestry (recalculated within ancestry strata for stratified models), child sex, maternal income-to-poverty ratio at birth, immune cell proportion estimated from methylation data, child age

```{r supplemental table age and linear mixed effect}
#my_main_outcomes=c('AHRR: cg05575921', 'Sustained smoking polymethylation score (newborns, cell-type controlled)', 'Sustained smoking polymethylation score (older children)', 'Pediatric clock', 'Global methylation')

test=rbind(global_models, longitudinal_global_models %>% mutate(childteen='Longitudinal'))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(childteen=factor(case_when(childteen=='Age 9'~'Age 9 models', 
                             childteen=='Age 15'~'Age 15 models', 
                             childteen=='Longitudinal'~'Longitudinal models'),levels=c('Age 9 models', 'Age 15 models', 'Longitudinal models')), 
         tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$childteen)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_age=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1
```

```{r supplemental table age and mixed effect print, eval=as_supp}
regression_table_age
```

# Supplemental Table `r sup_tab`: Associations between prenatal maternal smoking and DNA methylation summary measures in multivariable models among `r n_total` US children by ancestry using linear mixed effect models. All models controlled for:  first two principal components of child genetic ancestry (recalculated within ancestry strata for stratified models), child sex, maternal income-to-poverty ratio at birth, immune cell proportion estimated from methylation data, child age

```{r supplemental table by ancestry using linear mixed effect models}
test=rbind(longitudinal_global_models %>% mutate(ancestry='Unstratified'),
           longitudinal_local_models%>%dplyr::select(-data))%>%
  filter(predictors=='base model')%>%
  filter(outcome %in% my_outcomes)%>%
  mutate(tbl=map(model, ~tbl_regression(.x, include='smkPreg_binary', 
                                        show_single_row='smkPreg_binary', 
                                        label=list(smkPreg_binary~'Prenatal maternal smoke exposed'))%>%add_n()))

table_list=test %>%split(f=.$ancestry)%>%map(~tbl_merge(.x$tbl, 
                                                         tab_spanner = as.character(.x$outcome)))

regression_table_ancestry=tbl_stack(table_list, group_header=names(table_list))

sup_tab=sup_tab+1

```

```{r supplemental table by ancestry using lme print, eval=as_supp}
regression_table_ancestry
```

# Supplemental Figure `r sup_fig`: Associations between prenatal maternal smoking and polymethylation scores calculated using alternative coefficient sets among `n n_total` children in the Fragile Families and Child Wellbeing study. Base model: Child sex, maternal income to poverty ratio, immune cell proportion, plate, 1st two principal components from PCA on child genetic data, child age. Prenatal exposures model: base model + yes/no any prenatal maternal alcohol use, yes/no any prenatal other drug use. Secondhand smoking models: prenatal exposures model + yes/no any postnatal maternal/primary care giver smoking at age 1 or 5, pk/day in month prior to primary caregiver/maternal interview at age 9 or 15. Surrogate variable: including all surrogate variables

```{r supplemental figure compare across pms coefficients} 
 compare_pms=all_models%>%
    mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
    unnest(tidied)%>%
    filter(term=='smkPreg_binaryYes')%>%
    filter(str_detect(outcome, 'smoking'))%>%
  filter(ancestry=='Unstratified')%>%
 ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=outcome))+
    geom_point(position = position_dodge(width = 1))+
    geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
    facet_wrap(predictors~., scales='free_x')+
    geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
    theme_classic()+
  scale_color_discrete(name='')+scale_x_continuous("Estimate (95% CI)")+scale_y_discrete('')

ggexport(compare_pms, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1

```

```{r compare across pms print, eval=as_supp, fig.width=10, fig.height=12}
compare_pms
```

# Supplemental Figure `r sup_fig` - Associations between prenatal maternal smoking and other *a priori* CpG sites selected from previous meta-analyses among `r n_total` children in the Fragile Families and Child Wellbeing study

```{r supplemental figure compare to top cpgs}
topCPGs<-read.csv(paste0(datadir, "/TopCpGsWiklundEtAl2019.csv"))
topCPGs<-topCPGs %>% mutate(estimate=as.numeric(gsub("−\\s", "-", gsub(" .*", "", β..SE.))), std.error=as.numeric(sub("\\).*", "", sub(".*\\(", "", β..SE.))), p.value=P.value, source="Wiklund et al. 2019", conf.low=(estimate-1.96*std.error), conf.high=(estimate+1.96*std.error), set ='')%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)%>%mutate(estimate=estimate*100, conf.low=100*conf.low, conf.high=100*conf.high)

#read in Joubert meta top cpgs 
#read in cpgs from PMC4833289
cpgs<-read.csv(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), stringsAsFactors = F, header=F, skip=2)
#make headers
header1<-scan(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), nlines=1, what=character(), sep=',')
header1[7:26]<-c(rep("SSnewborn", 5), rep("SSnewbornCT", 5), rep("SSolder", 5), rep("anynewborn", 5))
header2<-paste(header1, scan(paste0(datadir, '/PMC4833289_SuppFile2CpGs.csv'), nlines=1, skip=1, what=character(), sep=','), sep='_')
names(cpgs)<-header2

#four meta-EWASes: SS_newborn (sustained smoking in newborns); SS_newbornCT (sustained smoking in newborns controlling for cell type) SS_older (sustained smoking in older children); any_newborn (any smoking in newborns)
#pivot_longer to get single column of coefs and a column for the analysis type
#and group into a list by analysis type
cpgs<-cpgs %>% pivot_longer(cols = c(matches("_Coef|_SE|_P|_FDR|_Direction")), names_to = c("set", ".value"), names_pattern="(.+)_(.+)")%>%rename_with(~gsub('_', '', .x))%>%mutate(Gene=`Mapped Gene`, estimate=Coef, conf.low=(estimate-1.96*SE), conf.high=(estimate+1.96*SE), source='Joubert et al. 2016', std.error=SE, p.value=P, Chr=CHR) %>%mutate(source=paste0(source, ': ', set))%>%dplyr::select(CpG, Chr, Position, Gene, estimate, std.error, p.value, conf.low, conf.high, source, set)%>%filter(set %in% c('SSolder', 'SSnewbornCT'))%>%mutate(estimate=estimate*100, conf.low=100*conf.low, conf.high=100*conf.high)

myCpGs=all_models%>%filter(str_detect(outcome, 'cg') & predictors=='base model')%>%unnest(tidied)%>%filter(term=='smkPreg_binaryYes')%>%
    filter(ancestry=='Unstratified')%>%mutate(CpG=gsub( '.*: ', '', outcome), Gene=gsub(':.*', '', outcome), source=paste0('Fragile Families:', age))

top_cpgs=myCpGs$CpG

cpg_all<-rbind(myCpGs%>%filter(CpG %in% top_cpgs), cpgs%>%filter(CpG %in% top_cpgs)%>%dplyr::select(-Chr), topCPGs%>%filter(CpG%in%top_cpgs)%>%mutate_at(c('p.value'), as.numeric))

#geom blank to give good axes limits
blank_data <- cpg_all %>%dplyr::select(Gene, CpG, estimate, std.error)%>%group_by(Gene)%>%mutate(min=min(estimate), max=max(as.numeric(std.error), na.rm=T))%>%dplyr::select(-estimate, -std.error)%>%unique()%>%mutate(estimate=min*-1+max, conf.high=estimate, conf.low=estimate)%>%dplyr::select(-max, -min)%>%arrange(Gene)%>%mutate(source='Wiklund et al. 2019')

blank_data2 <- cpg_all %>%dplyr::select(Gene, CpG, estimate, std.error)%>%group_by(Gene)%>%mutate(min=min(estimate))%>%dplyr::select(-estimate, -std.error)%>%unique()%>%mutate(estimate=min*-1, conf.high=estimate, conf.low=estimate)%>%dplyr::select(-min)%>%arrange(Gene)%>%mutate(source='Fragile Families: Age_9')

plot_aprioriCpG<-ggplot(cpg_all, aes(x=estimate, xmin=conf.low, xmax=conf.high, color=source, y=CpG))+geom_point(position=position_dodge(width=1))+geom_errorbarh(height=0.1, position=position_dodge(width=1))+scale_x_continuous()+geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+facet_wrap(.~Gene, scales='free', ncol=2)+theme_classic()+theme(legend.position = 'bottom')+guides(color=guide_legend(nrow=4,byrow=TRUE))+geom_blank(data = blank_data, aes(x = estimate))+scale_color_discrete('', labels=c('Fragile families saliva: age 15', 'Fragile families saliva: age 9', 'Fragile families: mixed effect model', 'Joubert et al. 2016 newborn cord blood', 'Joubert et al. 2016 peripheral blood older children', 'Wicklund et al. peripheral blood >16 years'))+scale_x_continuous('Estimate (95% CI)')

ggexport(plot_aprioriCpG, filename=file.path(outdir, paste0('SupplementalFigure', sup_fig, '.pdf')), width=9)
sup_fig=sup_fig+1
```

```{r compare to top cpgs print, eval=as_supp, fig.width=10}
plot_aprioriCpG
```

```{r main figure 4 auc}

age_labels=c('Age 9', 'Age 15'); names(age_labels)=c('Age 9', 'Age 15')
methyl_labels_roc=paste0('Base model+', all_methyl_labels)
names(methyl_labels_roc)=paste0('Base model+', names(all_methyl_labels))

fig4a= global_roc%>%
    dplyr::select(predictors, covariates, methylation, childteen, test, modeltype)%>%
    unnest(test) %>% 
    filter(str_detect(methylation, 'SSnewbornCT|cg0557|global') | modeltype=='Base model')%>%
    ggplot(aes(x=spec, y=sens, color=modeltype))+geom_line()+
    facet_wrap(childteen~., nrow=1, labeller=labeller(childteen=age_labels))+
    theme_classic()+
    scale_color_manual('', values=c('red', 'darkred', 'darkblue', 'darkgreen', 'blue', 'green'), 
                       function(x) str_wrap(x, width = 30))+
    theme(legend.position = 'right')+guides(color=guide_legend(nrow=4, byrow=TRUE))+
    geom_abline(intercept=1, slope=-1, color='black')+
    xlab('Specificity')+ylab('Sensitivity')+
    theme(legend.position='bottom')+ coord_fixed()

fig4b=global_roc%>%
  dplyr::select(covariates, methylation, childteen, test)%>% 
  filter(str_detect(methylation, 'SSnewbornCT') & covariates!='')%>%unnest(test)%>%
  rbind(long_global_roc%>% 
          mutate(childteen='Age 9 & Age 15', methylation=gsub('_Age_9.*', '', methylation))%>%
          dplyr::select(covariates, methylation, childteen, test)%>% 
          filter(str_detect(methylation, 'SSnewbornCT') 
                 & covariates!='')%>%
          unnest(test))%>%
  ggplot(aes(x=spec, y=sens, color=childteen, linetype=childteen))+geom_line(size=2)+
  facet_wrap(~methylation, labeller = labeller(methylation=c('SSnewbornCT_center_scale'='Polymethylation score')))+
  theme_classic()+theme(legend.position = 'right')+guides(color=guide_legend(nrow=4,byrow=TRUE))+
  geom_abline(intercept=1, slope=-1, color='black')+
  scale_color_grey('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  scale_linetype('', labels=c('Age 9 only', 'Age 9 & 15', 'Age 15 only'))+
  xlab('Specificity')+ylab('Sensitivity')+
  theme(legend.position='bottom')+ coord_fixed()

auc_fig=cowplot::plot_grid(fig4a, fig4b, labels=c('A', 'B'), nrow=1, rel_widths=c(0.63, 0.37), align='h', axis='tb')

ggexport(auc_fig, filename=file.path(outdir, paste0('Figure', main_fig, '.pdf')), width=9)
main_fig=main_fig+1
```

```{r main figure 4 auc print, eval=stand_alone, fig.width=12}
auc_fig
```

# Supplemental Table `r sup_tab`: Area under the curve for models shown in Main Figure 4 for classifying prenatal maternal smoke exposure, with *P* value when comparing to base model only or to model using polymethylation score + base model. Base model: Child sex, maternal income to poverty ratio, immune cell proportion, plate, 1st two principal components from PCA on child genetic data, child age.

```{r supplemental table auc}
roc_table=global_roc %>% dplyr::select(childteen, methylation, modeltype, roc, auc)
roc_base=roc_table %>% filter(modeltype=='Base model+No methylation')%>%mutate(roc_basemodel=roc)%>%dplyr::select(childteen, roc_basemodel)
roc_ss=roc_table %>% filter(modeltype=='Base model+Sustained smoking polymethylation score (newborns, cell-type controlled)')%>%mutate(roc_pms=roc)%>%dplyr::select(childteen, roc_pms)

rct1=roc_table%>% 
  filter(methylation %in% c('globalmethylation', 'cg05575921', 'SSnewbornCT_center_scale', ''))%>%
  left_join(roc_base)%>%
  mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
  left_join(roc_ss)%>%
  mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
  mutate(auc=round(auc, 2))%>%
  mutate(`P value vs polymethylation score`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
  mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
  dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score`)

sup_tab=sup_tab+1  
```

```{r supplemental table auc print, eval=as_supp}
rct1%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs base model + polymethylation score'))
```

# Supplemental Table `r sup_tab`: Area under the curve when using different coefficient sets to construct polymethylation scores for predicting prenatal maternal smoke exposure. With *P* value when comparing to base model only or to model using polymethylation score + base model. Base model: Child sex, maternal income to poverty ratio, immune cell proportion, plate, 1st two principal components from PCA on child genetic data, child age.

```{r supplemental table auc other coefficients} 
rct2=roc_table%>% 
    filter(str_detect(modeltype, 'smoking|No methylation'))%>%
    left_join(roc_base)%>%
    mutate(roctest_base=map2(roc, roc_basemodel, ~roc.test(.x, .y)))%>%
    left_join(roc_ss)%>%
    mutate(roctest_ss=map2(roc, roc_pms, ~roc.test(.x, .y)))%>%
    mutate(auc=round(auc, 4))%>%
    mutate(`P value vs polymethylation score with cell-type control`=map_dbl(roctest_ss, ~round(.x$p.value, 4)))%>%
    mutate(`P value vs base model`=map_dbl(roctest_base, ~round(.x$p.value, 4)))%>%
    dplyr::select(childteen, modeltype, auc, `P value vs base model`, `P value vs polymethylation score with cell-type control`)
```

```{r supplemental table auc other coefficients print, eval=as_supp}
rct2%>%kbl(col.names=c('Age', 'Model', 'AUC', 'P value vs base model', 'P value vs base model + polymethylation score with cell-type control'))
```