### 2021-11-29 Papers/Polymethylation_scores/07_conclusions.Rmd

# Conclusions

In a large, prospective study of diverse participants, we showed that DNA methylation in children's saliva had strong associations with and reasonable classification accuracy for prenatal maternal smoke exposure. Further, we demonstrated that polymethylation scores could be applied as a biomarker of prenatal maternal smoke exposure across genetic ancestry groups, an important consideration for the equitable biomarker development. The development and application of biomarkers for prenatal maternal smoke exposure has important implications for epidemiological research and clinical practice. Given the difficulty of measuring prenatal maternal smoke exposure, such a biomarker could allow for confounder control in research areas where such control is currently impossible. Prenatal maternal smoke exposure is prevalent and has negative health consequences, thus an exposure biomarker could be used to provide support and health interventions for children.