### 2022-02-14 Papers/Polymethylation_scores/04_background.Rmd
revise intro and discussion

---
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
latex_engine: xelatex
---

# Background

In 2018, 11% of pregnant women in the United States aged 15-44 reported any past-month cigarette use [@SAMHSA2018]. Children exposed to cigarette smoking in utero are more likely to have low-birth weight, negative neurodevelopmental outcomes and asthma [@Banderali:2015vu]. Measuring prenatal maternal smoking behavior is challenging. Maternal prenatal smoking is underreported due to stigma [@Kvalvik:2012tk; @Dietz:2011ua]. The gold-standard biomarker, serum cotinine levels, have a half-life of nine hours in pregnant women, and are rarely available outisde of birth cohorts[@Dempsey:2002wr]. Alternative biomarkers of prenatal maternal smoke exposure could improve research and clinical practice.

DNA methylation is a possible biomarker. Prenatal maternal smoking is associated with DNA methylation signatures in placental, cord blood, and peripheral blood samples [@Joubert:2016ue; @Everson:2021td; @Cardenas2019; @Richmond:2018tu; @Tehranifar:2018tb]. Cord blood and placental samples are rarely available outside of birth cohorts. While peripheral blood is a more realistic sample, saliva is easier to collect [] and increasingly used in large epidemiological cohorts []. Since DNA methylation drives cell differentiation and cell type proportions differ across tissue types, it is important to verify associations in new tissues. To our knowledge, no study has evaluated associations between maternal prenatal cigarette smoking and salivary DNA methylation.

Additionally, DNA methylation can vary by genetics [@Galanter2017; @Fraser2012; @Moen2013; @Rahmani2017]. Own-smoking DNA methylation signatures differ by genetic ancestry [@Barcelona2019; @Sun2013; @Philibert2013; @Philibert2012; @Liu2019]. Most work on prenatal maternal smoking and child DNA methylation has been conducted in European-ancestry cohorts [@Joubert:2016ue], with some notable exceptions [@Ladd-Acosta:2016aa; @Tehranifar:2018tb; @Zhang2018; @Gonseth2016; @Neophytou:2019ty]. Some CpG sites associated with prenatal maternal smoking in European-ancestry cohorts have been replicated in non-European-ancestry or multi-ethnic populations [@Tehranifar:2018tb; @Neophytou:2019ty]. Children of both Black and non-Black identifying mothers exhibited similar directions of association between prenatal maternal smoking and peripheral blood DNA methylation at 38 CpG sites [@Xu2021]. However, this analysis stratified by self-reported ethnicity rather than genetic ancestry. Very studies have evaluated the portability of DNA methylation signatures of prenatal maternal smoke exposure across genetic ancestry populations within a single cohort [@Tehranifar:2018tb; @Gonseth2016].

After association testing, DNA methylation-based biomarkers of prenatal maternal smoking can be assessed by evaluating their classification accuracy. Biomarkers range from single DNA methylation sites to multi-site summary measures, such as polymethylation scores [@Philibert2020 @Philibert2013; @Reese:2017ts; @Richmond:2018tu; @Ladd-Acosta:2016aa]. In a polymethylation score, CpG sites are weighted by their effect size from a previous, independent sample and then summed to a single score [@Chen2020; @Bakulski2021a]. DNA polymethylation scores for prenatal maternal smoke exposure have performed well (area under the curves (AUCs)\>0.8) in newborns and young children [@Reese:2017ts; @Ladd-Acosta:2016aa] and fairly in middle-aged adults (AUC of 0.72 (95% confidence interval 0.69, 0.76)) [@Richmond:2018tu]. DNA methylation signatures of own smoking change post-cessation [@keshawarz2021]. Associations with prenatal maternal smoking change at some CpG sites as children age [@Richmond2015]. Yet few studies have assessed the longitudinal persistence of DNA-methylation biomarker accuracy.

DNA methylation differences hold promise as a biomarker for prenatal maternal smoking. We must understand how and if the signal varies across tissues sample, genetic ancestry, and age. In the Fragile Families and Child Wellbeing study, a diverse longitudinal birth cohort, we aimed to assess salivary DNA methylation biomarkers for prenatal maternal cigarette smoking. We tested associations between prenatal maternal smoke exposure and saliva DNA methylation, including polymethylation scores for smoke exposure, individual a priori CpG sites, global DNA methylation, and epigenetic clocks, at ages 9 and 15. We then performed ancestry- and age-stratified analyses to test the hypothesis that DNA methylation could serve as a portable and persistent biomarker of prenatal maternal smoking.