### 2022-02-14 Papers/Polymethylation_scores/06_discussion.Rmd

---
output:
  word_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
---

# Discussion

In the longitudinal Fragile Families and Child Wellbeing birth cohort, we observed that prenatal maternal smoking was associated with several characterizations of DNA methylation in childrens saliva samples from ages 9 and 15. Prenatal maternal smoking was associated with polymethylation scores for prenatal smoke exposure across strata of genetic ancestry and age. Global methylation and epigenetic clocks were not associated with maternal smoking exposure. Polymethylation scores for prenatal smoke exposure improved classification of self-reported prenatal maternal smoking over a base model, while a single *a priori* CpG site, *AHRR*:cg05575921, did not.

Our findings are consistent with the previous literature on associations between prenatal maternal smoking and DNA methylation. We replicated top hits from previous epigenome wide association analyses in cord and peripheral blood [@Joubert:2016ue; @wiklund2019]. There are no previous studies of prenatal maternal smoking and saliva DNA methylation. However, similar populations of immune cells are present in both blood and saliva, and saliva samples in our study were primarily composed of immune cells. Additionally, previous work has shown that the majority of CpG sites are similarly methylated in blood and saliva (reviewed in @Langie2017). Our results suggest salivary DNA methylation can be used as an alternative to peripheral blood DNA methylation for biomarkers of prenatal maternal smoking.

We advance prenatal smoking - DNA methylation literature by evaluating the portability of association between prenatal maternal smoking and DNA methylation across ancestry groups. The portability of polygenic risk scores across genetic ancestries is a complex research area [@Martin2019] and evaluating the portability of epigenetic summary measures has been identified as a key area for evaluation [@Huls2020]. Our results suggest polymethylation scores for prenatal maternal smoking are portable across genetic ancestry groups. In the samples from age 15, admixed Latin and European ancestry samples had reduced effect estimates. Genetic ancestry correlates with race, and White and Latino teens have much higher rates of teen smoking and earlier ages at initiation than Black children [@El-Toukhy2016]. Though we excluded children who reported own-smoking from our analytic sample, differential initiation and reporting of own-smoking could result in more similar DNA methylation profiles of children exposed vs unexposed to prenatal maternal smoking.

We contribute to the understanding of the impact of time-since-exposure on the classification accuracy of DNA methylation biomarkers for prenatal maternal smoking. Our saliva-based classification of 9- and 15-year-olds (AUC=0.78 and 0.77 respectively) was less accurate than previous cord-blood- and peripheral-blood-based classifications in newborns (AUCs ranging from 0.82-0.97) [@Reese:2017ts] and 3-5-year-old children (AUC=0.87) [@Ladd-Acosta:2016aa]. The difference in classification accuracy could reflect lower proportion of immune cells of saliva as compared to blood. However, they could also reflect a diminishing capacity to distinguish exposure status as time since exposure increases and new environmental exposures are encountered. Indeed, our classification was better than peripheral-blood-based classification in 30-year-old adults (AUC=0.72) [@Richmond:2018tu]. Notably, a previous longitudinal analysis of children's blood samples from birth to 17 years of age found that larger changes in DNA methylation occured between 0 and 7 years of age than between 7 and 17 years of age [@Richmond2015]. In that cohort, the association between prenatal smoke exposure and DNA methylation changed between childhood and adolescence at some but not all CpG sites. This may explain why we found that the polymethylation score improved classification accuracy while *AHRR*:cgg05575921 methylation did not. While *AHRR*:cg05575921 hypomethylation is a consistent marker of prenatal maternal smoking in newborns and children, the association did not persist in samples from adult women [@Richmond:2018tu]. Similarly, while *AHRR*:cg05575291 methylation measured in both blood and saliva can very accurately classify own-smoking [@Dawes2019; @Philibert2020], the magnitude of hypomethylation is reduced post-cessation[@keshawarz2021]. As the accuracy of *AHRR*:cg05575921 and other single CpG site biomarkers may be influenced by time-since-exposure and new environmental exposures, incorporating information across multiple sites of DNA methylation may yield a a more robust biomarker.

We note that while the pediatric DNA methylation clock accurately recapitulated subjects chronological ages, the GRIM-age values in our population were overestimated by 10-15 years. In an analysis of 4 pediatric cohorts of children under 20 years of age, GRIM age estimates were 30-35 years greater than chronological in buccal cell samples and 10-15 years greater than chronological age in blood samples [@okazaki2021]. GRIM-age was neither trained nor validated in a pediatric population [@lu2019], which may explain this overestimation.

Our analysis is not without its limitations. We used maternal self-report of prenatal maternal smoke exposure, as serum cotinine levels were not available. Due to social desirability bias, this could result in exposure misclassification. We would expect this to bias our results towards the null. In any analysis of a prenatal exposure and postnatal outcome, there is the possibility of selection bias into the cohort due to live birth bias. Selection bias is also possible due to loss-to-follow-up between birth and age 15. Additionally, while we controlled for postnatal secondhand smoke exposure and excluded children who reported any own-smoking, we cannot exclude the possibility of residual confounding in this observational cohort.

However, our analysis also has several strengths. We analyzed samples from a large cohort of diverse participants underrepresented in genetic and epigenetic research [@Sirugo2019; @Bentley2017]. While our exposure measurement of prenatal maternal smoking was self-reported, it was assessed prospectively and preceded outcome measurements. We analyzed repeated measures of DNA methylation with reproducible array measures conducted in a single batch. We tested associations between prenatal maternal smoking and multiple DNA summary measures to evaluate the specificity of the polymethylation scores. We adjusted for prenatal maternal drinking, other prenatal drug use and postnatal smoke exposure to examine the specificity of the biomarker to nature and timing of exposure.