### 2022-03-18 Papers/Polymethylation_scores/01_titlepage.Rmd
post submission revisions

---
output:
  pdf_document: default
  word_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
always_allow_html: yes
---

# Title Page

**Title:** Polymethylation Scores for Prenatal Maternal Smoke Exposure Persist Until Age 15 and Are Detected in Saliva in the Fragile Families and Child Wellbeing Cohort

**Authors:** Freida A. Blostein$^1$, Jonah Fisher$^2$,  John Dou$^1$, Lisa Schenper$^3$, Erin Ware$^2$ Daniel A. Notterman$^{3*}$, Colter Mitchell$^{2*}$, Kelly M. Bakulski$^{1*+}$

**Affiliations:**

$^1$Department of Epidemiology, University of Michigan School of Public Health, Ann Arbor, Michigan, United States of America

$^2$Institute for Social Research, University of Michigan, Ann Arbor, Michigan, United States of America

$^3$Department of Molecular Biology, Princeton University, Princeton, New Jersey, United States of America

*Co-senior authors

$^+$Corresponding author