### 2022-03-18 Papers/Polymethylation_scores/04_background.Rmd

---
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
latex_engine: xelatex
csl: springer-vancouver.csl
---

# Introduction

In 2018, 11% of pregnant people in the United States aged 15-44 reported any past-month cigarette use [@SAMHSA2018]. Children exposed to cigarette smoking *in utero* are more likely to have low-birth weight, negative neurodevelopmental outcomes, and asthma [@Banderali:2015vu]. Measuring prenatal maternal smoking behavior can be challenging. Maternal prenatal smoking is underreported due to stigma [@Kvalvik:2012tk; @Dietz:2011ua]. The gold-standard smoking biomarker, serum cotinine levels, have a half-life of nine hours in pregnant women, and are rarely available outside of birth cohort research [@Dempsey:2002wr]. Alternative biomarkers of prenatal maternal smoke exposure with a larger window of measurement could improve research and clinical practice.

DNA methylation is a possible biomarker. Prenatal maternal smoking is associated with offspring DNA methylation signatures at birth in placental and cord blood samples, and postnatally in peripheral blood samples [@Joubert:2016ue; @Everson:2021td; @Cardenas2019; @Richmond:2018tu; @Tehranifar:2018tb]. Another postnatal biospecimen option is saliva, which is easier to collect than blood and increasingly used in large epidemiological cohorts [@Langie2017]. Since DNA methylation drives cell differentiation and cell type proportions differ across tissue types, it is important to verify associations in new tissues [@Bakulski2016]. In addition, most studies of prenatal maternal smoking and childhood DNA methylation have been conducted in primarily European genetic ancestry populations [@Joubert:2016ue; @Ladd-Acosta:2016aa; @Tehranifar:2018tb; @Zhang2018; @Gonseth2016; @Neophytou:2019ty], and it will be important to test generalizability to other groups. To our knowledge, few studies have assessed multiple genetic ancestry groups in the same cohort and no study has evaluated associations between maternal pregnancy cigarette smoking and salivary DNA methylation in children.

Using the findings from large-scale association testing, in independent populations, DNA methylation-based biomarkers of prenatal maternal smoking can be developed and assessed by evaluating their classification accuracy. Biomarkers can range from single DNA methylation sites to summary measures across multiple sites, termed polymethylation scores [@Philibert2020 @Philibert2013; @Reese:2017ts; @Richmond:2018tu; @Ladd-Acosta:2016aa]. In a polymethylation score, DNA methylation sites are weighted by their effect size from a previous, independent sample and then summed to a single score [@Chen2020; @Bakulski2021a]. DNA polymethylation scores for prenatal maternal smoke exposure have performed well (area under the curves (AUCs)\>0.8) in cord blood samples from newborns and peripheral blood samples from young children [@Reese:2017ts; @Ladd-Acosta:2016aa] and fairly well in peripheral blood from middle-aged adults (AUC of 0.72 (95% confidence interval 0.69, 0.76)) [@Richmond:2018tu]. Associations with prenatal maternal smoking change at some DNA methylation sites as children age [@Richmond2015]. Yet few studies have assessed the longitudinal persistence of DNA-methylation biomarker accuracy.

DNA methylation differences hold promise as a biomarker for prenatal maternal smoke exposure. We must understand how the signal varies across tissue, genetic ancestry, and age. In the Fragile Families and Child Wellbeing Study, a diverse longitudinal birth cohort, we aimed to assess salivary DNA methylation biomarkers for prenatal maternal cigarette smoking. We tested associations between prenatal maternal smoke exposure and saliva DNA methylation, including polymethylation scores for smoke exposure, individual a priori DNA methylation sites, global DNA methylation, and epigenetic clocks, at ages 9 and 15. We then performed genetic ancestry- and age-stratified analyses to test the hypothesis that DNA methylation could serve as a portable and persistent biomarker of prenatal maternal smoking.