### 2022-03-18 Papers/Polymethylation_scores/08_methods.Rmd

---
output:
  word_document: 
    reference_docx: "../Template.docx"
  html_document: default
bibliography: [Methylation_SmokingBib.bib, packages.bib]
---

## Methods

### Cohort

The Fragile Families and Child Wellbeing Study is a birth cohort of nearly 5,000 children born in 20 cities in the United States between 1998 and 2000 [@Reichman:2001wv]. Participants were selected at delivery using a three-stage stratified random sample design which oversampled unmarried mothers by a ratio of 3:1 [@Reichman:2001wv]. Participants were excluded at the baseline wave on the following criteria: the parents who planned to place the child for adoption, the father was deceased, the parents did not speak English or Spanish well enough to be interviewed, or the mother or baby were too ill to complete the interview, or the baby died before the interview could take place. Children were followed longitudinally with assessments at ages 1, 3, 5, 9 and 15; an age 22 follow up is ongoing. Across certain child ages, assessments included medical record extraction, biosample collection, in-home assessments, and surveys of the mother, father, primary caregiver, teacher and child. At ages nine and fifteen a saliva sample was taken from the child [@Reichman:2001wv]. A subsample of the Fragile Families cohort was selected for saliva DNA and DNA methylation processing [@Reichman:2001wv]. To be eligible for the DNA methylation assessment children had to have participated and given saliva at age 9 and 15; after which a random sample of these eligible participants was then selected. One exception to this rule is that all participants in the Study of Adolescent Neurodevelopment (SAND) were assayed even if they only provided a sample at one time-point. All analyses account for this oversample of these participants.

### Pre- and postnatal exposure measurement

Prenatal maternal substance use variables were derived from maternal self-report questionnaire data at baseline (child's birth).

At baseline, mother's answered categorical questions about their prenatal smoking, drug and alcohol use. For maternal prenatal smoking, mother's were asked: "During your pregnancy, how many cigarettes did you smoke?". Responses were categorical: 2 or more packs a day, 1 or more but less than 2 packs per day, less than 1 pack a day or none. Few participants reported smoking a pack or more a day, thus for our primary exposure variable we created a dichotomous any versus no prenatal smoke exposure variable.

In sensitivity analyses we controlled for other types of prenatal exposures. Mothers were also asked how frequently they drank during their pregnancy. We created a dichotomous any versus no prenatal alcohol exposure variable. Similarly, mothers were asked how frequently they used any other drug during their pregnancy. We again created a binary variable for any versus no prenatal use of other drugs.

In sensitivity analyses, we controlled for postnatal second hand smoke exposure. When the child was 1, 5, 9 and 15 years of age, primary caregivers responded to questions about smoking in the household. To encapsulate early life postnatal secondhand smoke exposure, we created a binary variable for postnatal exposure at ages 1 or 5. To encapsulate recent postnatal smoke exposure to the age 9 and 15 interview where DNA methylation was collected, we used a categorical variable for packs per day (no smoking, less than one pack/day, one or more packs/day) in the month prior to each interview.

To ensure we measured the direct effect of prenatal exposure to cigarette smoke, rather than the effect of personal postnatal smoking we excluded children who reported own smoking. At ages 9 children were asked if they had ever smoked a cigarette or used tobacco (yes/no) and at age 15 they were asked if they had ever smoked an entire cigarette (yes/no).

### Covariate measurement

We used household income-to-poverty ratio at baseline as a proxy for socioeconomic status. Income-to-poverty ratio is the ratio of total household income (as self-reported by the mother) to the official poverty thresholds designated by the United States Census Bureau for the year preceding the interview.

Mothers reported sex (male/female) of their child at baseline.

Child genetic ancestry was calculated. Grants R01 HD36916, R01 HD073352, and R01 HD076592 provided support for the collection, assay, and analysis of the genetic data in the cohort. Principal components (PC) of child genetic ancestry were calculated from genetic data using Illumina PsychChip_v1-1 with child saliva samples. Genetic ancestry was assigned by comparing PC loadings to 1000 Genomes super population clusters. Samples with PC1 \> 0.018 and PC2 \> -0.0075 were assigned to European ancestry. Samples with PC1 \< -0.005 and PC2 \> 0.007+0.75(PC1) were assigned to African ancestry. Samples with PC1 \> 0.018 and -0.055 \< PC2 \< 0.025 were assigned to Admixed ancestry [@FFPGS2021]. Samples assigned to Admixed ancestry were primarily from individuals whose mothers self-reported Hispanic ethnicity (`r modeldata%>%group_by(ancestry, cm1ethrace)%>%summarise(n=n())%>%ungroup()%>%group_by(ancestry)%>%mutate(sum=sum(n), percent=round(100*(n/sum), 0))%>%filter(ancestry=='Admixed ancestry - Latin heritage' & cm1ethrace=='3 hispanic')%>%pull(percent)`% of samples), so we label this group Admixed - Latin ancestry.

### DNA methylation measurement

Salivary samples from the children were collected at ages nine and fifteen using the Oragene•DNA sample collection kit (DNA Genotek Inc., Ontario). Saliva DNA was extracted manually using DNA Genotek's purification protocol using prepIT L2P. DNA was bisulfite treated and cleaned using the EZ DNA Methylation kit (Zymo Research, California). Samples were randomized and plated across slides by demographic characteristics. Saliva DNA methylation was measured using the Illumina HumanMethylation 450k BeadArray [@Sandoval2011] and imaged using the Illumina iScan system. All samples were run in a single batch to minimize technical variability.

DNA methylation image data were processed in R statistical software (3.5) using the minfi package [@Aryee2014]. We used the February 2022 data freeze of the Fragile Families and Child Wellbeing DNA methylation data. The red and green image pairs (n=1811) were read into R and the minfi *preprocessNoob* function was used to normalize dye bias and apply background correction. Further quality control was applied using the ewastools packages [@Heiss2018]. We dropped samples using the following criteria: if \>10% of DNA methylation sites had detection p-value \>0.01 (n=`r probe_fail`), if there was sex discordance between DNA methylation predicted sex and recorded sex (n=`r sex_mismatch`), if samples had outlier methylation values (n=`r meth_outlier`) or if two sequential samples from the same individual exhibited genetic discordance between visits (n=`r genetic_discordance`) (Figure 1). Technical replicates were removed (n=`r tech_rep`). DNA methylation sites were removed if they had detection p-value \>0.01 in 5% of samples (n=33376) [@Chen2013]. Relative proportions of immune and epithelial cell types were estimated from DNA methylation measures using a childhood saliva reference panel [@Middleton2021].

We investigated several summary measures of DNA methylation, including polymethylation scores for smoke exposure, single DNA methylation sites, mean global DNA methylation, and DNA methylation clocks.

Our primary outcome variable was a polymethylation score for prenatal maternal smoke exposure. From an independent meta-analysis of prenatal smoke exposure and DNA methylation (the Joubert meta-analysis), we extracted the regression coefficients of 6,074 DNA methylation sites associated with prenatal maternal smoking at a false discovery ratio corrected $P$ value \<0.05 [@Joubert:2016ue]. We mean-centered the DNA methylation beta values in our study, weighted them by the independent regression coefficients and took the sum. For our main analysis we used coefficients from a regression of sustained smoking exposure and DNA methylation in newborn cord blood with cell-type control.

To assess the robustness of polymethylation scores to the source of coefficients for weights, we generated two sensitivity polymethylation scores from the Joubert meta-analysis [@Joubert:2016ue]. First, we used the regression of 568 cord blood DNA methylation sites associated with prenatal maternal smoking at a strict Bonferroni threshold, when analyses were not adjusted for cell proportions [@Joubert:2016ue]. Second, we used the regression coefficients of 19 child peripheral blood DNA methylation sites associated with prenatal maternal smoking at the strict Bonferroni threshold when analyses were not adjusted for cell proportions [@Joubert:2016ue].

To assess the robustness of polymethylation scores to the selection of DNA methylation sites, we generated two additional sensitivity polymethylation scores. One score was generated using 28 DNA methylation sites and coefficients prioritized from a LASSO regression of prenatal maternal smoking and cord blood methylation in the Norwegian Mother and Child Cohort Study [@Reese:2017ts]. Another score was generated using 204 DNA methylation sites and coefficients prioritized from an elastic net regression of prenatal maternal smoking and DNA methylation in adolescent peripheral blood [@rauschert2020]. This score was calculated using the coefficients and code provided by the authors in the R code accompanying the manuscript [@R-DNAsmokeR].

To compare the utility of polymethylation scores relative to single site biomarkers, we selected the top DNA metyhlation site from the Joubert meta-analysis, *AHRR:* cg05575921. As a sensitivity analyses, we also selected four other *a priori* DNA methylation sites which were identified from meta-analyses of prenatal maternal smoking and DNA methylation in children's cord and peripheral blood samples: *GFI1*:cg14179389, *CYP1A1*: cg05549655, *MYO1G*: cg22132788, and *MYO1G*:cg04180046 [@Joubert:2016ue; @wiklund2019].

To assess the specificity of the impact of prenatal smoke exposure on DNA methylation, we compared the polymethylation scores to mean global and regulatory regional DNA methylation and methylation-derived age measures. Global DNA methylation was calculated for each sample as the mean methylation value of each sample across the cleaned probe set. Mean DNA methylation restricted to probes in genomic regions (CpG island, shore, shelf or open sea, as identified in the R package IlluminaHumanMethylation450kanno.ilmn12.hg19 v 0.6.0) was also calculated. Pediatric epigenetic age was calculated for each sample using the coefficients and methods provided by the creators (see [\<https://github.com/kobor-lab/Public-Scripts\>](https://github.com/kobor-lab/Public-Scripts){.uri}) [@mcewen2019]. As a sensitivity analysis, the GRIM-age clock, which was neither trained nor validated in a pediatric population, was calculated as previously described [@lu2019].

To control for unmeasured confounding and potential batch effects, we calculated surrogate variables from DNA methylation data using the function sva from the R package sva version 3.38 [@R-sva]. The relationship between prenatal maternal smoking and DNA methylation was protected.

### Statistical analyses

#### Inclusion exclusion criteria

We excluded samples missing any data on the following variables: maternal prenatal smoking, maternal prenatal alcohol use, maternal prenatal other drug use, maternal income to poverty ratio, postnatal secondhand smoke exposure variables, child sex, or child age. We also excluded samples that were missing genetic data. Age 9 samples from children who reported ever smoking a cigarette or using tobacco age 9 or were missing own-smoking data were excluded. Age 15 samples from children who reported smoking a cigarette at age 9 or at age 15 or were missing own-smoking data were excluded. Children who were missing a response to the question at age 9 but answered that they had never smoked a whole cigarette at age 15 were kept in the sample.

#### Descriptive statistics

We described the distribution of continuous variables using the mean and standard deviation. We described the distribution of categorical variables using percents. We compared the distribution of included to excluded samples using the Welch's t-test for continuous variables and the Chi-Square test for categorical variables. We report the correlation between continuous measures of DNA methylation (including polymethylation scores, single sites, mean global DNA methylation and estimated cell type proportions) using the Pearson correlation coefficient . We report the correlation of these DNA methylation measures between visits using the Pearson correlation coefficient and scatter plots. We compared the distribution of variables by exposure status to prenatal maternal smoking using the Welch's t-test for continuous variables and the Chi-Square test for categorical variables. For select continuous DNA methylation measures we compare the distribution across prenatal smoke exposure status using violin plots with captive box plots.

#### Regression analyses

To assess the association between our primary exposure variable of prenatal maternal smoking and salivary DNA methylation, we performed age-stratified multivariable linear regressions with the primary polymethlation score as our outcome. To facilitate interpretation, we z-score standardized the polymethylation score. We adjusted for an indicator variable for maternal residence in Chicago, Detroit or Toledo (over sampled cities), maternal income to poverty ratio at baseline, child sex, child age, plate from DNA methylation processing, estimated immune cell proportion from DNA methylation data, and the first two components of genetic ancestry from principal component analysis. While child sex, child age, plate from DNA methylation processing and immune cell proportions are not confounders (as they cannot casually effect prenatal maternal smoke exposure), these variables can strongly effect DNA methylation and so were adjusted for as precision variables. We reported the effect estimate for polymethylation score term and the 95% confidence interval (CI). We visualized our findings using forest plots.

To assess the portability of the association between our primary exposure variable of prenatal maternal smoking and salivary DNA methylation across genetic ancestry, we additionally stratified by genetic ancestry and reran the models described above. In ancestry-stratified models, the first two principal components from principal component analysis run within each ancestry strata were used.

To compare the utility of the polymethylation score to a single site biomarkers, we used percent methylation at AHRR:cg05575921 as the outcome in the models described above. To assess the specificity of the impact of prenatal maternal smoking on DNA methylation, we used mean percent global DNA methylation and the pediatric DNA methylation clock as outcome variables.

#### Regression sensitivity analyses

To assess the specificity of the DNA methylation signal to the type and timing of exposure, we additionally adjusted for the following variables in sensitivity analyses: 1) prenatal drug and alcohol use and 2) postnatal second-hand smoke exposure in addition to prenatal drug and alcohol use. To assess the robustness of the association between prenatal maternal smoking and DNA methylation to batch effects and residual confounding, we also performed regression analyses adjusting for surrogate variables calculated from the DNA methylation data.

To assess the robustness of the association to choice of polymethylation score, we performed the same regression analyses described above with z-score standardized sensitivity polymethylation scores as the outcome. Similarly, we replaced the outcome variable of *AHRR:*cg05575921 with other *a priori* selected DNA methylation sites in sensitivity analyses. We plotted the estimates and 95% CI for each DNA methylation site along with the previously reported estimates from the literature using forest plots [@Joubert:2016ue; @wiklund2019]

#### Receiver operator curve analysis

To evaluate the accuracy of the DNA methylation summary measures as biomarkers of prenatal maternal smoking, we used a receiver operating curve. First, we regressed exposure to prenatal maternal smoking (outcome) against the DNA methylation summary measures in individual logistic regressions, while adjusting for the base model variables listed above. Next, we calculated receiver operating curves (ROC) and area under the curves (AUCs) using the function roc from the R library pROC version 1.18.0 [@pROC2011]. We compared ROC curves and AUCs from models with only base model variables to those which included base model variables and the methylation summary measure using the Delong method and the function roc.test from the R library pROC version 1.18.0. This is appropriate as the base model variable only model is nested within the base model variable + methylation summary measures model.

#### Receiver operator curve sensitivity analyses

To assess the robustness of polymethylation scores classification accuracy to choice of coefficients, we used polymethylation scores calculated using different coefficient sets as predictors for the receiver operating curve analysis in sensitivity analyses.

To compare the performance of saliva-based polymethylation scores to previous reports of blood-based polymethylation scores, we used a forest plot. We plotted the AUC and 95% CI (when available) for each polymethylation score in our cohort and in the cohorts previously reported in the literature [@rauschert2020; @Richmond:2018tu].

Code to perform all analyses is available (www.github.com/bakulskilab)