### 2022-03-18 Papers/Polymethylation_scores/99_FigureCaptions.Rmd

---
output:
  html_document: default
  word_document: default
bibliography: Methylation_SmokingBib.bib
always_allow_html: true
---

```{r, echo=F, include=FALSE}
print_figs=TRUE
```

#### Table 1: Bivariate associations between prenatal maternal smoking (no/yes) and selected DNA methylation summary measures and important covariates among a diverse sample of `r n_total` children in the Fragile Families and Child Wellbeing study

```{r,  eval=print_figs, include=T}
smoking_tbl
```

#### **Figure 1 - Selection of samples from the Fragile Families and Child Wellbeing study into analytic subset.** N represents the number of individuals at each step in the selection procedure, M represents the number of samples. Individuals with repeated measures can have more than one sample.

```{r, echo=F, eval=print_figs, include=T, fig.height=11, fig.width=7}
graph1
```

#### **Figure 2 - Differences in selected DNA methylation summary measures by self-report of prenatal maternal smoking among `r n_total` children in the Fragile Families and Child Wellbeing study at ages 9 and 15.** Samples from children exposed to prenatal maternal smoke in grey, samples from children unexposed to prenatal maternal smoke in black. From top-left, clockwise: Polymethylation scores for prenatal maternal smoke exposure, constructed using regression coefficients for prenatal smoke exposure predicting DNA methylation in newborn cordblood samples, accounting for cell-type control. DNA methylation values from samples in the Fragile Families and ChildWellbeing study were mean-centered, then multiplied by these regression coefficients and summed. Pediatric epigentic clock (years). AHRR gene: percent cg05575921 methylation. Percent global DNA methylation.

```{r, echo=F, eval=print_figs, include=T, fig.height=5, fig.width=7}
fig2
```

#### **Figure 3 - Prenatal maternal smoke exposure is consistently associated with polymethylation scores at ages 9 and 15 and is portable across genetic ancestry groups in a sample of `r n_total` children in the Fragile Families and Child Wellbeing study** All models shown controlled for: first two principal components of child genetic ancestry (from ancestry-stratified principal components for ancestry stratified models), child sex, maternal income-to-poverty ratio at birth, immune cell proportion estimated from methylation data, yes/no other maternal prenatal drug use, yes/no maternal prenatal alcohol use, postnatal maternal smoking when child age 1 or age 5, postnatal maternal/primary care give smoking packs/day in month prior to saliva sample.

```{r, echo=F, eval=print_figs, include=T, fig.width=12}
forest_plot
```

####  **Figure 4 - Polymethylation scores accurately classify prenatal maternal smoke exposure at ages 9 and 15 among `r n_total` children in the Fragile Families and Child Wellbeing study** A) Receiver operator curve for select DNA methylation measures for predicting prenatal smoke exposure using no other variables (light colors) or using base model variables (dark colors, other variables included: child sex, maternal income-poverty ratio at birth, immune cell proportion and batch of methylation data processing). B) Receiver operator curve for include polymethylation scores individually at each visit (black & light grey) or jointly (dark grey).

```{r, echo=F, eval=print_figs, include=T,  fig.width=12}
auc_fig
```

#### **Figure 5 - Comparison of different polymethylation scores and their accuracy at classifying prenatal smoke exposure among `r n_total` children with saliva samples in the Fragile Families and Child Wellbeing study as opposed to previously reported results in cord and peripheral blood samples

```{r, echo=F, eval=print_figs, include=T,  fig.width=12}
auc_compare
```