### 2022-03-18 Papers/Polymethylation_scores/index_paper.Rmd

---
title: "PAPER DRAFT"
author: '------------'
date: "`r Sys.Date()`"
output:
  word_document: 
    reference_docx: "../Template.docx"
  html_document:
    df_print: paged
  pdf_document: default
bibliography: [Methylation_SmokingBib.bib, packages.bib]
csl: springer-vancouver.csl
always_allow_html: yes
---

```{r, echo=F, include=F, message=F, warning=F}
library(stringr)
library(knitr)
library(tidyverse)
library(tidyr)
library(dplyr)

coefCI=function(df, filter_statement, my_term, si){
  my_df=df%>%
    filter(eval(parse(text = filter_statement)))%>%
    unnest(tidied)%>%
    mutate(val=paste0(round(estimate, si), ' (95%CI: ', 
                      round(conf.low, si), ', ',
                      round(conf.high, si), ')'))%>%
    filter(term==my_term)%>%
    pull(val)
  return(my_df)
}
```

```{r load data, echo=F, include=F, warning=F, message=F, cache=TRUE, cache.lazy=F}
results_run=str_detect(here::here(), 'nfs/turbo/bakulski1')
knitr::knit_child(here::here("Reports", "Paper1.Rmd"))
```

```{r, child = here::here("Papers", "Polymethylation_scores", "01_titlepage.Rmd")}
```

```{r, child = here::here("Papers", "Polymethylation_scores", "02_abstract.Rmd")}
```

```{r, child = here::here("Papers", "Polymethylation_scores", "03_keywords.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "04_background.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "08_methods.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "05_results.Rmd"), eval=results_run}
```

```{r child = here::here("Papers", "Polymethylation_scores", "06_discussion.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "07_conclusions.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "98_declarations.Rmd")}
```

```{r child = here::here("Papers", "Polymethylation_scores", "99_FigureCaptions.Rmd"), eval=results_run}
```

# References