### 2022-04-29 Papers/Polymethylation_scores/04_background.Rmd

---
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: Methylation_SmokingBib.bib
latex_engine: xelatex
csl: springer-vancouver.csl
---

# Introduction

Children exposed to cigarette smoking *in utero* are more likely to have low-birth weight, negative neurodevelopmental outcomes, and asthma [@Banderali:2015vu]. Although 11% of US pregnant people in 2018 reported any past-month cigarette use [@SAMHSA2018], this is considered an underestimate. When compared to biomarker measures, maternal prenatal smoking is underreported - presumably due to stigma [@Kvalvik:2012tk; @Dietz:2011ua]. The gold-standard smoking biomarker, serum cotinine levels, has a half-life of nine hours in pregnant people, and is rarely available outside of birth cohort research [@Dempsey:2002wr]. Alternative biomarkers of prenatal maternal smoke exposure with a larger window of measurement could improve measurement of prenatal smoke exposure in older children and adults.

DNA methylation is a possible biomarker of prenatal smoke exposure. Smoking during pregnancy is associated with offspring DNA methylation signatures at birth in placental and cord blood samples, and postnatally in peripheral blood samples [@Joubert:2016ue; @Everson:2021td; @Cardenas2019; @Richmond:2018tu; @Tehranifar:2018tb]. Another postnatal biospecimen option is saliva, which is easier to collect than blood and increasingly collected in large epidemiological cohorts [@Langie2017]. However, associations of pregnancy smoking with offspring salivary DNA methylation have not been measured. Since DNA methylation contributes to cell differentiation and cell type proportions differ across tissue types, it is important to verify associations in new tissues [@Bakulski2016]. In addition, most studies of prenatal maternal smoking and childhood DNA methylation have been conducted in primarily European genetic ancestry populations [@Joubert:2016ue; @Ladd-Acosta:2016aa; @Tehranifar:2018tb; @Zhang2018; @Gonseth2016; @Neophytou:2019ty], and it will be important to test generalizability to other groups. Few studies have assessed multiple genetic ancestry groups in the same cohort and no study has evaluated associations between pregnancy cigarette smoking and salivary DNA methylation in children.

Using the findings from large-scale association testing in independent populations, DNA methylation-based exposure biomarkers can be developed and assessed by evaluating their classification accuracy. Biomarkers can range from single DNA methylation sites to summary measures across multiple sites, termed polymethylation scores [@Philibert2020 @Philibert2013; @Reese:2017ts; @Richmond:2018tu; @Ladd-Acosta:2016aa]. In a polymethylation score, DNA methylation sites are weighted by their effect size from a previously conducted analysis in an independent sample and then summed to a single score [@Chen2020; @Bakulski2021a]. DNA polymethylation scores for *in utero* smoke exposure have performed well (area under the curves (AUCs)\>0.8) in cord blood samples from newborns and peripheral blood samples from young children [@Reese:2017ts; @Ladd-Acosta:2016aa]. Polymethylation scores have also classified *in utero* smoke exposure when using from DNA methylation in peripheral blood from middle-aged adults (AUC of 0.72 (95% confidence interval 0.69, 0.76)) [@Richmond:2018tu]. Yet few studies have assessed the longitudinal persistence of DNA-methylation-based smoking biomarker accuracy.

DNA methylation differences hold promise as an exposure biomarker for *in utero* cigarette smoke. We must understand how the signal varies across tissue, genetic ancestry, and age. In the Fragile Families and Child Wellbeing Study, a diverse longitudinal birth cohort, we aimed to assess salivary DNA methylation biomarkers for pregnancy cigarette smoking. We tested associations between pregnancy smoke exposure, measured at ages 9 and 15. We compared the magnitude and specificity of these signals to other DNA methylation measures including individual a priori DNA methylation sites, global DNA methylation, and epigenetic clocks. We then performed genetic ancestry- and age-stratified analyses to test the hypothesis that DNA methylation could serve as a portable and persistent biomarker of prenatal maternal smoking.